---
title: "policy_learning_nitrate_regional"
author: "Xindi Lin"
date: "2025-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
library(ranger)
library(SuperLearner)
library(sf)
library(dplyr)
library(ggplot2)
library(WeightIt)
library(GpGp)
library(Metrics)
library(rnaturalearth)
library(ggplot2)
library(see)
library(yardstick)
library(xgboost)
library(glmnet)
library(kernlab)
library(glmnet)
library(numDeriv)
library(MASS)
```

```{r}
source('data_generation.R')
source('funcs_kriging.R')
source('loss_functions.R')
source('loss_functions_RKHS.R')
```

#### load or preprocess data

```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv"
## we will use the one with the proportion
# file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_median_well"
command <- paste("brctl download", shQuote(file_path))
system(command)
data <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv")
```

#### Add the newly constructed covariates: landuse proportion

```{r}
df_plss_sections_landuse_prop <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/plss_sections_landuse_prop.csv")
```

```{r}
data <- left_join(data,df_plss_sections_landuse_prop,by = c("Range.Direction", "Township", "Range", "Section")) 
```

```{r}
write.csv(data, "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_median_well.csv",row.names = FALSE)
```

```{r}
data <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_median_well.csv")
```


I hope these two columns are the same, but it turned out that they are very different.
If everything works as what I want to do in the data collection, covariate construction process, I think the land use proportion is more accurate. Because when we construct combine, we decrease the resolution of the rasters, and then match the PLSS section.

```{r}
cbind(data$crop_type_combine, colnames(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])[max.col(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])])
```

```{r}
# visual check
colnames(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])[max.col(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])][21:50]

data[21:50,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")]
```


But I will visualize the dominant proportion

```{r}
data$max_landuse_prop = colnames(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])[max.col(data[,c("Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined")])]
```

```{r}
usa_states <- ne_states(country = "united states of america", returnclass = "sf")
wisconsin <- usa_states[usa_states$name == "Wisconsin", ]
```

```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "white", color = "black") +  # Wisconsin boundary
  geom_sf(data = st_as_sf(data, coords = c('longitude','latitude'), crs = 4326), aes(color= max_landuse_prop),size=0.1) +
  scale_color_material()
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()
```

I think it looks reasonable


#### check well Use

```{r}
table(data$WellUse)
```
10mg/L for private wells

5mg/L for public wells, maybe set 5mg/L as the threshold. So the imbalanced issue might be mitigated. 

##### Add historical data as a covariate

Separate data into most recent measurement/the historical measurement at every location (grid). 

Take average/median historical of neighbors in township/range. 

We will first separate the historical data and most recent data

```{r}
# only by well depth
latest_records <- data %>%
  # Make sure the date column is a true Date type for correct sorting
  mutate(SampleDate = as.Date(SampleDate)) %>%
  # Group the data by the well ID
  group_by(WellID) %>%
  # Within each group, keep only the row with the maximum date
  slice_max(order_by = SampleDate, n = 1) %>%
  # Optional: remove the grouping
  ungroup()
```

```{r}
dim(latest_records)
```

keep the latest record by "welldepth", "latitude", "longitude", "WellUse"

```{r}
# by sample date
latest_records_depth <- latest_records %>%
  # Ensure the date column is a true Date type
  mutate(SampleDate = as.Date(SampleDate)) %>%
  # Group by all four identifying columns
  group_by(WellDepth, latitude, longitude) %>%
  # Keep the row with the latest date within each group
  slice_max(order_by = SampleDate, n = 1) %>%
  ungroup()

## by sample year
# latest_records_depth <- latest_records %>%
#   # Ensure the date column is a true Date type
#   # mutate(SampleDate = as.Date(SampleDate)) %>%
#   # Group by all four identifying columns
#   group_by(WellDepth, latitude, longitude) %>%
#   # Keep the row with the latest date within each group
#   slice_max(order_by = SampleYear, n = 1) %>%
#   ungroup()
```

We will then find the median/mean of the historical nitrate levels within each township/range

```{r}
data_latest_oldest <- data %>% 
  mutate(SampleDate = as.Date(SampleDate)) %>%
  group_by(WellDepth, latitude, longitude, WellUse) %>%
  mutate(sampling_order = row_number(desc(SampleDate))) %>% 
  arrange(latitude, longitude, WellDepth, sampling_order) %>% 
  ungroup()
```

```{r}
data_latest <- data_latest_oldest[data_latest_oldest$sampling_order == 1,]
data_historical <- data_latest_oldest[data_latest_oldest$sampling_order != 1,]
```

Then, we aggregate the historical result in every township/

```{r}
# data_historical_agg <-  data_historical %>% 
#   group_by(Township, Range, Range.Direction) %>% 
#   mutate(Median_Historical = median(concentration_plus_median)) 

data_historical_agg <- data_historical %>%
  group_by(Township, Range, Range.Direction) %>%
  summarise(
    Median_Historical = median(concentration_plus_median),
    Count = n(),
    .groups = 'drop'  # This is good practice to ungroup the result
  )
```

We add the aggregated median to the original dataset

```{r}
data_latest <- data_latest %>% 
  left_join(data_historical_agg %>% 
              select(Township, Range, Range.Direction, Median_Historical), 
            by = c("Township", "Range", "Range.Direction"))
```

Some blocks do not have historical data. We can simply drop them.

```{r}
sum(is.na(data_latest$Median_Historical))
```

```{r}
data_latest <- na.omit(data_latest)
```

```{r}
write.csv(data_latest, "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_histLevel_median_well.csv",row.names = FALSE)
```

```{r}
data_latest <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_histLevel_median_well.csv")
```


```{r}
data <- data_latest
```

#### remove zero-inflated

```{r}
data <- data[data$concentration_plus_median!=0.5,]
```

#### Choose Area

```{r}
sort(unique(data$County))
```

```{r}
data$area = ""

indices_central <- data$County %in% c("Wood", "Portage", "Waupaca", "Juneau", "Adams", "Waushara", "Marquette", "Green Lake")
data[indices_central, "area"] = "Central"

indices_westCentral <- data$County %in% c("Buffalo", "Trempealeau", "La Crosse", "Monroe")
data[indices_westCentral, "area"] = "West Central"

indices_southCentral <- data$County %in% c("Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock")
data[indices_southCentral, "area"] = "South Central"

indices_eastCentral <- data$County %in% c("Outagamie", "Winnebago", "Door", "Fond du Lac", "Brown", "Calumet","Sheboygan","Manitowoc","Kewaunee")
data[indices_eastCentral, "area"] = "East Central"

indices_southWest <- data$County %in% c("Vernon", "Crawford", "Grant", "Richland", "Sauk", "Iowa","Lafayette")
data[indices_southWest, "area"] = "South West"

indices_southEast <- data$County %in% c("Washington", "Waukesha", "Walworth", "Ozaukee", "Milwaukee", "Racine","Kenosha")
data[indices_southEast, "area"] = "South East"
```

```{r}
table(data$area)
```


```{r}
data <- data[data$area == "Central",]
data <- data[data$area == "East Central",] # choose this
data <- data[data$area == "South Central",]
data <- data[data$area == "South East",]
data <- data[data$area == "South West",]
data <- data[data$area == "West Central",]
```

#### Visualization

##### Distribution of Contaminants and Depth

```{r}
plot(density(log(data$WellDepth)))
plot(density(data$WellDepth))
# so we can roughly assume the log of wellDepth, i.e., the treatment, is normally distributed.
```

```{r}
hist(data[data$WellUse %in% c("PRIVATE NON-POTABLE", "PRIVATE POTABLE"), "concentration_plus_median"], main = 'private wells')
```

```{r}
plot(density(log(data$concentration_plus_median)))
```

```{r}
plot(density(log(data$concentration_plus_median[data$concentration_plus_median!=0.5])))
```


#### Down/sub sampling


#### Split to Training and Test Data

```{r}
set.seed(1998)

data_full <- data
train_indices <- sample(nrow(data_full), nrow(data_full) * 0.7)
data <- data_full[train_indices,]
data_test <- data_full[-train_indices,]
```

We will define a function to implement the training/test data

#### Estimate nuisance parameter

```{r}
data$logWellDepth = log(data$WellDepth)
data$logconcentration_plus_median = log(data$concentration_plus_median)
data_test$logWellDepth = log(data_test$WellDepth)
data_test$logconcentration_plus_median = log(data_test$concentration_plus_median)
```

##### using defined function to collect result comparing data/covariates

```{r}
# data_full <- data
```

```{r}
unique(data_full$area)
```

```{r}
covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")
sub_region <- "South West"
seed <- 1998
```


```{r}
training_test_function <- function(data_full, 
                                   # spatial = c("GP","Coords","None")
                                   covariate_names, 
                                   sub_region = c("Whole","Central","South West","West Central","South Central","South East","East Central"),
                                   # method = c('xgboost'), 
                                   seed = 1998){
  
  
  sub_region <- match.arg(sub_region)
  ## Split to Training and Test Data
  data <- data_full[data_full$area == sub_region,]
  set.seed(seed)
  train_indices <- sample(nrow(data), nrow(data) * 0.7)
  data_train <- data[train_indices,]
  data_test <- data[-train_indices,]
  
  ## xgboost 
  ### get design matrix. generate two design matrix separately might incur different column number if one category
  
  # 1. Define a recipe that learns from the FULL dataset
  my_recipe <- recipe(~ ., data = data[, covariate_names]) %>%
  step_dummy(all_nominal_predictors()) # This creates the dummy variables
  
  # 2. "Prep" the recipe. This learns and finalizes the rules (including all factor levels).
  trained_recipe <- prep(my_recipe)
  
  # 3. "Bake" (apply) the trained recipe to your subsets
  design_matrix <- as.matrix(bake(trained_recipe, new_data = data_train[, covariate_names]))
  design_matrix_test <- as.matrix(bake(trained_recipe, new_data = data_test[, covariate_names]))
  
  # design_matrix <- model.matrix(~ . , data = data_train[, covariate_names])
  # design_matrix_test <- model.matrix(~ . , data = data_test[, covariate_names])
  
  ### select nrounds
  dtrain <- xgb.DMatrix(data = design_matrix, label = data_train$logconcentration_plus_median)

  params <- list(
    objective = "reg:squarederror", # Specify regression task
    eta = 0.1,                      # Learning rate
    max_depth = 4,                  # Maximum depth of a tree
    monotone_constraints = c(0, -1, rep(0,15)), # adding this actually improve the performance
    eval_metric = "rmse"            # The metric to evaluate
  )

  cv_results <- xgb.cv(
    params = params,
    data = dtrain,
    nrounds = 400,
    nfold = 5,
    early_stopping_rounds = 50, # Stop if performance doesn't improve
    verbose = 0                 # Suppress verbose output
  )
  
  ### train the model
  
  OR_mean_model_xgboost <- xgboost(data = design_matrix, 
                                   label = data_train$logconcentration_plus_median,
                                   eta = 0.1,                      # Learning rate
                                   max_depth = 4, 
                                   nrounds = cv_results$best_iteration,
                                   monotone_constraints = c(0, -1, rep(0,15)))
  
  ## GP Regression of Residual
  
  ### get cross-validation residual
  loocv_model <- xgb.cv(
    data = dtrain,
    nrounds = cv_results$best_iteration,
    eta = 0.1,                      # Learning rate
    max_depth = 4, 
    monotone_constraints = c(0, -1, rep(0,15)),
    nfold = 10,                 # Key step for LOOCV
    prediction = TRUE,         # Important: tells the function to return predictions
    verbose = 0                # Suppress progress messages
  )
  
  gpfit = GpGp::fit_model(data_train$logconcentration_plus_median - loocv_model$pred,
                        locs = data_train[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")

  krige_values <- leave_one_out_kriging(locs = data_train[,c("longitude", "latitude")], 
                                      y_obs = data_train$logconcentration_plus_median - loocv_model$pred, 
                                      gp_model = gpfit$covfun_name,
                                      gp_params = gpfit$covparms,
                                      order = c("coordinate")) 

  krige_values_test <- GpGp::predictions(fit = gpfit, 
                       locs_pred = data_test[,c("longitude", "latitude")], 
                       X_pred = rep(1,nrow(data_test)))
  
  ## calculate the training/prediction rMSE
  
  ### non-spatial rMSE
  rmse_train_nonsp <- Metrics::rmse(data_train$logconcentration_plus_median, predict(OR_mean_model_xgboost, design_matrix))
  rmse_test_nonsp <- Metrics::rmse(data_test$logconcentration_plus_median, predict(OR_mean_model_xgboost, design_matrix_test))
  
  ### spatial rMSE
  rmse_train_sp <- Metrics::rmse(data_train$logconcentration_plus_median, predict(OR_mean_model_xgboost, design_matrix) + krige_values)
  rmse_test_sp <- Metrics::rmse(data_test$logconcentration_plus_median, predict(OR_mean_model_xgboost, design_matrix_test) + krige_values_test)
  df_result <- data.frame('Geographic Scope' = rep(sub_region, 2), 'Model Type' = c("With Kriging","No Kriging"),
             'Training rMSE' = c( rmse_train_sp, rmse_train_nonsp),
             'Test rMSE' = c(rmse_test_sp, rmse_test_nonsp))
  return(df_result)
}  
```


```{r}
data_whole <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_median_well.csv")
```

```{r}
data_whole$area = ""

indices_central <- data_whole$County %in% c("Wood", "Portage", "Waupaca", "Juneau", "Adams", "Waushara", "Marquette", "Green Lake")
data_whole[indices_central, "area"] = "Central"

indices_westCentral <- data_whole$County %in% c("Buffalo", "Trempealeau", "La Crosse", "Monroe")
data_whole[indices_westCentral, "area"] = "West Central"

indices_southCentral <- data_whole$County %in% c("Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock")
data_whole[indices_southCentral, "area"] = "South Central"

indices_eastCentral <- data_whole$County %in% c("Outagamie", "Winnebago", "Door", "Fond du Lac", "Brown", "Calumet","Sheboygan","Manitowoc","Kewaunee")
data_whole[indices_eastCentral, "area"] = "East Central"

indices_southWest <- data_whole$County %in% c("Vernon", "Crawford", "Grant", "Richland", "Sauk", "Iowa","Lafayette")
data_whole[indices_southWest, "area"] = "South West"

indices_southEast <- data_whole$County %in% c("Washington", "Waukesha", "Walworth", "Ozaukee", "Milwaukee", "Racine","Kenosha")
data_whole[indices_southEast, "area"] = "South East"
```


```{r}
data_whole$logWellDepth = log(data_whole$WellDepth)
data_whole$logconcentration_plus_median = log(data_whole$concentration_plus_median)
```

for data with prop covariate

```{r}
for(sub_region in c("Central","South West","West Central","South Central","South East","East Central")){
  df_result <- training_test_function(data_full = data_whole, 
                                      covariate_names = c("logWellDepth","WellUse","Corn","Soybeans",
                                                          "Vegtables","Other.Crop","Cover.Crop",
                                                          "Grass","Forest","Developed","drainagecl",
                                                          "precipitation","cafolog","StaticLevel"), 
                                      sub_region = sub_region)
  print(df_result)
  if(sub_region == "Central"){
    whole_data_prop_result <- df_result
  } else {
    whole_data_prop_result <- rbind(whole_data_prop_result, df_result)
  }
}

```
```{r}
whole_data_prop_result
```

For data with combined crop

```{r}
for(sub_region in c("Central","South West","West Central","South Central","South East","East Central")){
  df_result <- training_test_function(data_full = data_whole, 
                                      covariate_names = covariate_names, 
                                      sub_region = sub_region)
  print(df_result)
  if(sub_region == "Central"){
    whole_data_result <- df_result
  } else {
    whole_data_result <- rbind(whole_data_result, df_result)
  }
}

```
```{r}
whole_data_result
```

Now we want to get result for latest data
```{r}
data_latest$area = ""

indices_central <- data_latest$County %in% c("Wood", "Portage", "Waupaca", "Juneau", "Adams", "Waushara", "Marquette", "Green Lake")
data_latest[indices_central, "area"] = "Central"

indices_westCentral <- data_latest$County %in% c("Buffalo", "Trempealeau", "La Crosse", "Monroe")
data_latest[indices_westCentral, "area"] = "West Central"

indices_southCentral <- data_latest$County %in% c("Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock")
data_latest[indices_southCentral, "area"] = "South Central"

indices_eastCentral <- data_latest$County %in% c("Outagamie", "Winnebago", "Door", "Fond du Lac", "Brown", "Calumet","Sheboygan","Manitowoc","Kewaunee")
data_latest[indices_eastCentral, "area"] = "East Central"

indices_southWest <- data_latest$County %in% c("Vernon", "Crawford", "Grant", "Richland", "Sauk", "Iowa","Lafayette")
data_latest[indices_southWest, "area"] = "South West"

indices_southEast <- data_latest$County %in% c("Washington", "Waukesha", "Walworth", "Ozaukee", "Milwaukee", "Racine","Kenosha")
data_latest[indices_southEast, "area"] = "South East"
```

```{r}
data_latest$logWellDepth = log(data_latest$WellDepth)
data_latest$logconcentration_plus_median = log(data_latest$concentration_plus_median)
```


for crop mode

```{r}
for(sub_region in c("Central","South West","West Central","South Central","South East","East Central")){
  df_result <- training_test_function(data_full = data_latest, 
                                      covariate_names = c("logWellDepth","WellUse","crop_type_combine",
                                                          "drainagecl","precipitation","cafolog",
                                                          "StaticLevel", "Median_Historical"), 
                                      sub_region = sub_region)
  print(df_result)
  if(sub_region == "Central"){
    recent_data_result <- df_result
  } else {
    recent_data_result <- rbind(recent_data_result, df_result)
  }
}

```
```{r}
recent_data_result
```

for crop proportion

```{r}
data_latest <- left_join(data_latest, df_plss_sections_landuse_prop,by = c("Range.Direction", "Township", "Range", "Section")) 
```

```{r}
for(sub_region in c("West Central","Central","South West","South Central","South East","East Central")){
  df_result <- training_test_function(data_full = data_latest, 
                                      covariate_names = c("logWellDepth","WellUse","Corn","Soybeans",
                                                          "Vegtables","Other.Crop","Cover.Crop",
                                                          "Grass","Forest","Developed","drainagecl",
                                                          "precipitation","cafolog","StaticLevel", "Median_Historical"),
                                      sub_region = sub_region)
  print(df_result)
  if(sub_region == "West Central"){
    recent_data_prop_result <- df_result
  } else {
    recent_data_prop_result <- rbind(recent_data_prop_result, df_result)
  }
}

```




##### Test if Land Use Proportion Improve the Result

```{r}
OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")
# we drop the proportion of undefined to avoid colinearility with other proportion
OR_covariate_names_prop <- c("logWellDepth","WellUse","Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","drainagecl","precipitation","cafolog","StaticLevel")
```

superlearner is extremely slow, I guess it is cross-validation in SuperLearner

```{r}
OR_mean_model <- SuperLearner(
  Y = data$logconcentration_plus_median,
  X = data[, OR_covariate_names],
  family = gaussian(),
  SL.library = c("SL.ranger")
)

OR_mean_prop_model <- SuperLearner(
  Y = data$logconcentration_plus_median,
  X = data[, OR_covariate_names_prop],
  family = gaussian(),
  SL.library = c("SL.ranger")
)
```

compare training loss plot

```{r}
plot(log(data$concentration_plus_median), OR_mean_model$library.predict, cex = 0.1)
abline(0,1,col='red')

plot(log(data$concentration_plus_median), OR_mean_prop_model$library.predict, cex = 0.1)
abline(0,1,col='red')
```

compare training rMSE

```{r}
Metrics::rmse(log(data$concentration_plus_median), OR_mean_model$library.predict)

Metrics::rmse(log(data$concentration_plus_median), OR_mean_prop_model$library.predict)
```

compare test plot

```{r}
plot(log(data_test$concentration_plus_median), predict(OR_mean_model, newdata = data_test)$pred)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), predict(OR_mean_prop_model, newdata = data_test)$pred)
abline(0,1,col='red')
```

compare test rmse

```{r}
plot(log(data_test$concentration_plus_median), predict(OR_mean_model, newdata = data_test)$pred)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), predict(OR_mean_prop_model, newdata = data_test)$pred)
abline(0,1,col='red')
```

```{r}
plot(log(data_test$concentration_plus_median), test_OR_pred)
abline(0,1,col='red')
```



##### Cross-validated kernel regression

##### Estimate Outcome Regression: kernel ridge regression (RKHS)

###### Using covariate to form the grand matrix

####### Manually

```{r}
library(kernlab)
library(glmnet)
```

```{r}
OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")
```

```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])[,-1]
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])[,-1]
```

```{r}
train_means <- colMeans(design_matrix)
train_sds <- apply(design_matrix, 2, sd)

design_matrix <- scale(design_matrix)
design_matrix_test <- scale(design_matrix_test, center = train_means, scale = train_sds)
```

```{r}
m <- - 1
gamma <- 2^(m) * median(rdist(design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))
k_matrix <- kernelMatrix(rbf, design_matrix)
# build kernel matrix for test data
k_matrix_test <- kernelMatrix(rbf, design_matrix_test,design_matrix)

# k_matrix <- as.numeric(k_matrix)
# k_matrix_test <- as.numeric(k_matrix_test)
```

```{r}
hist(k_matrix[col(k_matrix)!=row(k_matrix)],xlim = c(0,1))
```


glmnet will automatically add the intercept for us, we we don't need to add the intercept

```{r}
OR_mean_model_RKHS.cv <- cv.glmnet(x = k_matrix, y = data$logconcentration_plus_median, alpha = 0, keep = TRUE)

OR_mean_model_RKHS <- glmnet(x = k_matrix, y = data$logconcentration_plus_median, alpha = 0, lambda = OR_mean_model_RKHS.cv$lambda.min)
```

```{r}
OR_pred_RKHS <- predict(OR_mean_model_RKHS, newx = k_matrix, s=OR_mean_model_RKHS.cv$lambda.min)
OR_pred_RKHS_test <- predict(OR_mean_model_RKHS, newx = k_matrix_test, s=OR_mean_model_RKHS.cv$lambda.min)
```

```{r}
plot(data$logconcentration_plus_median, OR_pred_RKHS, cex = 0.2)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), OR_pred_RKHS_test,cex = 0.2)
abline(0,1,col='red')

Metrics::rmse(data$logconcentration_plus_median, OR_pred_RKHS)
Metrics::rmse(log(data_test$concentration_plus_median), OR_pred_RKHS_test)
```


```{r}
cor(log(data_test$concentration_plus_median), OR_pred_RKHS_test)
```


####### using package


```{r}
library(e1071)
```




```{r}
OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")
```

```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])[,-1]
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])[,-1]
```

```{r}
svm_auto <- svm(x = design_matrix, y = data$logconcentration_plus_median, type='nu-regression')
```

```{r}
plot(data$logconcentration_plus_median, predict(svm_auto, newdata = design_matrix), cex = 0.2)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test), cex = 0.2)
abline(0,1,col='red')

Metrics::rmse(data$logconcentration_plus_median, predict(svm_auto, newdata = design_matrix))
Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test))
```

```{r}
svm_tunning <- e1071::tune.svm(x = design_matrix, y = data$logconcentration_plus_median)
```

```{r}
plot(data$logconcentration_plus_median, predict(svm_tunning$best.model, newdata = design_matrix), cex = 0.2)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), predict(svm_tunning$best.model, newdata = design_matrix_test), cex = 0.2)
abline(0,1,col='red')

Metrics::rmse(data$logconcentration_plus_median, predict(svm_tunning$best.model, newdata = design_matrix))
Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_tunning$best.model, newdata = design_matrix_test))
```


```{r}
predict(svm_tunning$best.model, newdata = design_matrix_test)
```

```{r}
library(gKRLS)
```

```{r}
gam(y ~ s(x1, x2, bs = "krr", k = "rbf"), data = data)
```

###### Using coordinates to form the grand matrix

```{r}

```


##### Estimate Residual using Gaussian Process: RKHS


```{r}
# gpfit_RKHS = GpGp::fit_model(data$logconcentration_plus_median -
#                                OR_mean_model_RKHS.cv$fit.preval[,which(OR_mean_model_RKHS.cv$lambda==OR_mean_model_RKHS.cv$lambda.min)],
#                         locs = data[,c("longitude", "latitude")],
#                         covfun_name = "matern_sphere")

gpfit_RKHS = GpGp::fit_model(svm_auto$residuals,
                        locs = data[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")
```

```{r}
# krige_values_RKHS <- leave_one_out_kriging(locs = data[,c("longitude", "latitude")], 
#                                       y_obs = data$logconcentration_plus_median -
#                                OR_mean_model_RKHS.cv$fit.preval[,which(OR_mean_model_RKHS.cv$lambda==OR_mean_model_RKHS.cv$lambda.min)], 
#                                       gp_model = gpfit_RKHS$covfun_name,
#                                       gp_params = gpfit_RKHS$covparms,
#                                       order = c("coordinate")) 

krige_values_RKHS <- leave_one_out_kriging(locs = data[,c("longitude", "latitude")], 
                                      y_obs = svm_auto$residuals, 
                                      gp_model = gpfit_RKHS$covfun_name,
                                      gp_params = gpfit_RKHS$covparms,
                                      order = c("coordinate")) 
```

```{r}
krige_values_test_RKHS <- GpGp::predictions(fit = gpfit_RKHS, 
                      locs_pred = data_test[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_test)))
```

```{r}
cat('Test rMSE: with kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test) + krige_values_test_RKHS))
cat('\n')
cat('Test rMSE: No kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test)))
cat('\n')

cat('Training rMSE: with kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), predict(svm_auto, newdata = design_matrix) + krige_values_RKHS))
cat('\n')
cat('Training rMSE: No kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), predict(svm_auto, newdata = design_matrix)))
cat('\n')
```

```{r}
cat('Test rMSE: with kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), OR_pred_RKHS_test + krige_values_test_RKHS))
cat('\n')
cat('Test rMSE: No kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), OR_pred_RKHS_test))
cat('\n')

cat('Training rMSE: with kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), OR_pred_RKHS + krige_values_RKHS))
cat('\n')
cat('Training rMSE: No kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), OR_pred_RKHS))
cat('\n')
```
The improvement is tremendous for RKHS

```{r}
plot(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test) + krige_values_test_RKHS,cex = 0.2, main = 'kriging test prediction')
abline(0,1,col='red')
plot(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test),cex = 0.2, main = ' no kriging test prediction')
abline(0,1,col='red')
```

```{r}
plot(log(data_test$concentration_plus_median), OR_pred_RKHS_test + krige_values_test_RKHS,cex = 0.2, main = 'kriging test prediction')
abline(0,1,col='red')
plot(log(data_test$concentration_plus_median), OR_pred_RKHS_test,cex = 0.2, main = ' no kriging test prediction')
abline(0,1,col='red')
```


It is not easier to predict in small area. It actually depends on which areas we choose
The GP regression fails in small area. It actually depends on which areas we choose.

##### Compute OR smoothers: RKHS 

This step is very cumbersome for mannually SVM. We might directly use SVM

```{r}
boundary_tol <- 0
depth_range <- c(min(data$logWellDepth) - boundary_tol, max(data$logWellDepth) + boundary_tol)
depth_range
```

```{r}
# xgboost use the design matrix
cf_predictions_RKHS <- precompute_cf_predictions(data_obs = design_matrix, 
                              fit_outcome = svm_auto,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
cf_predictions_RKHS <- isotomic_correction(treatment_grid = cf_predictions_RKHS$treatment_values,
                                          pred_matrix_cf = cf_predictions_RKHS$pred_matrix)
```

```{r}
smoothers_RKHS <- make_prediction_smoothers(treatment_grid = cf_predictions_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_RKHS$pred_matrix,
                                            # smooth_method = "gam",
                                            # smooth_method = "smooth.spline",
                                            smooth_method = "splinefun",
                                            spar = NULL)
```

```{r}
cumint_smoothers_RKHS <- make_cumint_smoothers(treatment_grid = cf_predictions_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_RKHS$pred_matrix,
                                         smooth_method = "splinefun",
                                         # smooth_method = "gam",
                                         # smooth_method = "smooth.spline",
                                         spar = NULL)
```

```{r}
# xgboost use the design matrix
cf_predictions_test_RKHS <- precompute_cf_predictions(data_obs = design_matrix_test, 
                              fit_outcome = svm_auto,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
cf_predictions_test_RKHS <- isotomic_correction(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                          pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix)
```

```{r}
smoothers_test_RKHS <- make_prediction_smoothers(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix,
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers_test_RKHS <- make_cumint_smoothers(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                               pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix,
                                               # smooth_method = "splinefun",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```


We can clearly see that some of it is not monotonically decreasing, so we add iso

```{r}
i <- 11
# when i = 20, it decrease first then increase, 
plot(seq(0,8,length.out = 100), sapply(seq(0,8,length.out = 100), function(knot) {smoothers_RKHS[[i]](knot)}), main = 'Smooth')
plot(cf_predictions_RKHS$treatment_values, sapply(cf_predictions_RKHS$treatment_values, function(knot) {smoothers_RKHS[[i]](knot)}), main = 'Smoothers RKHS')

plot(cf_predictions_RKHS$treatment_values, sapply(cf_predictions_RKHS$treatment_values, function(knot) {cumint_smoothers_RKHS[[i]](knot)}), main = 'cumint Smoothers RKHS')

plot(cf_predictions_RKHS$treatment_values,  cf_predictions_RKHS$pred_matrix[,i], main = 'Non-smooth original')
```

```{r}
mean(sapply(1:nrow(data), function(i){smoothers_RKHS[[i]](depth_range[1])>= log(5)}))
mean(sapply(1:nrow(data), function(i){smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] >= log(5)}))
```

```{r}
mean(data$logconcentration_plus_median >= log(5))
mean(data$logconcentration_plus_median >= log(2))
```



##### Estimate Outcome Regression: XGBoost

```{r}
OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")

# OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel","Median_Historical")

# OR_covariate_names_prop <- c("logWellDepth","WellUse","Corn","Soybeans","Vegtables","Other.Crop","Cover.Crop","Grass","Forest","Developed","Undefined","drainagecl","precipitation","cafolog","StaticLevel","Median_Historical")

# OR_covariate_names <- c("logWellDepth","precipitation","cafolog","crop_type_combine","drainagecl")
```

Adding latitude, longitude does not change the prediction result a lot.

```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])
```

we might do cross-validation 

```{r}
dtrain <- xgb.DMatrix(data = design_matrix, label = data$logconcentration_plus_median)

params <- list(
  objective = "reg:squarederror", # Specify regression task
  eta = 0.1,                      # Learning rate
  max_depth = 4,                  # Maximum depth of a tree
  monotone_constraints = c(0, -1, rep(0,15)), # adding this actually improve the performance
  eval_metric = "rmse"            # The metric to evaluate
)

cv_results <- xgb.cv(
  params = params,
  data = dtrain,
  nrounds = 400,
  nfold = 5,
  early_stopping_rounds = 50, # Stop if performance doesn't improve
  verbose = 0                 # Suppress verbose output
)
```

```{r}
# 4. Evaluate the results
print("Cross-validation results (last 5 rounds):")
# print(tail(cv_results$evaluation_log))
print((cv_results$evaluation_log))

# Find the best number of rounds based on the lowest test RMSE
best_iteration <- cv_results$best_iteration
print(paste("Best iteration:", best_iteration))
```


```{r}
OR_mean_model_xgboost <- xgboost(data = design_matrix, 
                                 label = data$logconcentration_plus_median,
                                 eta = 0.1,                      # Learning rate
                                 max_depth = 4, 
                                 # label = data$concentration_plus_median,
                                 nrounds = cv_results$best_iteration,
                                 monotone_constraints = c(0, -1, rep(0,15)))
```

```{r}
importance_matrix <- xgb.importance(feature_names = colnames(design_matrix), model = OR_mean_model_xgboost)
xgb.plot.importance(importance_matrix = importance_matrix)
```

```{r}
plot(data$logconcentration_plus_median, predict(OR_mean_model_xgboost,design_matrix), cex = 0.2)
abline(0,1,col='red')

plot(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost,design_matrix_test),cex = 0.2)
abline(0,1,col='red')

Metrics::rmse(data$logconcentration_plus_median, predict(OR_mean_model_xgboost,design_matrix))
Metrics::rmse(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost,design_matrix_test))
```


```{r}
cor(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix_test))
```

Percentage of agricultural

Separate data into most recent measurement/the historical measurement at every location (grid). 

Take average/median historical of neighbors in township/range. 

Find the region with strong spatial correlation. Subset of the sample. 

##### Estimate Residual using Gaussian Process

```{r}
set.seed(1998)
loocv_model <- xgb.cv(
  data = dtrain,
  nrounds = cv_results$best_iteration,
  eta = 0.1,                      # Learning rate
  max_depth = 4, 
  monotone_constraints = c(0, -1, rep(0,15)),
  nfold = 10,                 # Key step for LOOCV
  prediction = TRUE,         # Important: tells the function to return predictions
  verbose = 0                # Suppress progress messages
)
```

```{r}
gpfit = GpGp::fit_model(data$logconcentration_plus_median - loocv_model$pred,
                        locs = data[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")
```

```{r}
krige_values <- leave_one_out_kriging(locs = data[,c("longitude", "latitude")], 
                                      y_obs = data$logconcentration_plus_median - loocv_model$pred, 
                                      gp_model = gpfit$covfun_name,
                                      gp_params = gpfit$covparms,
                                      order = c("coordinate")) 
```

```{r}
krige_values_test <- GpGp::predictions(fit = gpfit, 
                      locs_pred = data_test[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_test)))
```

```{r}
cat('Test rMSE: with kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix_test) + krige_values_test))
cat('\n')
cat('Test rMSE: No kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix_test)))
cat('\n')

cat('Training rMSE: with kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix) + krige_values))
cat('\n')
cat('Training rMSE: No kriging: ')
cat(Metrics::rmse(log(data$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix)))
cat('\n')
```
```{r}
plot(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost,design_matrix_test) + krige_values_test,cex = 0.2, main = 'kriging test prediction')
abline(0,1,col='red')
plot(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost,design_matrix_test),cex = 0.2, main = ' no kriging test prediction')
abline(0,1,col='red')
```


It is not easier to predict in small area. It actually depends on which areas we choose
The GP regression fails in small area. It actually depends on which areas we choose.

##### Compute OR smoothers: XGBoost 

```{r}
boundary_tol <- 0
depth_range <- c(min(data$logWellDepth) - boundary_tol, max(data$logWellDepth) + boundary_tol)
depth_range
```

```{r}
# xgboost use the design matrix
cf_predictions <- precompute_cf_predictions(data_obs = design_matrix, 
                              fit_outcome = OR_mean_model_xgboost,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
smoothers <- make_prediction_smoothers(treatment_grid = cf_predictions$treatment_values,
                                            pred_matrix_cf = cf_predictions$pred_matrix,
                                            # smooth_method = "gam",
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers <- make_cumint_smoothers(treatment_grid = cf_predictions$treatment_values,
                                         pred_matrix_cf = cf_predictions$pred_matrix,
                                         smooth_method = "splinefun",
                                         # smooth_method = "gam",
                                         # smooth_method = "smooth.spline",
                                         spar = NULL)
```

```{r}
# xgboost use the design matrix
cf_predictions_test <- precompute_cf_predictions(data_obs = design_matrix_test, 
                              fit_outcome = OR_mean_model_xgboost,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
smoothers_test <- make_prediction_smoothers(treatment_grid = cf_predictions_test$treatment_values,
                                            pred_matrix_cf = cf_predictions_test$pred_matrix,
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers_test <- make_cumint_smoothers(treatment_grid = cf_predictions_test$treatment_values,
                                               pred_matrix_cf = cf_predictions_test$pred_matrix,
                                               # smooth_method = "splinefun",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```

```{r}
i <- 10
plot(seq(0,8,length.out = 100), sapply(seq(0,8,length.out = 100), function(knot) {smoothers[[i]](knot)}))
```

##### Estimate GPS

###### Normal Density

```{r}
GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
```

```{r}
gps_mean_model <- SuperLearner(
  Y = data$logWellDepth,
  X = data[, GPS_covariate_names],
  family = gaussian(),
  SL.library = c("SL.ranger")
)
```

```{r}
gpfit_gps <- GpGp::fit_model(data$logWellDepth - gps_mean_model$Z,
                        locs = data[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")
```

```{r}
krige_values_depth <- leave_one_out_kriging(locs = data[,c("longitude", "latitude")], 
                                      y_obs = data$logWellDepth - gps_mean_model$Z, 
                                      gp_model = gpfit_gps$covfun_name,
                                      gp_params = gpfit_gps$covparms,
                                      order = c("coordinate")) 
```

```{r}
krige_values_depth_test <- GpGp::predictions(fit =  gpfit_gps, 
                      locs_pred = data_test[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_test)))
```

```{r}
Metrics::rmse(data_test$logWellDepth, predict(gps_mean_model, data_test[, GPS_covariate_names])$pred + krige_values_depth_test)
Metrics::rmse(data_test$logWellDepth, predict(gps_mean_model, data_test[, GPS_covariate_names])$pred)
```

Plot the result

```{r}
# visualize the residual of the fitted value
plot(data$logWellDepth, gps_mean_model$SL.predict)
abline(0,1,col='red')

# visualize the error of prediction
plot(data_test$logWellDepth, predict(gps_mean_model, data_test[, GPS_covariate_names])$pred)
abline(0,1,col='red')

# visualize the error of prediction
plot(data_test$logWellDepth, predict(gps_mean_model, data_test[, GPS_covariate_names])$pred + krige_values_depth_test)
abline(0,1,col='red')
```


```{r}
gps_sigma = sd(data$logWellDepth - gps_mean_model$Z)
gps_sigma
gps_sigma = sd(data$logWellDepth - gps_mean_model$Z - krige_values_depth)
gps_sigma
```


###### GPS: Covariate Balancing

```{r}
library(independenceWeights)
```

```{r}
weights <- independence_weights(A = data$logWellDepth, X = data[, GPS_covariate_names])
```

```{r}
plot(weights$weights,1/gps_est)
```

```{r}
library(WeightIt)
```

```{r}
weights_cbps <- weightit(logWellDepth ~ StaticLevel+WellUse+crop_type_combine+drainagecl+precipitation+cafolog, 
                         data = data, method = "cbps")
```

```{r}
plot(weights_cbps$weights, weights$weights)
```


#### indirect method

```{r}
# threshold_val = log(10)
threshold_val = log(5)
# threshold_val = log(3)
```

Approximately 1:3, should we do stratified sample splitting when we split train/test data

```{r}
table(data$logconcentration_plus_median<=threshold_val)
table(data_test$logconcentration_plus_median<=threshold_val)
```

##### indirect method by OR: RKHS

###### non-spatial

get non-spatial policy

```{r}
indirect_policy_nonsp_RKHS <- sapply(1:nrow(data), function(i) {
  if(smoothers_RKHS[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers_RKHS[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_RKHS[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
indirect_policy_nonsp_test_RKHS <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test_RKHS[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers_test_RKHS[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test_RKHS[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

get non-spatial policy evaluation

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_nonsp_RKHS > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(indirect_policy_nonsp_RKHS > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_nonsp_test_RKHS > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_nonsp_test_RKHS > data_test$logWellDepth))
```

###### spatial indirect

```{r}
indirect_policy_RKHS <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(smoothers_RKHS[[i]](depth_range[2]) + krige_values_RKHS[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_RKHS[[i]](t) + krige_values_RKHS[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```


```{r}
indirect_policy_test_RKHS <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test_RKHS[[i]](depth_range[1]) + krige_values_test_RKHS[i]<= threshold_val) return(depth_range[1])
  if(smoothers_test_RKHS[[i]](depth_range[2]) + krige_values_test_RKHS[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test_RKHS[[i]](t) + krige_values_test_RKHS[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_RKHS > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_RKHS > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_test_RKHS > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_test_RKHS > data_test$logWellDepth))
```

##### indirect method by OR: gradient boost

```{r}
plot(log(data$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix) + krige_values, cex = 0.2, main = "nitrate prediction: training data")
abline(0,1,col='lightblue')
abline(h = threshold_val,col='lightblue')
abline(v = threshold_val,col='lightblue')

plot(log(data_test$concentration_plus_median), predict(OR_mean_model_xgboost, design_matrix_test) + krige_values_test, cex = 0.2, main = "nitrate prediction: test data")
abline(0,1,col='lightblue')
abline(h = threshold_val,col='lightblue')
abline(v = threshold_val,col='lightblue')
```

###### non-spatial

get non-spatial policy

```{r}
indirect_policy_nonsp <- sapply(1:nrow(data), function(i) {
  if(smoothers[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
indirect_policy_nonsp_test <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers_test[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

get non-spatial policy evaluation

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_nonsp > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(indirect_policy_nonsp > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_nonsp_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_nonsp_test > data_test$logWellDepth))
```

###### spatial indirect

```{r}
indirect_policy <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(smoothers[[i]](depth_range[1]) + krige_values[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(smoothers[[i]](depth_range[2]) + krige_values[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers[[i]](t) + krige_values[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```


```{r}
indirect_policy_test <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test[[i]](depth_range[1]) + krige_values_test[i]<= threshold_val) return(depth_range[1])
  if(smoothers_test[[i]](depth_range[2]) + krige_values_test[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test[[i]](t) + krige_values_test[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_test > data_test$logWellDepth))
```

##### indirect method by IPW


We first set the bandwidth

```{r}
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
```

```{r}
gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)
```


We try different weights, but in general it does not work

###### Independence energy weights

```{r}
indirect_policy_ipw <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma = 5* kernel_bw) * weights$weights[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[2], data$logWellDepth[i], sigma = 5* kernel_bw) * weights$weights[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {data$logconcentration_plus_median[i] * gaussian_kernel(t, data$logWellDepth[i], sigma = 5*  kernel_bw) * weights$weights[i] - threshold_val},
         lower = depth_range[1],
         upper = depth_range[2])
  return(res$root)
})
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_ipw > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_ipw > data$logWellDepth))
```

almost all of them goes to the lower boundary

```{r}
mean(indirect_policy_ipw == depth_range[1])
```

###### gaussian propensity weights

```{r}
indirect_policy_ipw <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma = 5* kernel_bw)/gps_est[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[2], data$logWellDepth[i], sigma = 5* kernel_bw)/gps_est[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {data$logconcentration_plus_median[i] * gaussian_kernel(t, data$logWellDepth[i], sigma = 5*  kernel_bw)/gps_est[i] - threshold_val},
         lower = depth_range[1],
         upper = depth_range[2])
  return(res$root)
})
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_ipw > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_ipw > data$logWellDepth))
```

almost all of them goes to the lower boundary

```{r}
mean(indirect_policy_ipw == depth_range[1])
```

###### covariate balancing weights

```{r}
indirect_policy_ipw <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma = 5* kernel_bw) * weights_cbps$weights[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[2], data$logWellDepth[i], sigma = 5* kernel_bw) * weights_cbps$weights[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {data$logconcentration_plus_median[i] * gaussian_kernel(t, data$logWellDepth[i], sigma = 5*  kernel_bw)* weights_cbps$weights[i] - threshold_val},
         lower = depth_range[1],
         upper = depth_range[2])
  return(res$root)
})
```

```{r}
i <- 1
data$logconcentration_plus_median[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma =  10*kernel_bw)/gps_est[i]
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_ipw > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_ipw > data$logWellDepth))
```

almost all of them goes to the lower boundary

```{r}
mean(indirect_policy_ipw == depth_range[1])
```

The IPW initialization does not work.

##### indirect method by DB

###### indreict DB RKHS

```{r}
resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
scaled_resids <- resids * weights_cbps$weights
plot(scaled_resids)
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
```

```{r}
indirect_policy_DB <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma = kernel_bw) <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(smoothers_RKHS[[i]](depth_range[2]) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(depth_range[2], data$logWellDepth[i], sigma = kernel_bw)> threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_RKHS[[i]](t) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(t, data$logWellDepth[i], sigma = kernel_bw)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

When I isotonic the RKHS, the improvement from DB indirect than the indirect suddenly becomes larger in training data.

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_DB> data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_DB > data$logWellDepth))
```

compared to the 0.85/0.56 (acc, mcc) in training data of indirect method, this is much better, however, technically, we are not able to get test indirect method in test data because we don't know that obs in test data

```{r}
hist(indirect_policy_RKHS)
mean(indirect_policy_RKHS == depth_range[1])
```


```{r}
hist(indirect_policy_DB)
mean(indirect_policy_DB == depth_range[1])
```


###### indirect DB XGBoost

#### Direct Method: RKHS SVM

```{r}
design_matrix[, "logWellDepth"] <- (design_matrix[, "logWellDepth"] - mean_logWellDepth)/sd_logWellDepth
design_matrix[, "precipitation"] <- (design_matrix[, "precipitation"] - mean_precipitation)/sd_precipitation
design_matrix[, "cafolog"] <- (design_matrix[, "cafolog"] - mean_cafolog)/sd_cafolog
design_matrix[, "StaticLevel"] <- (design_matrix[, "StaticLevel"] - mean_StaticLevel)/sd_StaticLevel
# standardize for test data
design_matrix_test[, "logWellDepth"] <- (design_matrix_test[, "logWellDepth"] - mean_logWellDepth)/sd_logWellDepth
design_matrix_test[, "precipitation"] <- (design_matrix_test[, "precipitation"] - mean_precipitation)/sd_precipitation
design_matrix_test[, "cafolog"] <- (design_matrix_test[, "cafolog"] - mean_cafolog)/sd_cafolog
design_matrix_test[, "StaticLevel"] <- (design_matrix_test[, "StaticLevel"] - mean_StaticLevel)/sd_StaticLevel
```

#### Direct Method: RKHS SVM coordinates

###### build grand matrix 

```{r}
coords_matrix <- model.matrix(~ . , data = data[, c("longitude", "latitude")])[,-1]
coords_matrix_test <- model.matrix(~ . , data = data_test[, c("longitude", "latitude")])[,-1]

train_means <- colMeans(coords_matrix)
train_sds <- apply(coords_matrix, 2, sd)

coords_matrix <- scale(coords_matrix)
coords_matrix_test <- scale(coords_matrix_test, center = train_means, scale = train_sds)
```

```{r}
m <- 0
gamma <- 2^(m) * median(rdist(coords_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))

k_coords_matrix <- kernelMatrix(rbf, coords_matrix)
# build kernel matrix for test data
k_coords_matrix_test <- kernelMatrix(rbf, coords_matrix_test, coords_matrix)
```

```{r}
hist(-log(k_coords_matrix[col(k_coords_matrix)!=row(k_coords_matrix)]))
abline(v = 1,col='blue')

hist(k_coords_matrix[col(k_coords_matrix)!=row(k_coords_matrix)],xlim = c(0,1))
abline(v = exp(-1),col='blue')
```

When using coordinate as covariance, it is either many zeros or many ones in the grand matrix

```{r}
# check rank
qr(k_coords_matrix)$rank # this is very slow, takes a very long time to run
# when m = 0.3, the rank is 276
dim(k_coords_matrix)
```

###### add intercept

```{r}
k_coords_matrix <- cbind(rep(1,nrow(k_coords_matrix)), k_coords_matrix)
k_coords_matrix_test <- cbind(rep(1,nrow(k_coords_matrix_test)), k_coords_matrix_test)
```

###### initialize: stretch indirect_policy

Let's see how many of indirect_policy are in the boundary

```{r}
depth_range
mean(indirect_policy == depth_range[1])
mean(indirect_policy == depth_range[2])
mean(indirect_policy_test == depth_range[1])
mean(indirect_policy_test == depth_range[2])
```

Let' see the histogram of indirect_policy

```{r}
hist(indirect_policy)
hist(data$logconcentration_plus_median)
abline(v = threshold_val,col='blue')
hist(indirect_policy_test)
hist(data_test$logconcentration_plus_median)
abline(v = threshold_val,col='blue')
```

As mentioned in the Chan's MRTP, since the indirect method is bounded in the depth range, it may not be suitable approximation of the SVM solution before the winsorization. Using indirect_policy may lead to an estimate policy shrinking to a certain alue, i.e., policy does not reflect the heterogeneity induced by covariate X.
Thus, we need to stretch the in_direct policy outside the boundary

```{r}
indices_non_depth_boundary <- (indirect_policy!= depth_range[1]) & (indirect_policy!= depth_range[2])

depth_knots <- seq(depth_range[1], depth_range[2], length.out = 10)

phi_lst <- vector("list", length(depth_knots) + 1)
stretch_models <- vector("list", length(depth_knots) + 1)

for(i in 1:length(depth_knots)){
  depth <- depth_knots[i]
  # fine the fitted value at that depth
  phi_lst[[i]] <- sapply(1:nrow(data), function(indice){smoothers[[indice]](depth)})
  # only use non-boundary value to fit linear model
  stretch_models[[i]] <- lm(phi_lst[[i]][indices_non_depth_boundary] ~ indirect_policy[indices_non_depth_boundary])
}
gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

phi_lst[[length(depth_knots) + 1]] <- (data$logconcentration_plus_median - sapply(1:nrow(data), function(indice){smoothers[[indice]](data$logWellDepth[indice])}))/gps_est

stretch_models[[length(depth_knots) + 1]] <- lm(phi_lst[[length(depth_knots) + 1]][indices_non_depth_boundary] ~ indirect_policy[indices_non_depth_boundary])
```

Then, we calculate the stretch value of the policy at the boundary

```{r}
indirect_policy_stretch_lst <- vector("list", length(depth_knots) + 1)

for(i in 1:(length(depth_knots) + 1)){
  intercept <- coef(stretch_models[[i]])[1]
  slope <- coef(stretch_models[[i]])[2]
  indirect_policy_stretch_lst[[i]] <- (phi_lst[[i]] - intercept)/slope
}

indirect_policy_stretch_lst_min <- do.call(pmin, indirect_policy_stretch_lst)
indirect_policy_stretch_lst_max <- do.call(pmax, indirect_policy_stretch_lst)
```

```{r}
indices_depth_lower_boundary <- indirect_policy == depth_range[1]
indices_depth_upper_boundary <- indirect_policy == depth_range[2]
```

```{r}
indirect_policy_stretch <- indirect_policy
indirect_policy_stretch[indices_depth_lower_boundary] <- indirect_policy_stretch_lst_min[indices_depth_lower_boundary]
indirect_policy_stretch[indices_depth_upper_boundary] <- indirect_policy_stretch_lst_max[indices_depth_upper_boundary]
```

```{r}
plot(indirect_policy, indirect_policy_stretch)
plot(1:nrow(data), indirect_policy_stretch)
summary(indirect_policy_stretch)
indirect_policy_stretch <- pmax(-10, indirect_policy_stretch)
hist(indirect_policy_stretch)
```

Let's check the accuracy and mcc for the stretched policy. We found that the stretched policy even improve the performance a little bit.

Also, reverse true/false does not affect the acc/mcc that much, if there is no indirect_policy == data$logWellDepth, I think the result should be the same.

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) < threshold_val, indirect_policy < data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) < threshold_val), estimate = as.factor( indirect_policy < data$logWellDepth))
```


```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_stretch > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_stretch > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) < threshold_val, indirect_policy_stretch < data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) < threshold_val), estimate = as.factor( indirect_policy_stretch < data$logWellDepth))
```


Then, we find the initial value of the DC algorithm

Tune the bandwidth is important, exp(-1), median bandwidth might not work

Try to get the param_initial

$$\min\|alpha\|_2^2 s.t. K\alpha = \theta +\epsilon$$. \epsilon is a relexation. K omits the intercept. only use it to find the non-intercept coefficient. Using optiSolve.

Two way: (1) tune hyper-parameter using cross-validation (2) do not tune hyper-parameter, find a good initial value

What we want is simplest alpha that satisfy the initial. 

-log() of off -diagnol grand matrix entries should be central around 1. Sanity check of RKHS. off-diagnal too small, over-fitted.
The problem is in the initial point. 


###### initialze: generalized inverse

Use indirect policy

```{r}
params_coords_initial_ginv <- rep(0, ncol(k_coords_matrix))
params_coords_initial_ginv[1] <- mean(indirect_policy)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy - mean(indirect_policy))
```

```{r}
plot(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv))
summary(indirect_policy - as.vector(k_coords_matrix %*%params_coords_initial_ginv))
summary(indirect_policy)
hist(params_coords_initial_ginv, main = "histogram of ginv initialize")
```

Let's check the acc and mcc of the initialized parameter. Why they decreased so much.

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_coords_matrix %*%params_coords_initial_ginv) > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.vector(k_coords_matrix %*%params_coords_initial_ginv) > data$logWellDepth))
```

Use indirect policy stretch

```{r}
params_coords_initial_stretch_ginv <- rep(0, ncol(k_coords_matrix))
params_coords_initial_stretch_ginv[1] <- mean(indirect_policy_stretch)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_stretch_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy_stretch - mean(indirect_policy_stretch))
```

```{r}
plot(indirect_policy_stretch, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy_stretch, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv))
summary(indirect_policy_stretch - as.vector(k_coords_matrix %*%params_coords_initial_stretch_ginv))
summary(indirect_policy_stretch)
hist(params_coords_initial_stretch_ginv, main = "histogram of ginv initialize")
## check accurary and mcc for the initial point
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv) > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv > data$logWellDepth)))
```


###### initialze: cv glmnet

```{r}

cv.glmnet(x = k_coords_matrix[,-1], y = indirect_policy)
params_coords_initial_cvglmet <- 
params_coords_initial_ginv[1] <- mean(indirect_policy)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy - mean(indirect_policy))
```

```{r}
plot(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv))
summary(indirect_policy - as.vector(k_coords_matrix %*%params_coords_initial_ginv))
summary(indirect_policy)
hist(params_coords_initial_ginv, main = "histogram of ginv initialize")
```

###### initialze: glmnet 

```{r}
params_coords_initial_ginv <- rep(0, ncol(k_coords_matrix))
params_coords_initial_ginv[1] <- mean(indirect_policy)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy - mean(indirect_policy))
```

```{r}
plot(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv))
summary(indirect_policy - as.vector(k_coords_matrix %*%params_coords_initial_ginv))
summary(indirect_policy)
hist(params_coords_initial_ginv, main = "histogram of ginv initialize")
```

##### DC algorithm

```{r}
max_iter <- 50
tol <- 1e-4
clip_epsilon_bar <- 300
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

loss_type <- "ipw"
resids <- data$logconcentration_plus_median
krige_adjust <- rep(0, nrow(data))

# loss_type <- 'or'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict
# krige_adjust <- krige_values

# loss_type <- 'db'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict
# krige_adjust <- krige_values


gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

params_old <- params_coords_initial
params <- params_coords_initial

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = params_old, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = params_old, 
                        K = k_coords_matrix, 
                        T = data$logWellDepth,
                        krige_adjust = krige_adjust,
                        outcome_resid = resids,
                        propensity_est = pmax(gps_est,0.001),
                        lambda = 0, 
                        smoothers = smoothers, 
                        cumint_smoothers = cumint_smoothers, 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = threshold_val, 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```

```{r}
plot(params_coords_initial, params_old)
```


```{r}
print('training')
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_coords_matrix %*% params_old)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_coords_matrix %*% params_old)  > data$logWellDepth))

print('test')
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, as.numeric(k_coords_matrix_test%*% params_old  > data_test$logWellDepth))
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_coords_matrix_test %*% params_old  > data_test$logWellDepth)))
```

Let's try non-DC, direct optimization algorithm

compute_total_loss_smooth <- function(assigned_trt,
                                         subject_idx,
                                         T,
                                         krige_adjust,
                                         outcome_resid,
                                         propensity_est,
                                         cumint_smoothers,
                                         trt_bounds,
                                         threshold_val,
                                         kernel_bw,
                                         clip_epsilon,
                                         surrogate_type = c("Gaussian", "Box"),
                                         loss_type = c("db","ipw","or"))

```{r}
optim(par = params_old, 
      fn = compute_total_loss_smooth, 
      gr = d_compute_total_loss_db_smooth, # we try the case with/without gr
      params_old = params_old, 
      K = k_coords_matrix, 
      T = data$logWellDepth,
      krige_adjust = krige_adjust,
      outcome_resid = resids,
      propensity_est = pmax(gps_est,0.001),
      lambda = 0, 
      smoothers = smoothers, 
      cumint_smoothers = cumint_smoothers, 
      kernel_bw = kernel_bw, 
      trt_bounds = depth_range,
      threshold_val = threshold_val, 
      clip_epsilon = clip_epsilon,
      clip_epsilon_bar = clip_epsilon_bar,
      surrogate_type = surrogate_type,
      loss_type = loss_type,
      method = "L-BFGS-B", 
      control = list(maxit = 60, trace = 3))
```


#### Direct Method: RKHS SVM covariate

##### preparation and initialize: for covariate

###### build grand matrix 

```{r}
# kernel_design_matrix <- model.matrix(~ . , data = data[, GPS_covariate_names])[,-1]
# kernel_design_matrix_test <- model.matrix(~ . , data = data_test[, GPS_covariate_names])[,-1]
# 
# kernel_design_matrix <- model.matrix(~ . , data = cbind(data[, GPS_covariate_names], data[, c("longitude", "latitude")]))[,-1]
# kernel_design_matrix_test <- model.matrix(~ . , data = cbind(data_test[, GPS_covariate_names], data_test[, c("longitude", "latitude")]))[,-1]

# we use the following one for now
kernel_design_matrix <- model.matrix(~ . , data = cbind(data[, GPS_covariate_names], U = krige_values_RKHS))[,-1]
kernel_design_matrix_test <- model.matrix(~ . , data = cbind(data_test[, GPS_covariate_names], U = krige_values_test_RKHS))[,-1]
```

```{r}
train_means <- colMeans(kernel_design_matrix)
train_sds <- apply(kernel_design_matrix, 2, sd)

kernel_design_matrix <- scale(kernel_design_matrix)
kernel_design_matrix_test <- scale(kernel_design_matrix_test, center = train_means, scale = train_sds)
```

```{r}
# we need to scale the continuous variables
# mean_logWellDepth <- mean(design_matrix[, "logWellDepth"])
# sd_logWellDepth <- sd(design_matrix[, "logWellDepth"])

mean_precipitation <- mean(kernel_design_matrix[, "precipitation"])
sd_precipitation <- sd(kernel_design_matrix[, "precipitation"])

mean_cafolog <- mean(kernel_design_matrix[, "cafolog"])
sd_cafolog <- sd(kernel_design_matrix[, "cafolog"])

mean_StaticLevel <- mean(kernel_design_matrix[, "StaticLevel"])
sd_StaticLevel <- sd(kernel_design_matrix[, "StaticLevel"])
```

```{r}
kernel_design_matrix[, "precipitation"] <- (kernel_design_matrix[, "precipitation"] - mean_precipitation)/sd_precipitation
kernel_design_matrix[, "cafolog"] <- (kernel_design_matrix[, "cafolog"] - mean_cafolog)/sd_cafolog
kernel_design_matrix[, "StaticLevel"] <- (kernel_design_matrix[, "StaticLevel"] - mean_StaticLevel)/sd_StaticLevel
# standardize for test data
kernel_design_matrix_test[, "precipitation"] <- (kernel_design_matrix_test[, "precipitation"] - mean_precipitation)/sd_precipitation
kernel_design_matrix_test[, "cafolog"] <- (kernel_design_matrix_test[, "cafolog"] - mean_cafolog)/sd_cafolog
kernel_design_matrix_test[, "StaticLevel"] <- (kernel_design_matrix_test[, "StaticLevel"] - mean_StaticLevel)/sd_StaticLevel
```

```{r}
m <- 0
gamma <- 2^(m) * median(rdist(kernel_design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))

k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
# build kernel matrix for test data
k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
```

```{r}
hist(-log(k_matrix[col(k_matrix)!=row(k_matrix)]))
abline(v = 1,col='blue')

hist(k_matrix[col(k_matrix)!=row(k_matrix)],xlim = c(0,1))
abline(v = exp(-1),col='blue')
```

This is like the least square loss

```{r}
svm_policy <- svm(x = kernel_design_matrix, y = indirect_policy_DB)
```

training metrics

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, predict(svm_policy, kernel_design_matrix)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(predict(svm_policy, kernel_design_matrix)  > data$logWellDepth))
```

test metrics

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, predict(svm_policy, kernel_design_matrix_test)  > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor(predict(svm_policy, kernel_design_matrix_test)  > data_test$logWellDepth))
```

It seems not too bad?

###### add intercept

```{r}
k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
```

###### initialize: stretch indirect_policy

As mentioned in the Chan's MRTP, since the indirect method is bounded in the depth range, it may not be suitable approximation of the SVM solution before the winsorization. Using indirect_policy may lead to an estimate policy shrinking to a certain alue, i.e., policy does not reflect the heterogeneity induced by covariate X.
Thus, we need to stretch the in_direct policy outside the boundary

```{r}
indices_non_depth_boundary <- (indirect_policy_DB!= depth_range[1]) & (indirect_policy_DB!= depth_range[2])

depth_knots <- seq(depth_range[1], depth_range[2], length.out = 10)

phi_lst <- vector("list", length(depth_knots) + 1)
stretch_models <- vector("list", length(depth_knots) + 1)

for(i in 1:length(depth_knots)){
  depth <- depth_knots[i]
  # fine the fitted value at that depth
  phi_lst[[i]] <- sapply(1:nrow(data), function(indice){smoothers_RKHS[[indice]](depth) + krige_values_RKHS[indice]})
  # only use non-boundary value to fit linear model
  stretch_models[[i]] <- lm(phi_lst[[i]][indices_non_depth_boundary] ~ indirect_policy_DB[indices_non_depth_boundary])
}

gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

phi_lst[[length(depth_knots) + 1]] <- (data$logconcentration_plus_median - sapply(1:nrow(data), function(indice){smoothers_RKHS[[indice]](data$logWellDepth[indice]) - krige_values_RKHS[indice]}))/gps_est

stretch_models[[length(depth_knots) + 1]] <- lm(phi_lst[[length(depth_knots) + 1]][indices_non_depth_boundary] ~ indirect_policy_DB[indices_non_depth_boundary])
```

Then, we calculate the stretch value of the policy at the boundary

```{r}
indirect_policy_DB_stretch_lst <- vector("list", length(depth_knots) + 1)

for(i in 1:(length(depth_knots) + 1)){
  intercept <- coef(stretch_models[[i]])[1]
  slope <- coef(stretch_models[[i]])[2]
  indirect_policy_DB_stretch_lst[[i]] <- (phi_lst[[i]] - intercept)/slope
}

indirect_policy_DB_stretch_lst_min <- do.call(pmin, indirect_policy_DB_stretch_lst)
indirect_policy_DB_stretch_lst_max <- do.call(pmax, indirect_policy_DB_stretch_lst)
```

```{r}
indices_depth_lower_boundary <- indirect_policy_DB == depth_range[1]
indices_depth_upper_boundary <- indirect_policy_DB == depth_range[2]
```

```{r}
indirect_policy_DB_stretch <- indirect_policy_DB
indirect_policy_DB_stretch[indices_depth_lower_boundary] <- indirect_policy_DB_stretch_lst_min[indices_depth_lower_boundary]
indirect_policy_DB_stretch[indices_depth_upper_boundary] <- indirect_policy_DB_stretch_lst_max[indices_depth_upper_boundary]
```

```{r}
plot(indirect_policy_DB, indirect_policy_DB_stretch)
plot(1:nrow(data), indirect_policy_DB_stretch)
summary(indirect_policy_DB_stretch)

indirect_policy_DB_stretch <- pmax(-20, indirect_policy_DB_stretch)
hist(indirect_policy_DB_stretch)
```

Let's check the accuracy and mcc for the stretched policy. We found that the stretched policy even improve the performance a little bit.

Also, reverse true/false does not affect the acc/mcc that much, if there is no indirect_policy == data$logWellDepth, I think the result should be the same.

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_DB > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_DB > data$logWellDepth))
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) < threshold_val, indirect_policy_DB_stretch < data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) < threshold_val), estimate = as.factor( indirect_policy_DB_stretch < data$logWellDepth))
```



###### initialze: generalized inverse

Use indirect policy

```{r}
params_coords_initial_ginv <- rep(0, ncol(k_matrix))
params_coords_initial_ginv[1] <- mean(indirect_policy)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy - mean(indirect_policy))
```

```{r}
plot(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy, as.vector(k_coords_matrix %*% params_coords_initial_ginv))
summary(indirect_policy - as.vector(k_coords_matrix %*%params_coords_initial_ginv))
summary(indirect_policy)
hist(params_coords_initial_ginv, main = "histogram of ginv initialize")
```

Let's check the acc and mcc of the initialized parameter. Why they decreased so much.

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_coords_matrix %*%params_coords_initial_ginv) > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.vector(k_coords_matrix %*%params_coords_initial_ginv) > data$logWellDepth))
```

Use indirect policy stretch

```{r}
params_coords_initial_stretch_ginv <- rep(0, ncol(k_coords_matrix))
params_coords_initial_stretch_ginv[1] <- mean(indirect_policy_stretch)
# generalized inverse rather than inverse, this is extremely slow 
params_coords_initial_stretch_ginv[2:ncol(k_coords_matrix)] <- MASS::ginv(k_coords_matrix[,-1]) %*% (indirect_policy_stretch - mean(indirect_policy_stretch))
```

```{r}
plot(indirect_policy_stretch, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv), main = "initialize: ginv v.s. indirect")
Metrics::rmse(indirect_policy_stretch, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv))
summary(indirect_policy_stretch - as.vector(k_coords_matrix %*%params_coords_initial_stretch_ginv))
summary(indirect_policy_stretch)
hist(params_coords_initial_stretch_ginv, main = "histogram of ginv initialize")
## check accurary and mcc for the initial point
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv) > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.vector(k_coords_matrix %*% params_coords_initial_stretch_ginv > data$logWellDepth)))
```


###### initialze: cv glmnet

```{r}
initial_cvglmnet <- cv.glmnet(x = k_matrix[,-1], y = indirect_policy_DB, alpha = 0)
params_initial_cvglmet <- coef(initial_cvglmnet)
params_initial_cvglmet <- as.numeric(params_initial_cvglmet)
```

```{r}
plot(indirect_policy_DB, as.vector(k_matrix %*% params_initial_cvglmet), main = "initialize: indirect v.s. indirect DB")
Metrics::rmse(indirect_policy_DB, as.vector(k_matrix %*% params_initial_cvglmet))
summary(indirect_policy_DB - as.vector(k_matrix %*% params_initial_cvglmet))
summary(indirect_policy_DB)
hist(params_initial_cvglmet, main = "histogram of cvglmnet initialize")
```

###### initialze: glmnet 

```{r}
# the original indirect policy
initial_glmnet <- glmnet(x = k_matrix[,-1], y = indirect_policy_DB, alpha = 0, lambda = 0)
params_initial_glmnet <- coef(initial_glmnet)
params_initial_glmnet <- as.numeric(params_initial_glmnet)

# # or the stretched indirect policy
# initial_glmnet <- glmnet(x = k_matrix[,-1], y = indirect_policy_DB_stretch, alpha = 0, lambda = 0)
# params_initial_glmnet <- coef(initial_glmnet)
# params_initial_glmnet <- as.numeric(params_initial_glmnet)
# 
# # it seems like the stretch actually does not help
```

```{r}
plot(indirect_policy_DB, as.vector(k_matrix %*% params_initial_glmnet), main = "initialize: indirect v.s. indirect DB")
Metrics::rmse(indirect_policy_DB, as.vector(k_matrix %*% params_initial_glmnet))
summary(indirect_policy_DB - as.vector(k_matrix %*% params_initial_glmnet))
summary(indirect_policy_DB)
hist(params_initial_glmnet, main = "histogram of cvglmnet initialize")
```


result of cv.glmnet initial

```{r}
print('training')
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_matrix %*% params_initial_glmnet)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix %*% params_initial_glmnet)  > data$logWellDepth))

print('test')
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, as.numeric(k_matrix_test%*% params_initial_glmnet  > data_test$logWellDepth))
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix_test %*% params_initial_glmnet  > data_test$logWellDepth)))
```

```{r}
print('training')
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_matrix %*% params_initial_cvglmet)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix %*% params_initial_cvglmet)  > data$logWellDepth))

print('test')
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, as.numeric(k_matrix_test%*% params_initial_cvglmet  > data_test$logWellDepth))
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix_test %*% params_initial_cvglmet  > data_test$logWellDepth)))
```

##### dc algorithm

```{r}
max_iter <- 50
tol <- 1e-4
clip_epsilon_bar <- 300
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

# loss_type <- "ipw"
# resids <- data$logconcentration_plus_median
# krige_adjust <- rep(0, nrow(data))

# loss_type <- 'or'
# resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
# krige_adjust <- krige_values_RKHS

loss_type <- 'db'
resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
krige_adjust <- krige_values_RKHS


gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

params_old <- params_initial_glmnet
params <- params_initial_glmnet

smoothers <- smoothers_RKHS
cumint_smoothers <- cumint_smoothers_RKHS
threshold_val <- log(5)

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = params_old, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = params_old, 
                        K = k_matrix, 
                        T = data$logWellDepth,
                        krige_adjust = krige_adjust,
                        outcome_resid = resids,
                        propensity_est = pmax(gps_est,0.001),
                        lambda = 10, 
                        smoothers = smoothers, 
                        cumint_smoothers = cumint_smoothers, 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = log(5), 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```


I think the IPW case just does not work

```{r}
# Some not when implement thr IPW case
# In "IPW" case: 
# When setting lambda = 0, At iterate     0  f=      -1.9501  |proj g|=       1.4006 when gr is defined, 
# At X0, 0 variables are exactly at the bounds At iterate     0  f=      -1.9501  |proj g|=       1.7171 when gr is not defined
# 
# 
# 
# It returns: At iterate     0  f=      -38.499  |proj g|=       2.3897, without specifying gr.
# But it returns At iterate     0  f=      -38.499  |proj g|=       1.3454, so there might be issue with the derivative
```

```{r}
# I think the initial looks too crazy
plot(params_initial_glmnet, params)
```



```{r}
cat('RKHS loss:')
sum(compute_total_loss_smooth(assigned_trt = k_matrix %*% params_old,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
cat('parametric loss:')
sum(compute_total_loss_smooth(assigned_trt = indirect_policy_DB,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
```


```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(k_matrix %*% params_old)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix %*% params_old > data$logWellDepth)))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, as.numeric(k_matrix_test%*% params_old  > data_test$logWellDepth))
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(k_matrix_test %*% params_old  > data_test$logWellDepth)))
```


```{r}
plot(indirect_policy_nonsp_test,  k_matrix_test %*% params_old)
plot(indirect_policy_nonsp,  k_matrix %*% params_old)
```


##### Let's use a 20 x 20 kernel to do a small test

```{r}
# we use the following one for now
kernel_design_matrix <- model.matrix(~ . , data = cbind(data[, GPS_covariate_names], U = krige_values_RKHS))[,-1]
kernel_design_matrix <- scale(kernel_design_matrix)
```

get different types of indirect method: OR based

```{r}
indirect_policy_RKHS <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(smoothers_RKHS[[i]](depth_range[2]) + krige_values_RKHS[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_RKHS[[i]](t) + krige_values_RKHS[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

get different types of indirect method: DB based

```{r}
resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
scaled_resids <- resids * weights_cbps$weights
# plot(scaled_resids)
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
```

```{r}
indirect_policy_DB <- sapply(1:nrow(data), function(i) {
  # if lower bound can alreay be lower than the threshold (satisfy the ES), we have to use lower bound
  if(smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(depth_range[1], data$logWellDepth[i], sigma = kernel_bw) <= threshold_val) return(depth_range[1])
  # if even upper bound can not be lower than the threshold, we have to use upper bound
  if(smoothers_RKHS[[i]](depth_range[2]) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(depth_range[2], data$logWellDepth[i], sigma = kernel_bw)> threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_RKHS[[i]](t) + krige_values_RKHS[i] + scaled_resids[i] * gaussian_kernel(t, data$logWellDepth[i], sigma = kernel_bw)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

check that the DB and OR indirect minimize the corresponding loss function

get the subset (sample size 20) for test

```{r}
set.seed(1998)
sub_indices <- sample(nrow(kernel_design_matrix), 20)
kernel_design_matrix_small <- kernel_design_matrix[sub_indices,]
indirect_policy_DB_small <- indirect_policy_DB[sub_indices]
indirect_policy_RKHS_small <- indirect_policy_RKHS[sub_indices]
```

```{r}
m <- 0
gamma <- 2^(m) * median(rdist(kernel_design_matrix_small))
rbf <- rbfdot(sigma = 1/(gamma^2))
k_matrix <- kernelMatrix(rbf, kernel_design_matrix_small)
k_matrix <- cbind(rep(1,20), k_matrix)
```

```{r}
hist(k_matrix)
```


```{r}
initial_glmnet <- glmnet(x = k_matrix[,-1], y = indirect_policy_DB_small, alpha = 0, lambda = 0)
params_initial_glmnet <- coef(initial_glmnet)
params_initial_glmnet <- as.numeric(params_initial_glmnet)
```

```{r}
plot(indirect_policy_DB_small, as.vector(k_matrix %*% params_initial_glmnet), main = "initialize: indirect v.s. indirect DB")
Metrics::rmse(indirect_policy_DB_small, as.vector(k_matrix %*% params_initial_glmnet))
summary(indirect_policy_DB_small - as.vector(k_matrix %*% params_initial_glmnet))
summary(indirect_policy_DB_small)
hist(params_initial_glmnet, main = "histogram of cvglmnet initialize")
```

```{r}

direct_optim_trt <- optim(par = cbind(rep(1,20),diag(20)) %*% params_initial_glmnet,
      fn = compute_total_loss_sum_smooth,
      subject_idx = 1:20,
      T = data$logWellDepth[sub_indices],
      krige_adjust = krige_values_RKHS[sub_indices],
      outcome_resid = rep(0,20),
      propensity_est = rep(1,20),
      cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
      trt_bounds = depth_range,
      threshold_val = log(5), 
      kernel_bw = kernel_bw, 
      clip_epsilon = clip_epsilon,
      surrogate_type = "Gaussian",
      loss_type = "or",
      method = "L-BFGS-B", 
      control = list(maxit = 60, trace =4))

```
```{r}
plot(indirect_policy_RKHS_small - direct_optim_trt$par)
```

```{r}
compute_total_loss_sum_smooth(assigned_trt = direct_optim_trt$par,
                              subject_idx = 1:20,
                              T = data$logWellDepth[sub_indices],
                              krige_adjust = krige_values_RKHS[sub_indices],
                              outcome_resid = rep(0,20),
                              propensity_est = rep(1,20),
                              cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                              trt_bounds = depth_range,
                              threshold_val = log(5), 
                              kernel_bw = kernel_bw, 
                              clip_epsilon = clip_epsilon,
                              surrogate_type = "Gaussian",
                              loss_type = "or")
```

```{r}
# for (indice in sub_indices){
#   plot(seq(depth_range[1],depth_range[2],length.out = 200), sapply(seq(depth_range[1],depth_range[2],length.out = 200),function(t){smoothers[[indice]](t)}))
# }
for (i in 1:length(sub_indices)){
  indice <- sub_indices[i]
  plot(seq(depth_range[1],depth_range[2],length.out = 200), sapply(seq(depth_range[1],depth_range[2],length.out = 200),function(t){threshold_val*(t-depth_range[1])- cumint_smoothers[[indice]](t)}), xlim = c(0,10), ylim = c(-3,3))
  lines(seq(depth_range[1],depth_range[2],length.out = 200), sapply(seq(depth_range[1],depth_range[2],length.out = 200),function(t){compute_total_loss_sum_smooth(assigned_trt = t,
                              subject_idx = i,
                              T = data$logWellDepth[sub_indices],
                              krige_adjust = krige_values_RKHS[sub_indices],
                              outcome_resid = rep(0,20),
                              propensity_est = rep(1,20),
                              cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                              trt_bounds = depth_range,
                              threshold_val = log(5), 
                              kernel_bw = kernel_bw, 
                              clip_epsilon = clip_epsilon,
                              surrogate_type = "Gaussian",
                              loss_type = "or")}), xlim = c(0,10),ylim = c(-3,3),col = 'grey')
  abline(v = direct_optim_trt$par[i],col='blue')
  abline(v = indirect_policy_RKHS_small[i],col='green')
}

```


! Try one sample loss function optimization.

```{r}
surrogate_type <- "Gaussian"

# loss_type <- "ipw"
# resids <- data$logconcentration_plus_median
# krige_adjust <- rep(0, nrow(data))

loss_type <- 'or'
# resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
# krige_adjust <- krige_values_RKHS

# loss_type <- 'db'
# resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
# krige_adjust <- krige_values_RKHS

# smoothers <- smoothers_RKHS
# cumint_smoothers <- cumint_smoothers_RKHS

direct_optim <- optim(#par = params_initial_glmnet, 
                      par =  c(mean(indirect_policy_DB_small), indirect_policy_DB_small - mean(indirect_policy_DB_small)),
                      fn = compute_total_loss_smooth_RKHS, # when doing this, I think it fails to calculate the derivative numerically by optim
                      gr = d_compute_total_loss_smooth_RKHS,
                      K = cbind(rep(1,20),diag(20)), 
                      T = data$logWellDepth[sub_indices],
                      krige_adjust = krige_values_RKHS[sub_indices],
                      outcome_resid = rep(0,20),
                      propensity_est = rep(1,20),
                      lambda = 0, 
                      smoothers = smoothers_RKHS[sub_indices], 
                      cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                      kernel_bw = kernel_bw, 
                      trt_bounds = depth_range,
                      threshold_val = log(5), 
                      clip_epsilon = clip_epsilon,
                      surrogate_type = surrogate_type,
                      loss_type = loss_type,
                      method = "L-BFGS-B",  # Try one sample loss function optimization. 
                      # method = "Nelder-Mead", # try non-gradient
                      control = list(maxit = 60, trace = 5))
```

```{r}
cauchy_x <- c(1.42474, 1.89435, -0.600596, -1.03981, -1.15887, -0.599502, 
              -0.958252, -0.343123, -1.1015, -0.623741, -0.658636, 
              -1.10584, 1.97502, -0.861156, 0.0160609, -0.529284, 
              0.519484, 1.9086, -0.870035, 0.046277, 1.13534)
```

```{r}
compute_total_loss_smooth_RKHS(params = cauchy_x,
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```

Is it the problem with cumint_smoothers?

```{r}
compute_total_loss_smooth_RKHS(params = direct_optim$par,
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)

compute_total_loss_smooth_RKHS(params = c(mean(indirect_policy_DB_small), indirect_policy_DB_small - mean(indirect_policy_DB_small)),
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)

compute_total_loss_smooth_RKHS(params = c(mean(indirect_policy_RKHS_small), indirect_policy_RKHS_small - mean(indirect_policy_RKHS_small)),
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```


```{r}
plot(c(mean(indirect_policy_DB_small), indirect_policy_DB_small - mean(indirect_policy_DB_small)), c(mean(indirect_policy_RKHS_small), indirect_policy_RKHS_small - mean(indirect_policy_RKHS_small)))
```


```{r}
max_iter <- 50
tol <- 1e-4
clip_epsilon_bar <- 300
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

loss_type <- 'or'
krige_adjust <- krige_values_RKHS


params_old <- params_initial_glmnet
params <- params_initial_glmnet

smoothers <- smoothers_RKHS[sub_indices]
cumint_smoothers <- cumint_smoothers_RKHS[sub_indices]
threshold_val <- log(5)

par_initial <- c(mean(indirect_policy_DB_small), indirect_policy_DB_small - mean(indirect_policy_DB_small)) + runif(21,-0.5,0.5)

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = par_initial, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        # gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = par_initial + 0.1, 
                        K = cbind(rep(1,20),diag(20)), 
                        T = data$logWellDepth[sub_indices],
                        krige_adjust = krige_values_RKHS[sub_indices],
                        outcome_resid = rep(0,20),
                        propensity_est = rep(1,20),
                        lambda = 0, 
                        smoothers = smoothers_RKHS[sub_indices], 
                        cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = log(5), 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```

```{r}
plot(params_old, par_initial)
```


```{r}
plot(indirect_policy_DB_small,cbind(rep(1,20),diag(20)) %*% params_old)
```

```{r}
compute_total_loss_smooth_RKHS(params = params_old,
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)

compute_total_loss_smooth_RKHS(params = c(mean(indirect_policy_DB_small), indirect_policy_DB_small - mean(indirect_policy_DB_small)),
                                   K = cbind(rep(1,20),diag(20)), 
                                    T = data$logWellDepth[sub_indices],
                                    krige_adjust = krige_values_RKHS[sub_indices],
                                    outcome_resid = rep(0,20),
                                    propensity_est = rep(1,20),
                                    lambda = 0, 
                                    smoothers = smoothers_RKHS[sub_indices], 
                                    cumint_smoothers = cumint_smoothers_RKHS[sub_indices], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```


##### if we set grand matrix as identity matrix, it should return the indirect method


```{r}
surrogate_type <- "Gaussian"

# loss_type <- "ipw"
# resids <- data$logconcentration_plus_median
# krige_adjust <- rep(0, nrow(data))

loss_type <- 'or'
resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
krige_adjust <- krige_values_RKHS

# loss_type <- 'db'
# resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
# krige_adjust <- krige_values_RKHS

smoothers <- smoothers_RKHS
cumint_smoothers <- cumint_smoothers_RKHS

direct_optim <- optim(par = params_initial_glmnet, 
                      fn = compute_total_loss_smooth_RKHS, 
                      gr = d_compute_total_loss_smooth_RKHS, 
                      K = cbind(rep(1,nrow(data)), diag(nrow(data))), 
                      T = data$logWellDepth,
                      krige_adjust = krige_adjust,
                      outcome_resid = resids,
                      propensity_est = pmax(gps_est,0.001),
                      lambda = 0, 
                      smoothers = smoothers, 
                      cumint_smoothers = cumint_smoothers, 
                      kernel_bw = kernel_bw, 
                      trt_bounds = depth_range,
                      threshold_val = log(5), 
                      clip_epsilon = clip_epsilon,
                      surrogate_type = surrogate_type,
                      loss_type = loss_type,
                      method = "L-BFGS-B", 
                      control = list(maxit = 60, trace = 4))
```

```{r}
3744 * compute_total_loss_smooth_RKHS(params = direct_optim$par,
                                   K = cbind(rep(1,nrow(data)), diag(nrow(data))),
                                   T = data$logWellDepth,
                                    krige_adjust = krige_adjust,
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 0, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)

3744 * compute_total_loss_smooth_RKHS(params = c(mean(indirect_policy_DB), indirect_policy_DB - mean(indirect_policy_DB)),
                                   K = cbind(rep(1,nrow(data)), diag(nrow(data))),
                                   T = data$logWellDepth,
                                    krige_adjust = krige_adjust,
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 0, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(5), 
                                    clip_epsilon = clip_epsilon,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)

d_compute_total_loss_smooth_RKHS(params = c(mean(indirect_policy_DB), indirect_policy_DB - mean(indirect_policy_DB)),
                                 K = cbind(rep(1,nrow(data)), diag(nrow(data))),
                                 T = data$logWellDepth,
                                  krige_adjust = krige_adjust,
                                  outcome_resid = resids,
                                  propensity_est = pmax(gps_est,0.001),
                                  lambda = 0, 
                                  smoothers = smoothers, 
                                  cumint_smoothers = cumint_smoothers, 
                                  kernel_bw = kernel_bw, 
                                  trt_bounds = depth_range,
                                  threshold_val = log(5), 
                                  clip_epsilon = clip_epsilon,
                                  surrogate_type = surrogate_type,
                                  loss_type = loss_type)
```


```{r}
cat('RKHS loss:')
sum(compute_total_loss_smooth(assigned_trt = cbind(rep(1,nrow(data)), diag(nrow(data))) %*% direct_optim$par,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
cat('parametric loss:')
sum(compute_total_loss_smooth(assigned_trt = indirect_policy_DB,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
```

```{r}
plot(indirect_policy_DB, cbind(rep(1,nrow(data)), diag(nrow(data))) %*% direct_optim$par)
```

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(cbind(rep(1,nrow(data)), diag(nrow(data))) %*% direct_optim$par)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(cbind(rep(1,nrow(data)), diag(nrow(data))) %*% direct_optim$par > data$logWellDepth)))
```

```{r}
max_iter <- 50
tol <- 1e-4
clip_epsilon_bar <- 300
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

# loss_type <- "ipw"
# resids <- data$logconcentration_plus_median
# krige_adjust <- rep(0, nrow(data))

loss_type <- 'or'
resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
krige_adjust <- krige_values_RKHS

# loss_type <- 'db'
# resids <- data$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
# krige_adjust <- krige_values_RKHS


gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

params_old <- params_initial_glmnet
params <- params_initial_glmnet

smoothers <- smoothers_RKHS
cumint_smoothers <- cumint_smoothers_RKHS
threshold_val <- log(5)

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = params_old, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = params_old, 
                        K = cbind(rep(1,nrow(data)), diag(nrow(data))), 
                        T = data$logWellDepth,
                        krige_adjust = krige_adjust,
                        outcome_resid = resids,
                        propensity_est = pmax(gps_est,0.001),
                        lambda = 0, 
                        smoothers = smoothers, 
                        cumint_smoothers = cumint_smoothers, 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = log(5), 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```

```{r}
cat('RKHS loss:')
sum(compute_total_loss_smooth(assigned_trt = cbind(rep(1,nrow(data)), diag(nrow(data))) %*% params_old,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
cat('parametric loss:')
sum(compute_total_loss_smooth(assigned_trt = indirect_policy_DB,
                         subject_idx = 1:nrow(k_matrix),
                         T = data$logWellDepth,
                         krige_adjust = krige_adjust,
                         outcome_resid = resids,
                         propensity_est = gps_est,
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         kernel_bw = kernel_bw, 
                         threshold_val = threshold_val, 
                         clip_epsilon = clip_epsilon,
                         surrogate_type = "Gaussian",
                         loss_type = "or"))
cat('\n')
```

```{r}
plot(indirect_policy_DB, cbind(rep(1,nrow(data)), diag(nrow(data))) %*% params_old)
```
```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, as.vector(cbind(rep(1,nrow(data)), diag(nrow(data))) %*% params_old)  > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor(as.vector(cbind(rep(1,nrow(data)), diag(nrow(data))) %*% params_old > data$logWellDepth)))
```


#### Visualize and Test the Loss functions

##### Test DC gradient

###### Test the parametric case


```{r}
# resids <- data$logconcentration_plus_median
# loss_type <- 'ipw'
# krige_adjust <- rep(0, nrow(data))

# resids <- data$logconcentration_plus_median - predict(OR_mean_model_xgboost, design_matrix) - krige_values
# loss_type <- 'or'
# krige_adjust <- krige_values

resids <- data$logconcentration_plus_median - predict(OR_mean_model_xgboost, design_matrix) - krige_values
loss_type <- 'db'
krige_adjust <- krige_values


params_old <- runif(ncol(design_matrix) ,min = -20, max = 50)
```

```{r}
# subset_indice <- c(1,2)
subset_indice <- 1:nrow(data)
# resids <- data$logconcentration_plus_median - predict(OR_mean_model_xgboost, design_matrix) - krige_values
# loss_type <- 'or'
# krige_adjust <- krige_values



d_compute_total_loss_smooth_dc_approx(params = params_old * 1.1, 
                                    params_old = params_old, 
                                    K = design_matrix[subset_indice,], 
                                    T = data$logWellDepth[subset_indice],
                                    krige_adjust = krige_adjust[subset_indice],
                                    outcome_resid = resids[subset_indice],
                                    propensity_est = pmax(gps_est[subset_indice],0.001),
                                    lambda = 0, 
                                    smoothers = smoothers[subset_indice], 
                                    cumint_smoothers = cumint_smoothers[subset_indice], 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```
```{r}
grad(func = function(params){
  compute_total_loss_smooth_dc_approx(params = params, 
                                    params_old = params_old, 
                                    K = design_matrix[subset_indice,], 
                                    T = data$logWellDepth[subset_indice],
                                    krige_adjust = krige_adjust[subset_indice],
                                    outcome_resid = resids[subset_indice],
                                    propensity_est = pmax(gps_est[subset_indice],0.001),
                                    lambda = 0, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)},
  x = params_old * 1.1)
```

When the value goes wild they are totally different, so the derivative might be wrong outside the depth_range
Let's see what happens

```{r}
params = params_old * 1.1
params_old = params_old
K = design_matrix[subset_indice,]
T = data$logWellDepth[subset_indice]
krige_adjust = krige_adjust[subset_indice]
outcome_resid = resids[subset_indice]
propensity_est = pmax(gps_est[subset_indice],0.001)
lambda = 0
smoothers = smoothers[subset_indice]
cumint_smoothers = cumint_smoothers[subset_indice]
kernel_bw = kernel_bw
trt_bounds = depth_range
threshold_val = log(10)
clip_epsilon = clip_epsilon
clip_epsilon_bar = clip_epsilon_bar
```

```{r}
n <- nrow(K)
  
assigned_trt <- as.vector(K %*% params)

assigned_trt_old <- as.vector(K %*% params_old)

d_total_loss_plus <- d_compute_total_loss_smooth_plus(assigned_trt = assigned_trt,
                                                      subject_idx = 1:n,
                                                      T = T,
                                                      krige_adjust = krige_adjust,
                                                      outcome_resid = outcome_resid,
                                                      propensity_est = propensity_est,
                                                      smoothers = smoothers,
                                                      trt_bounds = trt_bounds,
                                                      threshold_val = threshold_val,
                                                      kernel_bw = kernel_bw,
                                                      clip_epsilon = clip_epsilon,
                                                      clip_epsilon_bar = clip_epsilon_bar,
                                                      surrogate_type = surrogate_type,
                                                      loss_type = loss_type) 

d_total_loss_minus_old <- d_compute_total_loss_smooth_minus(assigned_trt = assigned_trt_old,
                                                          subject_idx = 1:n,
                                                          T = T,
                                                          krige_adjust = krige_adjust,
                                                          outcome_resid = outcome_resid,
                                                          propensity_est = propensity_est,
                                                          smoothers = smoothers,
                                                          trt_bounds = trt_bounds,
                                                          threshold_val = threshold_val,
                                                          kernel_bw = kernel_bw,
                                                          clip_epsilon = clip_epsilon,
                                                          clip_epsilon_bar = clip_epsilon_bar,
                                                          surrogate_type = surrogate_type,
                                                          loss_type = loss_type)

if(lambda == 0){
  penalty <- rep(0, ncol(K))
} else {
  if(ncol(K) < nrow(K)){
    penalty <- c(0, (1/n) * lambda * as.vector(params[-1]))  
  }
  else {
    penalty <- c(0, (1/n) * lambda * as.vector(K[,-1] %*% params[-1]))  
  }
}

d_total_loss_dc_approx <- as.numeric(t(K) %*% (d_total_loss_plus - d_total_loss_minus_old))/n + penalty
```

```{r}
d_total_loss_dc_approx
```
```{r}
grad(func = compute_total_loss_smooth_plus, x = assigned_trt,
                                                      subject_idx = 1:n,
                                                      T = T,
                                                      krige_adjust = krige_adjust,
                                                      outcome_resid = outcome_resid,
                                                      propensity_est = propensity_est,
                                                      cumint_smoothers = cumint_smoothers,
                                                      trt_bounds = trt_bounds,
                                                      threshold_val = threshold_val,
                                                      kernel_bw = kernel_bw,
                                                      clip_epsilon = clip_epsilon,
                                                      clip_epsilon_bar = clip_epsilon_bar,
                                                      surrogate_type = surrogate_type,
                                                      loss_type = loss_type)
```



###### Test RKHS case

```{r}
params = params_initial * 1.1
params_old = params_initial
K = k_matrix
T = data$logWellDepth
krige_adjust = rep(0, nrow(data))
outcome_resid = resids
propensity_est = pmax(gps_est,0.001)
lambda = 10
smoothers = smoothers
cumint_smoothers = cumint_smoothers
kernel_bw = kernel_bw
trt_bounds = depth_range
threshold_val = log(10)
clip_epsilon = clip_epsilon
clip_epsilon_bar = clip_epsilon_bar
surrogate_type = surrogate_type
loss_type = loss_type
```

```{r}
assigned_trt <- as.vector(K %*% params)
  
assigned_trt_old <- as.vector(K %*% params_old)
```

```{r}
subject_idx <- 1
```

```{r}
# compute with grad

a = grad(func = function(params){
  total_loss_plus <- compute_total_loss_smooth_plus(assigned_trt = assigned_trt[subject_idx],
                                                    subject_idx = subject_idx,
                                                    T = T,
                                                    krige_adjust = krige_adjust,
                                                    outcome_resid = outcome_resid,
                                                    propensity_est = propensity_est,
                                                    cumint_smoothers = cumint_smoothers,
                                                    trt_bounds = trt_bounds,
                                                    threshold_val = threshold_val,
                                                    kernel_bw = kernel_bw,
                                                    clip_epsilon = clip_epsilon,
                                                    clip_epsilon_bar = clip_epsilon_bar,
                                                    surrogate_type = surrogate_type,
                                                    loss_type = loss_type) 
  d_total_loss_minus_old <- d_compute_total_loss_smooth_minus(assigned_trt = assigned_trt_old[subject_idx],
                                                              subject_idx = subject_idx,
                                                              T = T,
                                                              krige_adjust = krige_adjust,
                                                              outcome_resid = outcome_resid,
                                                              propensity_est = propensity_est,
                                                              smoothers = smoothers,
                                                              trt_bounds = trt_bounds,
                                                              threshold_val = threshold_val,
                                                              kernel_bw = kernel_bw,
                                                              clip_epsilon = clip_epsilon,
                                                              clip_epsilon_bar = clip_epsilon_bar,
                                                              surrogate_type = surrogate_type,
                                                              loss_type = loss_type)
  return(mean(total_loss_plus - d_total_loss_minus_old * assigned_trt[subject_idx]))},
  x = params
  )
```


##### DC approx loss function

```{r}
compute_total_loss_smooth_dc_approx(params = params_old * 1.1, 
                                    params_old = params_old, 
                                    K = k_matrix, 
                                    T = data$logWellDepth,
                                    krige_adjust = rep(0, nrow(data)),
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 10, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```

```{r}
# This is too slow and cannot be implemented
grad(func = compute_total_loss_smooth_dc_approx, x = params_old * 1.1, params_old = params_old, 
                                    K = k_matrix, 
                                    T = data$logWellDepth,
                                    krige_adjust = rep(0, nrow(data)),
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 10, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```

```{r}
d = d_compute_total_loss_smooth_dc_approx(params = params_old * 1.1, 
                                    params_old = params_old, 
                                    K = k_matrix, 
                                    T = data$logWellDepth,
                                    krige_adjust = rep(0, nrow(data)),
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 10, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```


```{r}
surrogate_type = "Gaussian"
loss_type = "ipw"

i <- 10
clip_epsilon_bar <- 200

clip_epsilon <- 10
threshold_val <- log(10)

knots <-  seq(depth_range[1] - 2, depth_range[2] + 2, length.out = 100)

kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))


total_loss_plus <- compute_total_loss_smooth_plus(assigned_trt = assigned_trt,
                                                    subject_idx = 1:n,
                                                    T = T,
                                                    krige_adjust = krige_adjust,
                                                    outcome_resid = outcome_resid,
                                                    propensity_est = propensity_est,
                                                    cumint_smoothers = cumint_smoothers,
                                                    trt_bounds = trt_bounds,
                                                    threshold_val = threshold_val,
                                                    kernel_bw = kernel_bw,
                                                    clip_epsilon = clip_epsilon,
                                                    clip_epsilon_bar = clip_epsilon_bar,
                                                    surrogate_type = surrogate_type,
                                                    loss_type = loss_type) 
  
  
d_total_loss_minus_old <- d_compute_total_loss_smooth_minus(assigned_trt = assigned_trt_old,
                                                              subject_idx = 1:n,
                                                              T = T,
                                                              krige_adjust = krige_adjust,
                                                              outcome_resid = outcome_resid,
                                                              propensity_est = propensity_est,
                                                              smoothers = smoothers,
                                                              trt_bounds = trt_bounds,
                                                              threshold_val = threshold_val,
                                                              kernel_bw = kernel_bw,
                                                              clip_epsilon = clip_epsilon,
                                                              clip_epsilon_bar = clip_epsilon_bar,
                                                              surrogate_type = surrogate_type,
                                                              loss_type = loss_type)
  
total_loss_plus - d_total_loss_minus_old * assigned_trt
```

##### Loss function

```{r}
surrogate_type = "Gaussian"
loss_type = "ipw"
i <- 2
clip_epsilon_bar <- 200
clip_epsilon <- 10
threshold_val <- log(10)

knots <-  seq(depth_range[1] - 2, depth_range[2] + 2, length.out = 100)

kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

loss_type <- "ipw"

plot(knots, sapply(knots, compute_total_loss_smooth,  subject_idx = i,
                            T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = clip_epsilon,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                            T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = clip_epsilon,
                            clip_epsilon_bar = clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = clip_epsilon,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                             T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = clip_epsilon,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type) - sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = clip_epsilon,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

```



##### Visualize loss function

```{r}
surrogate_type = "Gaussian"
loss_type = "ipw"
i = 10
clip_epsilon_bar = 50
knots <- seq(-6, 6, length.out = 100)

plot(knots, sapply(knots, compute_total_loss_smooth,  subject_idx = i,
                            T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type) - sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

```


##### Test Gradient of DC

```{r}
grad(func = compute_total_loss_smooth_dc_approx, x = params_old * 1.1, params_old = params_old, 
                                    K = k_matrix, 
                                    T = data$logWellDepth,
                                    krige_adjust = rep(0, nrow(data)),
                                    outcome_resid = resids,
                                    propensity_est = pmax(gps_est,0.001),
                                    lambda = 10, 
                                    smoothers = smoothers, 
                                    cumint_smoothers = cumint_smoothers, 
                                    kernel_bw = kernel_bw, 
                                    trt_bounds = depth_range,
                                    threshold_val = log(10), 
                                    clip_epsilon = clip_epsilon,
                                    clip_epsilon_bar = clip_epsilon_bar,
                                    surrogate_type = surrogate_type,
                                    loss_type = loss_type)
```


```{r}

```


##### Test gradient os plus, minus

```{r}

knots <- seq(depth_range[1] - 20, depth_range[2] + 20, length.out = 200)
i <- 100

d1 <- sapply(knots, d_compute_total_loss_smooth_plus, subject_idx = i, T = data$logWellDepth, 
                   krige_adjust = rep(0, nrow(data)), outcome_resid =  data$logconcentration_plus_median, 
                   propensity_est = gps_est,
                        smoothers = smoothers, 
                        
                        kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                        trt_bounds = depth_range,
                        threshold_val = log(10), 
                        clip_epsilon = 1,
                        clip_epsilon_bar = 400,
                        surrogate_type = "Gaussian",
                        loss_type = "ipw")

plot(knots, d1)

d2 <- sapply(knots, function(assigned_trt){
  grad(func = compute_total_loss_smooth_plus, x = assigned_trt, subject_idx = i, T = data$logWellDepth, 
                  krige_adjust = rep(0, nrow(data)), outcome_resid =  data$logconcentration_plus_median, 
                  propensity_est = gps_est, cumint_smoothers = cumint_smoothers,
                  kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                  trt_bounds = depth_range,
                  threshold_val = log(10), 
                  clip_epsilon = 1,
                  clip_epsilon_bar = 400,
                  surrogate_type = "Gaussian",
                  loss_type = "ipw")
})
  
plot(knots, d2)

                                           
```

```{r}
d1 - d2
```


```{r}

knots <- seq(depth_range[1] - 20, depth_range[2] + 20, length.out = 200)
i <- 1
d3 <- sapply(knots, d_compute_total_loss_smooth_minus, subject_idx = i, T = data$logWellDepth, 
                   krige_adjust = rep(0, nrow(data)), outcome_resid =  data$logconcentration_plus_median, 
                   propensity_est = gps_est,
                        smoothers = smoothers, 
                        
                        kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                        trt_bounds = depth_range,
                        threshold_val = log(10), 
                        clip_epsilon = 1,
                        clip_epsilon_bar = 400,
                        surrogate_type = "Gaussian",
                        loss_type = "ipw")
plot(knots, d3)

d4 <- sapply(knots, function(assigned_trt){
  grad(func = compute_total_loss_smooth_minus, x = assigned_trt, subject_idx = i, T = data$logWellDepth, 
                  krige_adjust = rep(0, nrow(data)), outcome_resid =  data$logconcentration_plus_median, 
                  propensity_est = gps_est, cumint_smoothers = cumint_smoothers,
                  kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                  trt_bounds = depth_range,
                  threshold_val = log(10), 
                  clip_epsilon = 1,
                  clip_epsilon_bar = 400,
                  surrogate_type = "Gaussian",
                  loss_type = "ipw")
})

plot(knots, d4)

d3 - d4                                           
```


```{r}
grad(function(coefs){
  assigned_trt <- design_matrix %*% coefs
  return(sum(compute_total_loss_smooth_plus(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data$X),
                                 T = data$T,
                                 krige_adjust = krige_values,
                                 outcome_resid = resids,
                                 propensity_est = gps_est,
                                 cumint_smoothers = cumint_smoothers,
                                 trt_bounds = c(-4, 4),
                                 threshold_val = threshold_val,
                                 kernel_bw = kernel_bw,
                                 clip_epsilon = 1,
                                 clip_epsilon_bar,
                                 surrogate_type = "Gaussian",
                                 loss_type = "db")))},
  x = coefs)
```


