---
title: "direct_method_RKHS"
author: "Xindi Lin"
date: "2025-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(betareg)
library(ranger)
library(mgcv)
library(kernlab)
library(glmnet)
source('data_generation.R')
```

##### preparation for simulation

```{r}
set.seed(1998)
trt_range = c(0,1)
trt_step = 0.01

epsilon = 1 # expand loss function outside trt-range

#### data generation setting
range = 0.2
phi = 10 # precision for beta distribution 
dim = 4
nugget = 0.5
data_linearity = FALSE
# data_linearity = TRUE 
## when the outcome regression is linear, the beta distributed is not appropriate for treatment
## when the outcome regression is linear, the gamma distribution is more appropriate, but gamma distribution might violate positivity 

#### threshold

threshold = 1 # for non-linear simulation, the mean of observed outcome is about one 

#### outcome regression

outcome_fitted_model = 'gam'
# outcome_fitted_model = 'lm'
# outcome_fitted_model = 'rf'




### Step 1: Data Generation

data = data_generation_grid_non_randomized(a = 1, d = 50, range = 0.2, phi = 10, nu = 1.5, dim = 4, nugget = 0.5, linear = data_linearity)

#### re-order data

data = reorder_data(data,order = "coordinate")


### Step 2: Nuisance Estimation

fit.propensity = betareg(trt ~ x1 + x2 + x3 + x4 | 1, data = data, link = 'logit', link.phi = 'identity')
# fit.outcome = gam(logy~s(trt)+s(x1)+s(x2)+s(x3)+s(x4),data = data)
fit.outcome = gam(logy~trt+x1+x2+x3+x4,data = data)
## check fitted outcome regression model
# plot(data$logy, fit.outcome$fitted.values, xlab = 'observed outcome',ylab = 'fitted outcome')
# abline(0,1,col='red')


#### Step 3 Fit Gaussian Processes, Kriging Residuals

resid.outcome = data$logy - fit.outcome$fitted.values
fitGp = fit_model(y = resid.outcome, locs = data[,c('coord_x','coord_y')])
krige_values = leave_one_out_kriging(data = data, resid = resid.outcome)

# # check fitted GP
# plot(data$epsilon,krige_values,xlab='true GP',ylab = 'fitted GP')
# abline(0,1,col='red')


### Step 4: Get Pre-computed prediction matrix for integral, and its smoothed version

pred_matrix = get_precompute_pred_matrix(data = data, fit.outcome = fit.outcome, trt_range = trt_range, trt_step = trt_step)

cum_smoothers = build_cumint_smoothers(trt_values, pred_matrix, method = "splinefun", spar = NULL)

smoothers = build_trt_smoothers(trt_values, pred_matrix, smooth_method = "splinefun",spar = NULL)

### Step 5: Get Pre-computed Propensity, and residuals

prop_estimation = get_prop_score(data = data, fit.propensity.beta = fit.propensity, propensity_lower_bound = 0.5)
# sum(prop_estimation==0.5)
resids = (data$logy - fit.outcome$fitted.values - krige_values)

```


##### beta kernels

```{r}
#' Beta Kernel Function
#'
#' @param y A numeric vector of evaluation points (must be in [0, 1]).
#' @param x A single numeric value for the kernel's center (must be in (0, 1)).
#' @param h A single positive numeric value for the bandwidth.
#'
#' @return A numeric vector of the kernel density values.
#'
beta_kernel <- function(y, x, h) {
  if (h <= 0) {
    stop("Bandwidth 'h' must be positive.")
  }
  if (x <= 0 || x >= 1) {
    stop("Center 'x' must be in the interval (0, 1).")
  }
  
  # The shape parameters for the Beta distribution
  shape1 <- x / h
  shape2 <- (1 - x) / h
  
  # Use R's built-in dbeta function
  # It correctly returns 0 for any y outside [0, 1].
  return(dbeta(y, shape1 = shape1, shape2 = shape2))
}
```

```{r}
#' Indefinite Integral of the Beta Kernel (CDF)
#'
#' @param t A numeric vector of upper integration limits.
#' @param x A single numeric value for the kernel's center.
#' @param h A single positive numeric value for the bandwidth.
#'
#' @return A numeric vector of the integral values from 0 to t.
#'
integral_kernel_cdf <- function(t, x, h) {
  if (h <= 0) {
    stop("Bandwidth 'h' must be positive.")
  }
  if (x <= 0 || x >= 1) {
    stop("Center 'x' must be in the interval (0, 1).")
  }
  
  shape1 <- x / h
  shape2 <- (1 - x) / h
  
  return(pbeta(t, shape1 = shape1, shape2 = shape2))
}
```


##### integral/derivative of kernel functions

```{r}
kernel_integral_plus = function(trt,trt_obs,bandwidth){
  ifelse(trt >= trt_obs-bandwidth, (1 / (2 * bandwidth) * (trt - (trt_obs - bandwidth))), 0)
}

kernel_integral_plus_derivative = function(trt,trt_obs,bandwidth){
  ifelse(trt >= trt_obs-bandwidth, (1 / (2 * bandwidth)), 0)
}

kernel_integral_minus = function(trt,trt_obs,bandwidth){
  ifelse(trt >= trt_obs+bandwidth, (1 / (2 * bandwidth) * (trt - (trt_obs + bandwidth))), 0)
}

kernel_integral_minus_derivative = function(trt,trt_obs,bandwidth){
  ifelse(trt >= trt_obs+bandwidth, (1 / (2 * bandwidth)), 0)
}
```

```{r}
# test integral/derivative of kernel functions
kernel_integral_plus(trt = c(-0.6,0.5),trt_obs = c(0,0),bandwidth = 0.7) - kernel_integral_minus(trt = c(-0.6,0.5),trt_obs = c(0,0),bandwidth = 0.7)

kernel_integral_minus_derivative(c(0.9,0.6),c(0,0),0.7)
```


##### nu_plus and nu_minus and nu_minus's derivative

```{r}
nu_plus = function(trt_assigned, trt_obs, scaled_residual, bandwidth, threshold, col_idx, krige_values,trt_range, cum_smoothers){
  # term of threshold integral
  term_1 = threshold * (trt_assigned - trt_range[1])
  
  # term of doubly adjusted
  scaled_residual_plus = pmax(0, scaled_residual)
  scaled_residual_minus = -pmin(0, scaled_residual)
  kernel_int_plus = kernel_integral_plus(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  kernel_int_minus = kernel_integral_minus(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
    
  term_2 = scaled_residual_plus * kernel_int_minus + scaled_residual_minus * kernel_int_plus
  
  # term of integral of outcome regression
  term_3 = mapply(
      FUN = integrate_outcome_gam_storage_smooth, 
      upper = trt_assigned, 
      col_idx = col_idx, 
      krige_value = krige_values, 
        MoreArgs = list(
        lower = trt_range[1], # 'lower' is also constant in this case
        cum_smoothers = cum_smoothers,
        smooth_type = "cum"
      )
    )
  # here it is plus term_2 not minus term_2
  if(sum(is.na(term_1 + term_2 - term_3)>0)){
    cat("NA in nu_plus")
    browser()
  }
  return(term_1 + term_2 - term_3)
}
```

```{r}
nu_plus_derivative = function(trt_assigned, trt_obs, scaled_residual, bandwidth, threshold, col_idx, krige_values,trt_range, smoothers){
  # term of threshold integral
  term_1 = rep(threshold, length(trt_assigned))
  
  # term of doubly adjusted
  scaled_residual_plus = pmax(0, scaled_residual)
  scaled_residual_minus = -pmin(0, scaled_residual)
  kernel_int_plus_derivative = kernel_integral_plus_derivative(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  kernel_int_minus_derivative = kernel_integral_minus_derivative(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  
  term_2 = scaled_residual_minus * kernel_int_plus_derivative + scaled_residual_plus * kernel_int_minus_derivative
  
  # term of outcome regression
  term_3 = mapply(
      FUN = outcome_gam_storage_smooth, 
      upper = trt_assigned, 
      col_idx = col_idx, 
      krige_value = krige_values, 
        MoreArgs = list(
        lower = trt_range[1], # 'lower' is also constant in this case
        smoothers = smoothers
      )
    )
  # here it is plus term_2 not minus term_2
  
  if(sum(is.na(term_1 + term_2 - term_3)>0)){
    cat("NA in nu_plus_derivative")
    browser()
  }
  return(term_1 + term_2 - term_3)
}
```

```{r}

```


```{r}
nu_minus = function(trt_assigned, trt_obs, scaled_residual, bandwidth){
  # term of doubly adjusted
  scaled_residual_plus = pmax(0, scaled_residual)
  scaled_residual_minus = -pmin(0, scaled_residual)
  kernel_int_plus = kernel_integral_plus(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  kernel_int_minus = kernel_integral_minus(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
    
  term_2 = scaled_residual_plus * kernel_int_plus + scaled_residual_minus * kernel_int_minus
  if(sum(is.na(term_2)>0)){
    cat("NA in nu_minus")
    browser()
  }
  return(term_2)
}

nu_minus_derivative = function(trt_assigned, trt_obs, scaled_residual, bandwidth){
  # term of doubly adjusted
  scaled_residual_plus = pmax(0, scaled_residual)
  scaled_residual_minus = -pmin(0, scaled_residual)
  kernel_int_plus_derivative = kernel_integral_plus_derivative(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  kernel_int_minus_derivative = kernel_integral_minus_derivative(trt = trt_assigned, trt_obs = trt_obs, bandwidth = bandwidth)
  
  term_2 = scaled_residual_plus * kernel_int_plus_derivative + scaled_residual_minus * kernel_int_minus_derivative
  if(sum(is.na(term_2)>0)){
    cat("NA in nu_minus_derivative")
    browser()
  }
  return(term_2)
}
```

```{r}
scaled_residuals = resids/prop_estimation
```


```{r}
# check if it is the same as the previous function
nu_plus(trt_assigned = c(0.2,0.5), trt_obs = data$trt[c(1,2)], scaled_residual = scaled_residuals[c(1,2)], bandwidth = 0.07, threshold = threshold,col_idx = c(1,2), krige_values = krige_values[c(1,2)], trt_range = trt_range, cum_smoothers = cum_smoothers)

nu_minus(trt_assigned = c(0.2,0.5), trt_obs = data$trt[c(1,2)], scaled_residual = scaled_residuals[c(1,2)], bandwidth = 0.07)
```

```{r}
nu_minus_derivative(trt_assigned = c(10,10), trt_obs = data$trt[c(1,2)], scaled_residual = scaled_residuals[c(1,2)], bandwidth = 0.07)
```


```{r}
trt_1 = 0.5
trt_2 = 0.45

nu_plus(trt_assigned = c(trt_1), trt_obs = data$trt[c(1)], scaled_residual = scaled_residuals[c(1)], bandwidth = 0.07, threshold = threshold,col_idx = c(1), krige_values = krige_values[c(1)], trt_range = trt_range, cum_smoothers = cum_smoothers)

nu_plus(trt_assigned = c(trt_2), trt_obs = data$trt[c(1)], scaled_residual = scaled_residuals[c(1)], bandwidth = 0.07, threshold = threshold,col_idx = c(1), krige_values = krige_values[c(1)], trt_range = trt_range, cum_smoothers = cum_smoothers)  
  
nu_plus(trt_assigned = c(trt_2), trt_obs = data$trt[c(1)], scaled_residual = scaled_residuals[c(1)], bandwidth = 0.07, threshold = threshold,col_idx = c(1), krige_values = krige_values[c(1)], trt_range = trt_range, cum_smoothers = cum_smoothers) + nu_plus_derivative(trt_assigned = c(trt_2), trt_obs = data$trt[c(1)], scaled_residual = scaled_residuals[c(1)], bandwidth = 0.07, threshold = threshold,col_idx = c(1), krige_values = krige_values[c(1)], trt_range = trt_range, smoothers = smoothers) * (trt_1 - trt_2)

```


```{r}
total_loss_db_storage_smooth(trt_assigned = c(0.2,0.2), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation, cum_smoothers, trt_range, threshold, bandwidth)
```

##### loss_plus, loss_minus, loss_minus_derivative

```{r}
total_loss_db_storage_plus =  function(trt_assigned, col_idx, data, krige_values, resids, prop_estimation, cum_smoothers, bandwidth, threshold, epsilon){
  
  # preparation
  
  ## create the vector to store the loss
  loss_vec = rep(NA,length(col_idx))
  ## find the corresponding index
  index_left_clip = trt_assigned<trt_range[1]
  index_right_clip = trt_assigned>trt_range[2]
  index_no_clip = (trt_assigned >= trt_range[1]) & (trt_assigned<=trt_range[2])
  ## get right clipped trt, this one is used to calculate the integral
  trt_assigned_right_clipped = pmin(trt_assigned, trt_range[2])
  
  
  scaled_residuals = resids[col_idx]/prop_estimation[col_idx]
  
  # calculate the loss for index_left_clip
  if(sum(index_left_clip)>0){
    loss_vec[index_left_clip] = -2*epsilon*(trt_assigned[index_left_clip] - trt_range[1])
  }
  
# calculate the loss for index_no_clip, index right clip
  if(sum(!index_left_clip)>0){
    loss_vec[!index_left_clip] = nu_plus(trt_assigned = trt_assigned_right_clipped[!index_left_clip], 
                                        trt_obs = data[col_idx,"trt"][!index_left_clip], 
                                        scaled_residual = scaled_residuals[!index_left_clip], 
                                        bandwidth = bandwidth, 
                                        threshold = threshold,
                                        col_idx = col_idx[!index_left_clip], 
                                        krige_values = krige_values[col_idx][!index_left_clip],
                                        trt_range = trt_range, 
                                        cum_smoothers = cum_smoothers)
                                        
  }
  
  
  
  if(sum(index_right_clip)>0){
    # add the right clipped additional loss
    # pred_matrix is not the input!!
    epsilon_bar = pmax(threshold - pred_matrix[which(trt_values==trt_range[2]),col_idx][index_right_clip] + abs(scaled_residuals[index_right_clip]/(2*bandwidth)), abs(scaled_residuals[index_right_clip]/(2*bandwidth)))
    
    loss_vec[index_right_clip] = loss_vec[index_right_clip] +  (epsilon_bar +2*epsilon)*(trt_assigned[index_right_clip]-trt_range[2])
    
  }
  
  if(sum(is.na(loss_vec)>0)){
    cat("NA in total_loss_db_storage_plus")
    browser()
  }
  return(loss_vec)
}

```

```{r}
total_loss_db_storage_plus_derivative =  function(trt_assigned, col_idx, data, krige_values, resids, prop_estimation, smoothers, bandwidth, threshold, epsilon){
  
  # preparation
  
  ## create the vector to store the loss
  loss_vec = rep(NA,length(col_idx))
  ## find the corresponding index
  index_left_clip = trt_assigned<trt_range[1]
  index_right_clip = trt_assigned>trt_range[2]
  index_no_clip = (trt_assigned >= trt_range[1]) & (trt_assigned<=trt_range[2])
  ## get right clipped trt, this one is used to calculate the integral
  trt_assigned_right_clipped = pmin(trt_assigned, trt_range[2])
  
  
  scaled_residuals = resids[col_idx]/prop_estimation[col_idx]
  
  # calculate the loss for index_left_clip
  if(sum(index_left_clip)>0){
    loss_vec[index_left_clip] = -2*epsilon
  }
  
# calculate the loss for index_no_clip, index right clip
  if(sum(!index_left_clip)>0){
    loss_vec[!index_left_clip] = nu_plus_derivative(trt_assigned = trt_assigned_right_clipped[!index_left_clip], 
                                        trt_obs = data[col_idx,"trt"][!index_left_clip], 
                                        scaled_residual = scaled_residuals[!index_left_clip], 
                                        bandwidth = bandwidth, 
                                        threshold = threshold,
                                        col_idx = col_idx[!index_left_clip], 
                                        krige_values = krige_values[col_idx][!index_left_clip],
                                        trt_range = trt_range, 
                                        smoothers = smoothers)
  }
  
  if(sum(index_right_clip)>0){
    # add the right clipped additional loss
    
    epsilon_bar = pmax(threshold - pred_matrix[which(trt_values==trt_range[2]),col_idx][index_right_clip] + abs(scaled_residuals[index_right_clip]/(2*bandwidth)), abs(scaled_residuals[index_right_clip]/(2*bandwidth)))
    # it should not be
    # loss_vec[index_right_clip] = loss_vec[index_right_clip] +  (epsilon_bar +2*epsilon)
    # but rather be 
    loss_vec[index_right_clip] = epsilon_bar +2*epsilon
  }
  
  if(sum(is.na(loss_vec)>0)){
    cat("NA in total_loss_db_storage_plus_derivative")
    browser()
  }
  return(loss_vec)
}

```

```{r}
1
```


```{r}
total_loss_db_storage_minus =  function(trt_assigned, col_idx, data, krige_values, resids, prop_estimation,bandwidth,threshold,epsilon){
  
  # preparation
  
  ## create the vector to store the loss
  loss_vec = rep(NA,length(col_idx))
  ## find the corresponding index
  index_left_clip = trt_assigned<trt_range[1]
  index_right_clip = trt_assigned>trt_range[2]
  index_no_clip = (trt_assigned >= trt_range[1]) & (trt_assigned<=trt_range[2])
  ## get right clipped trt, this one is used to calculate the integral
  trt_assigned_right_clipped = pmin(trt_assigned, trt_range[2])
  ## get scaled residual
  scaled_residuals = resids[col_idx]/prop_estimation[col_idx]
  
  # calculate the loss for index_left_clip
  if(sum(index_left_clip)>0){
    loss_vec[index_left_clip] = -2*epsilon*(trt_assigned[index_left_clip] - trt_range[1]) - epsilon + epsilon * exp(trt_assigned[index_left_clip] - trt_range[1])
  }
  
# calculate the loss for index_no_clip, index right clip
  if(sum(!index_left_clip)>0){
    loss_vec[!index_left_clip] = nu_minus(trt_assigned = trt_assigned_right_clipped[!index_left_clip], 
                                        trt_obs = data[col_idx,"trt"][!index_left_clip], 
                                        scaled_residual = scaled_residuals[!index_left_clip], 
                                        bandwidth = bandwidth)
  }
  
  if(sum(index_right_clip)>0){
    # add the right clipped additional loss
    
    epsilon_bar = pmax(threshold - pred_matrix[which(trt_values==trt_range[2]),col_idx][index_right_clip] + abs(scaled_residuals[index_right_clip]/(2*bandwidth)), abs(scaled_residuals[index_right_clip]/(2*bandwidth)))
    
    loss_vec[index_right_clip] = loss_vec[index_right_clip] +  (epsilon_bar +2*epsilon)*(trt_assigned[index_right_clip] - trt_range[2]) - epsilon +epsilon*exp(-trt_assigned[index_right_clip] + trt_range[2])
  }
  
  if(sum(is.na(loss_vec)>0)){
    cat("NA in total_loss_db_storage_minus")
    browser()
  }
  return(loss_vec)
}

```

```{r}
total_loss_db_storage_minus_derivative =  function(trt_assigned, col_idx, data, krige_values, resids, prop_estimation,bandwidth,threshold,epsilon){
  # preparation
  ## create the vector to store the loss
  loss_vec = rep(NA,length(col_idx))
  ## find the corresponding index
  index_left_clip = trt_assigned<trt_range[1]
  index_right_clip = trt_assigned>trt_range[2]
  index_no_clip = (trt_assigned >= trt_range[1]) & (trt_assigned<=trt_range[2])
  ## get right clipped trt, this one is used to calculate the integral
  trt_assigned_right_clipped = pmin(trt_assigned, trt_range[2])
  ## get scaled residual
  scaled_residuals = resids[col_idx]/prop_estimation[col_idx]
  
  # calculate the loss for index_left_clip
  if(sum(index_left_clip)>0){
    loss_vec[index_left_clip] = -2*epsilon + epsilon * exp(trt_assigned[index_left_clip] - trt_range[1])
  }
  
# calculate the loss for index_no_clip, index right clip
  if(sum(!index_left_clip)>0){
    
    loss_vec[!index_left_clip] = nu_minus_derivative(trt_assigned = trt_assigned_right_clipped[!index_left_clip], 
                                            trt_obs = data[col_idx,"trt"][!index_left_clip], 
                                            scaled_residual = scaled_residuals[!index_left_clip], 
                                            bandwidth = bandwidth)
  }
  
  if(sum(index_right_clip)>0){
    # add the right clipped additional loss
    
    epsilon_bar = pmax(threshold - pred_matrix[which(trt_values==trt_range[2]),col_idx][index_right_clip] + abs(scaled_residuals[index_right_clip]/(2*bandwidth)), abs(scaled_residuals[index_right_clip]/(2*bandwidth)))
    # it should not be
    # loss_vec[index_right_clip] = loss_vec[index_right_clip] +  (epsilon_bar +2*epsilon) - epsilon*exp(-trt_assigned[index_right_clip]+trt_range[2])
    # but rather be 
    loss_vec[index_right_clip] =  (epsilon_bar +2*epsilon) - epsilon*exp(-trt_assigned[index_right_clip]+trt_range[2])
  }
  
  if(sum(is.na(loss_vec)>0)){
    cat("NA in total_loss_db_storage_minus_derivative")
    browser()
  }
  
  return(loss_vec)
  
}
```

```{r}
total_loss_db_storage_plus(trt_assigned = c(2,0.2), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation, cum_smoothers = cum_smoothers,threshold = threshold,bandwidth = 0.07, epsilon = 1)

total_loss_db_storage_minus(trt_assigned = c(2,0.2), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation,threshold = threshold,bandwidth = 0.07, epsilon = 1)
```


```{r}
# early version
total_loss_db_storage_plus(trt_assigned = c(2,0.5), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation, cum_smoothers = cum_smoothers,threshold = threshold,bandwidth = 0.07, epsilon = 1)

total_loss_db_storage_minus(trt_assigned = c(2,0.5), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation,threshold = threshold,bandwidth = 0.07, epsilon = 1)
```


```{r}
total_loss_db_storage_smooth(trt_assigned = c(2,0.5), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation, cum_smoothers = cum_smoothers, trt_range = trt_range, threshold = threshold,bandwidth = 0.07)
```

```{r}
total_loss_db_storage_minus_derivative(trt_assigned = c(0.2,0.2), col_idx = c(1,2), data, krige_values = krige_values[c(1,2)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1)
```



```{r}
# check the loss_minus derivative

trt_1 = 21
trt_2 = 15

total_loss_db_storage_minus(trt_assigned = c(trt_1), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1)

total_loss_db_storage_minus(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1) + total_loss_db_storage_minus_derivative(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1) * (trt_1 - trt_2)

total_loss_db_storage_minus_derivative(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1)

total_loss_db_storage_minus_derivative(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, threshold = threshold,bandwidth = 0.07, epsilon = 1) * (trt_1 - trt_2)
```
We notice that total_loss_db_storage_minus is not only convex, it is actually piecewise linear, so sometimes it is exact.


```{r}
# check the loss_plus derivative

trt_1 = 2
trt_2 = 1.4


total_loss_db_storage_plus(trt_assigned = c(trt_1), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation,cum_smoothers = cum_smoothers, threshold = threshold,bandwidth = 0.07, epsilon = 1)

total_loss_db_storage_plus(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation,cum_smoothers = cum_smoothers, threshold = threshold,bandwidth = 0.07, epsilon = 1) + total_loss_db_storage_plus_derivative(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, smoothers = smoothers, threshold = threshold,bandwidth = 0.07, epsilon = 1) * (trt_1 - trt_2)

total_loss_db_storage_plus_derivative(trt_assigned = c(trt_2), col_idx = c(1), data, krige_values = krige_values[c(1)], resids, prop_estimation, smoothers = smoothers, threshold = threshold,bandwidth = 0.07, epsilon = 1) * (trt_1 - trt_2)
```

##### DC algorithm


##### Step of SVM RKHS

```{r}
# Fix gamma as the median 
dist_matrix = rdist(data[,c('x1','x2','x3','x4')])
gamma = median(c(dist_matrix))
```


```{r}
#' Construct the Kernel Matrix
#' @param S A matrix of features (X, U).
#' @param gamma Bandwidth of the RBF kernel.
#' @return The n x n kernel matrix K.
construct_kernel_matrix <- function(S, gamma) {
  rbf <- rbfdot(sigma = 1/(gamma^2))
  return(kernelMatrix(rbf, S))
}
```

###### initialization

```{r}
initial_values_db_storage_smooth = mclapply(
  X = 1:nrow(data),
  FUN = function(idx){
    optim_result = optim(
      # Initial guess for the trt_assigned vector
      par = 0.5, 
      # The function to minimize
      fn = total_loss_db_storage_smooth, 
      # --- Arguments for the optimizer ---
      # method = "Brent", # Brent only gives two decimal result?
      method = "L-BFGS-B",
      lower = 0, # Lower bound for each element is 0
      upper = 1, # Upper bound for each element is 1
      col_idx = idx, 
      data = data, 
      krige_values = krige_values, 
      resids = resids, 
      prop_estimation = prop_estimation, 
      cum_smoothers = cum_smoothers,
      trt_range = trt_range,
      threshold = threshold,
      bandwidth = bandwidth
    )
    return(optim_result$par)
  }
)
```

```{r}
S = cbind(data[,c('x1','x2','x3','x4')], as.matrix(krige_values))
K = construct_kernel_matrix(S = as.matrix(S), gamma = gamma)

svm_initial_cv_fit =  cv.glmnet(K, initial_values_db_storage_smooth, alpha = 0)
```

```{r}
svm_initial_cv_fit
```

```{r}
as.numeric(coef(svm_initial_cv_fit, s = "lambda.min"))
coef(svm_initial)
```

```{r}
params_old = as.numeric(coef(svm_initial_cv_fit, s = "lambda.min"))
params = as.numeric(coef(svm_initial_cv_fit, s = "lambda.min"))
```

###### DC loss function

```{r}
#' Approximation loss function of DC algorithm
#' @param params initial value of parameters.
#' @param K The n x n kernel matrix K.
#' @param gamma Bandwidth of the RBF kernel.
#' @return The n x n kernel matrix K.
loss_total_DC_db_storage = function(params, params_old, K, data, krige_values, resids, 
                                    prop_estimation, lambda, smoothers, cum_smoothers, 
                                    bandwidth, threshold, epsilon){
  # print('calculate loss')
  
  n = nrow(K)
  b =  params[1]
  eta = params[2:(n+1)]
  trt_assigned = as.vector(K %*% eta + b)
  
  eta_old = params_old[2:(n+1)]
  b_old = params_old[1]
  trt_assigned_old = as.vector(K %*% eta_old + b_old)
  
  
  L_plus_total = total_loss_db_storage_plus(trt_assigned = trt_assigned, 
                                            col_idx = 1:n, 
                                            data = data, 
                                            krige_values = krige_values, 
                                            resids = resids, 
                                            prop_estimation = prop_estimation, 
                                            cum_smoothers = cum_smoothers,
                                            bandwidth = bandwidth,
                                            threshold = threshold,
                                            epsilon = epsilon)
  
  grad_L_minus_total = total_loss_db_storage_minus_derivative(trt_assigned = trt_assigned_old, 
                                                              col_idx = 1:n, 
                                                              data = data, 
                                                              krige_values = krige_values, 
                                                              resids = resids, 
                                                              prop_estimation = prop_estimation, 
                                                              bandwidth = bandwidth,
                                                              threshold = threshold,
                                                              epsilon = epsilon)
        
  # grad_L_minus_total[is.na(grad_L_minus_total) | is.infinite(grad_L_minus_total)] <- 0
  
  val = mean(L_plus_total -  grad_L_minus_total * (trt_assigned - trt_assigned_old)) + (lambda / 2) * (t(eta) %*% K %*% eta)
  if (val< (-100)){
    cat('loss_total_DC is way too large')
    browser() 
  }
  
  if (!is.finite(val)) {
    
    cat("NA in loss_total_DC_db_storage")
    cat("Non-finite loss at params[1:5]:", round(params[1:5], 5), "\n")
    browser() 
    val <- 1e10  # Penalize non-finite values heavily
  }
  # cat('Value of loss: ',val, '\n')
  
  return(as.numeric(val))
  
}
```

```{r}
# test the DC approximated loss function
loss_total_DC_db_storage = function(params, params_old, K, data, krige_values, resids, 
                                    prop_estimation, lambda, smoothers, cum_smoothers, 
                                    bandwidth, threshold, epsilon)
  
trt_assigned_old = as.vector(K %*% eta + b)
trt_assigned = trt_assigned_old * 1.1

L_plus_total = total_loss_db_storage_plus(trt_assigned = trt_assigned, 
                                          col_idx = 1:n, 
                                          data = data, 
                                          krige_values = krige_values, 
                                          resids = resids, 
                                          prop_estimation = prop_estimation, 
                                          cum_smoothers = cum_smoothers,
                                          bandwidth = bandwidth,
                                          threshold = threshold,
                                          epsilon = epsilon)
  
grad_L_minus_total = total_loss_db_storage_minus_derivative(trt_assigned = trt_assigned_old, 
                                                            col_idx = 1:n, 
                                                            data = data, 
                                                            krige_values = krige_values, 
                                                            resids = resids, 
                                                            prop_estimation = prop_estimation, 
                                                            bandwidth = bandwidth,
                                                            threshold = threshold,
                                                            epsilon = epsilon)

L_minus_total = total_loss_db_storage_minus(trt_assigned = trt_assigned_old, 
                                            col_idx = 1:n, 
                                            data = data, 
                                            krige_values = krige_values, 
                                            resids = resids, 
                                            prop_estimation = prop_estimation, 
                                            bandwidth = bandwidth,
                                            threshold = threshold,
                                            epsilon = epsilon)

a = L_plus_total -  ( L_minus_total + grad_L_minus_total * (trt_assigned - trt_assigned_old ))

b = total_loss_db_storage_smooth(trt_assigned = trt_assigned, col_idx = 1:n, data = data, krige_values = krige_values, resids = resids, prop_estimation = prop_estimation, cum_smoothers = cum_smoothers, trt_range = trt_range, threshold = threshold,bandwidth = 0.07)
```

```{r}
plot(b,a)
```

###### DC loss function derivative

```{r}
loss_grad_total_DC_db_storage = function(params, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon) {
  n <- nrow(K)
  b <- params[1]
  eta <- params[2:(n+1)]

  b_old <- params_old[1]
  eta_old <- params_old[2:(n+1)]

  trt_assigned     <- as.vector(K %*% eta + b)
  trt_assigned_old <- as.vector(K %*% eta_old + b_old)

  # Compute gradient of concave part at current params
  grad_L_plus_total <- total_loss_db_storage_plus_derivative(trt_assigned = trt_assigned,
                                             col_idx = 1:n,
                                             data = data,
                                             krige_values = krige_values,
                                             resids = resids,
                                             prop_estimation = prop_estimation,
                                             smoothers = smoothers,
                                             bandwidth = bandwidth,
                                             threshold = threshold,
                                             epsilon = epsilon)
  
  # Compute gradient of concave part at old params
  grad_L_minus_total_old <- total_loss_db_storage_minus_derivative(trt_assigned = trt_assigned_old,
                                                               col_idx = 1:n,
                                                               data = data,
                                                               krige_values = krige_values,
                                                               resids = resids,
                                                               prop_estimation = prop_estimation,
                                                               bandwidth = bandwidth,
                                                               threshold = threshold,
                                                               epsilon = epsilon)

  grad_plus <-  apply(sapply(1:length(grad_L_plus_total), function(i) c(1,K[i,]) * grad_L_plus_total[i]),1,mean)
  
  grad_minus <-  apply(sapply(1:length(grad_L_minus_total_old), function(i) c(1, K[i,]) * grad_L_minus_total_old[i]),1,mean)
  
  grad_penalty <- c(0,lambda * as.vector(K %*% eta)) 
  
  jacobian_trt_assigned <- cbind(rep(1, nrow(K)), K)

  grad_vec <- as.vector(t(grad_L_plus_total -  grad_L_minus_total_old) %*% jacobian_trt_assigned)/n + grad_penalty
  
  # grad_vec <- grad_plus + grad_minus + grad_penalty
                      
  # if (max(abs(grad_vec))>100){
  #   cat('grad_loss_total_DC is way too large')
  #   browser() 
  # }
  
  # Check validity
  if (any(!is.finite(grad_vec))) {
    cat("NA in loss_grad_total_DC_db_storage")
    cat("Non-finite gradient at params[1:5]:", round(params[1:5], 5), "\n")
    browser()
    grad_vec[!is.finite(grad_vec)] <- 0
  }

  return(grad_vec)
}
```


```{r}
# test the above function

params_1 = params_old * 1.05
params_2 = params_old * 1.1


loss_total_DC_db_storage(params_2, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon)

loss_total_DC_db_storage(params_1, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon) +
sum(loss_grad_total_DC_db_storage(params_1, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon) * (params_2 - params_1))

```

```{r}
sum(loss_grad_total_DC_db_storage(params_1, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon) * (params_2 - params_1))
```

```{r}
sum(loss_grad_total_DC_db_storage(params_1, params_old, K, data, krige_values, resids,
                                         prop_estimation, lambda, smoothers, cum_smoothers, bandwidth,
                                         threshold, epsilon) * (params_2 - params_1))
```

###### Test for DB loss function

```{r}
# we first get data from direct
```


###### DC loss function implementation

```{r}
#' Construct the Kernel Matrix
#' @param params_init initial value of parameters
#' @param K kernel matrix used as predictive variables
#' @param gamma Bandwidth of the RBF kernel.
#' @return The n x n kernel matrix K.
run_true_dc_algorithm <- function(params, K, data, krige_values, resids, prop_estimation, lambda, cum_smoothers, smoothers,bandwidth, threshold, epsilon, max_iter = 10, tol = 1e-5){
    n <- nrow(K)
    params_old <- params

    cat("--- Starting True Difference of Convex Algorithm ---\n")

    for (k in 1:max_iter) {
        
        # Solve the convex subproblem
        cat(k,'-th iteration of DC algorithm','\n')
        sub_result <- optim(par = params_old, 
                            fn = loss_total_DC_db_storage, 
                            gr = loss_grad_total_DC_db_storage,
                            params_old = params_old,
                            K = K, 
                            lambda = lambda, 
                            data = data,
                            krige_values = krige_values, 
                            resids = resids, 
                            prop_estimation = prop_estimation, 
                            cum_smoothers = cum_smoothers,
                            smoothers = smoothers,
                            bandwidth = bandwidth,
                            threshold = threshold,
                            epsilon = epsilon,
                            method = "L-BFGS-B", 
                            control = list(maxit = 300, trace = 3))
        
        params <- sub_result$par

        # 3. Check for convergence
        change <- sqrt(sum((params_old - params)^2))
        cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
        
        if (change < tol) {
            cat("--- DC Algorithm Converged ---\n")
            params_old <- params
            break
        }
        
        params_old <- params
    }
    
    if (k == max_iter) {
        cat("--- DC Algorithm reached max iterations ---\n")
    }

    return(list(par = params_old, convergence = ifelse(k < max_iter, 0, 1), iter = k))
}
```

```{r}
dc_result = run_true_dc_algorithm(params = params, 
                                  K = K, 
                                  data = data, 
                                  krige_values = krige_values, 
                                  resids = resids, 
                                  prop_estimation = prop_estimation, 
                                  lambda = lambda, 
                                  cum_smoothers = cum_smoothers, 
                                  bandwidth = bandwidth, 
                                  threshold = threshold, 
                                  epsilon = epsilon, 
                                  max_iter = 4, 
                                  tol = 1e-5)
```
```{r}
plot(params[-1], dc_result$par[-1]) 
abline(0,1,col = 'red')
```


##### Result Assessment

```{r}
# get estimated policy
trt_est = cbind(rep(1, nrow(K)), K) %*% sub_result$par
# get estimated policy at initial params
trt_assigned_old = as.vector(K %*% eta_old + b_old)

print('RMSE')
cat('db parametric:', sqrt(mean(db_error^2)),'\n')
cat('db RKHS:', sqrt(mean((true_policy-trt_est)^2)),'\n')
cat('db RKHS initial:', sqrt(mean((true_policy-trt_assigned_old)^2)),'\n')

print('Bias')
cat('db parametric:', mean(db_error),'\n')
cat('db RKHS:', mean(true_policy-trt_est),'\n')
cat('db RKHS initial :', mean(true_policy-trt_assigned_old),'\n')

print('Variance')
cat('db parametric:', var(db_error),'\n')
cat('db RKHS:', var(true_policy-trt_est),'\n')
cat('db initial:', var(true_policy-trt_assigned_old),'\n')

```

```{r}
plot(true_policy,trt_est, xlab = 'true policy', ylab = 'db predicted policy',cex = 0.5, main = 'RKHS')
abline(0,1,col = 'red')
plot(true_policy, model.matrix(param_model_initial_db_storage_smooth) %*% parametric_model_db_storage_smooth$par, xlab = 'true policy', ylab = 'db predicted policy',cex = 0.5, main = 'parametric')
abline(0,1,col = 'red')
```
```{r}
# difference of initial params and estimated params
plot(params_old[-1], sub_result$par[-1],cex= 0.5)
abline(0,1,col='red')
```

```{r}

```


#### cross-validation

