---
title: "nitrate_regional_cross_validation"
author: "Xindi Lin"
date: "2025-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
library(ranger)
library(SuperLearner)
library(sf)
library(dplyr)
library(ggplot2)
library(WeightIt)
library(GpGp)
library(Metrics)
library(rnaturalearth)
library(ggplot2)
library(see)
library(yardstick)
library(xgboost)
library(glmnet)
library(kernlab)
library(glmnet)
library(numDeriv)
library(MASS)
library(e1071)
library(WeightIt)
library(caret)
```

```{r}
source('data_generation.R')
source('funcs_kriging.R')
source('loss_functions.R')
source('loss_functions_RKHS.R')
source('miscellaneous.R')
```


### preparation

#### load regional data and preprocess


```{r}
file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv"
## we will use the one with the proportion
# file_path <- "/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_latest_with_covar_landuseProp_median_well"
command <- paste("brctl download", shQuote(file_path))
system(command)
data <- read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv")
```

remove zero-inflated
```{r}
data <- data[data$concentration_plus_median!=0.5,]
```

choose area

```{r}
data$area = ""

indices_central <- data$County %in% c("Wood", "Portage", "Waupaca", "Juneau", "Adams", "Waushara", "Marquette", "Green Lake")
data[indices_central, "area"] = "Central"

indices_westCentral <- data$County %in% c("Buffalo", "Trempealeau", "La Crosse", "Monroe")
data[indices_westCentral, "area"] = "West Central"

indices_southCentral <- data$County %in% c("Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock")
data[indices_southCentral, "area"] = "South Central"

indices_eastCentral <- data$County %in% c("Outagamie", "Winnebago", "Door", "Fond du Lac", "Brown", "Calumet","Sheboygan","Manitowoc","Kewaunee")
data[indices_eastCentral, "area"] = "East Central"

indices_southWest <- data$County %in% c("Vernon", "Crawford", "Grant", "Richland", "Sauk", "Iowa","Lafayette")
data[indices_southWest, "area"] = "South West"

indices_southEast <- data$County %in% c("Washington", "Waukesha", "Walworth", "Ozaukee", "Milwaukee", "Racine","Kenosha")
data[indices_southEast, "area"] = "South East"
```

```{r}
data <- data[data$area == "East Central",]
```

add log transformation

```{r}
data$logWellDepth = log(data$WellDepth)
data$logconcentration_plus_median = log(data$concentration_plus_median)
```

```{r}
set.seed(1998)

data_full <- data
train_indices <- createDataPartition(data$logconcentration_plus_median, list = FALSE, p = 0.7) # from caret package
data <- data_full[train_indices,]
data_test <- data_full[-train_indices,]
```

we try new partition method such that the distribution of nitrate level are the same

```{r}
set.seed(1998)
# 2. First Split: Training (60%) vs. Temp (40%)
# Create indices for the training set. 'p = 0.6' means 60% of the data.
data_full <- data
train_indices <- createDataPartition(data_full$logconcentration_plus_median, p = 0.6, list = FALSE, times = 1)
# Separate the training data and the temporary data
data_train <- data_full[train_indices, ]
data_tmp  <- data_full[-train_indices, ]

# 3. Second Split: Validation (50% of Temp) vs. Test (50% of Temp)
# Split the temporary data in half to get validation and test sets.
# Note: We stratify on the 'y' variable of the temp_data.
validation_indices <- createDataPartition(data_tmp$logconcentration_plus_median, p = 0.5, list = FALSE, times = 1)

# Separate the validation and test data
data_validation <- data_tmp[validation_indices, ]
data_test <- data_tmp[-validation_indices, ]
```

```{r}
dim(data_train)
dim(data_validation)
dim(data_test)
```


#### Estimate nuisance parameter

##### OR RKHS

```{r}
OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","StaticLevel")
```

```{r}
design_matrix <- model.matrix(~ . , data = data_train[, OR_covariate_names])[,-1]
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])[,-1]
design_matrix_validation <- model.matrix(~ . , data = data_validation[, OR_covariate_names])[,-1]
```

```{r}
svm_auto <- svm(x = design_matrix, y = data_train$logconcentration_plus_median, type='nu-regression')
```

```{r}
gpfit_RKHS = GpGp::fit_model(svm_auto$residuals,
                        locs = data_train[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")

krige_values_RKHS <- leave_one_out_kriging(locs = data_train[,c("longitude", "latitude")], 
                                      y_obs = svm_auto$residuals, 
                                      gp_model = gpfit_RKHS$covfun_name,
                                      gp_params = gpfit_RKHS$covparms,
                                      order = c("coordinate")) 

krige_values_validation_RKHS <- GpGp::predictions(fit = gpfit_RKHS, 
                      locs_pred = data_validation[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_validation)))

krige_values_test_RKHS <- GpGp::predictions(fit = gpfit_RKHS, 
                      locs_pred = data_test[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_test)))
```

```{r}
cat('Test rMSE: with kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test) + krige_values_test_RKHS))
cat('\n')
cat('Test rMSE: No kriging: ')
cat(Metrics::rmse(log(data_test$concentration_plus_median), predict(svm_auto, newdata = design_matrix_test)))
cat('\n')

cat('Validation rMSE: with kriging: ')
cat(Metrics::rmse(log(data_validation$concentration_plus_median), predict(svm_auto, newdata = design_matrix_validation) + krige_values_validation_RKHS))
cat('\n')
cat('Validation rMSE: No kriging: ')
cat(Metrics::rmse(log(data_validation$concentration_plus_median), predict(svm_auto, newdata = design_matrix_validation)))
cat('\n')

cat('Training rMSE: with kriging: ')
cat(Metrics::rmse(log(data_train$concentration_plus_median), predict(svm_auto, newdata = design_matrix) + krige_values_RKHS))
cat('\n')
cat('Training rMSE: No kriging: ')
cat(Metrics::rmse(log(data_train$concentration_plus_median), predict(svm_auto, newdata = design_matrix)))
cat('\n')
```

##### Covariate Balancing Weight

```{r}
weights_cbps <- weightit(logWellDepth ~ StaticLevel+WellUse+crop_type_combine+drainagecl+precipitation+cafolog, 
                         data = data_train, method = "cbps")
```

#### Compute OR smoothers: RKHS

```{r}
boundary_tol <- 0
depth_range <- c(min(data_train$logWellDepth) - boundary_tol, max(data_train$logWellDepth) + boundary_tol)
depth_range
```

for training 

```{r}
# xgboost use the design matrix
cf_predictions_RKHS <- precompute_cf_predictions(data_obs = design_matrix, 
                              fit_outcome = svm_auto,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
cf_predictions_RKHS <- isotomic_correction(treatment_grid = cf_predictions_RKHS$treatment_values,
                                          pred_matrix_cf = cf_predictions_RKHS$pred_matrix)
```

```{r}
smoothers_RKHS <- make_prediction_smoothers(treatment_grid = cf_predictions_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_RKHS$pred_matrix,
                                            # smooth_method = "gam",
                                            smooth_method = "smooth.spline",
                                            # smooth_method = "splinefun",
                                            spar = NULL)
```

```{r}
cumint_smoothers_RKHS <- make_cumint_smoothers(treatment_grid = cf_predictions_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_RKHS$pred_matrix,
                                         # smooth_method = "splinefun",
                                         # smooth_method = "gam",
                                         smooth_method = "smooth.spline",
                                         spar = NULL)
```

for test 

```{r}
# xgboost use the design matrix
cf_predictions_test_RKHS <- precompute_cf_predictions(data_obs = design_matrix_test, 
                              fit_outcome = svm_auto,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
cf_predictions_test_RKHS <- isotomic_correction(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                          pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix)
```

```{r}
smoothers_test_RKHS <- make_prediction_smoothers(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix,
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers_test_RKHS <- make_cumint_smoothers(treatment_grid = cf_predictions_test_RKHS$treatment_values,
                                               pred_matrix_cf = cf_predictions_test_RKHS$pred_matrix,
                                               # smooth_method = "splinefun",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```

for validation

```{r}
# xgboost use the design matrix
cf_predictions_validation_RKHS <- precompute_cf_predictions(data_obs = design_matrix_validation, 
                              fit_outcome = svm_auto,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
cf_predictions_validation_RKHS <- isotomic_correction(treatment_grid = cf_predictions_validation_RKHS$treatment_values,
                                          pred_matrix_cf = cf_predictions_validation_RKHS$pred_matrix)
```

```{r}
smoothers_validation_RKHS <- make_prediction_smoothers(treatment_grid = cf_predictions_validation_RKHS$treatment_values,
                                            pred_matrix_cf = cf_predictions_validation_RKHS$pred_matrix,
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers_validation_RKHS <- make_cumint_smoothers(treatment_grid = cf_predictions_validation_RKHS$treatment_values,
                                               pred_matrix_cf = cf_predictions_validation_RKHS$pred_matrix,
                                               # smooth_method = "splinefun",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```

##### Visualizetion of OR Smoothers

```{r}
i <- 11
# when i = 20, it decrease first then increase, 
plot(seq(0,8,length.out = 100), sapply(seq(0,8,length.out = 100), function(knot) {smoothers_RKHS[[i]](knot)}), main = 'Smooth')
plot(cf_predictions_RKHS$treatment_values, sapply(cf_predictions_RKHS$treatment_values, function(knot) {smoothers_RKHS[[i]](knot)}), main = 'Smoothers RKHS')

plot(cf_predictions_RKHS$treatment_values, sapply(cf_predictions_RKHS$treatment_values, function(knot) {cumint_smoothers_RKHS[[i]](knot)}), main = 'cumint Smoothers RKHS')

plot(cf_predictions_RKHS$treatment_values,  cf_predictions_RKHS$pred_matrix[,i], main = 'Non-smooth original')
```

```{r}
mean(sapply(1:nrow(data), function(i){smoothers_RKHS[[i]](depth_range[1])>= log(5)}))
mean(sapply(1:nrow(data), function(i){smoothers_RKHS[[i]](depth_range[1]) + krige_values_RKHS[i] >= log(5)}))
```

```{r}
mean(data_train$logconcentration_plus_median >= log(5))
mean(data_train$logconcentration_plus_median >= log(2))
```

```{r}
mean(data_train$logconcentration_plus_median >= log(5))
mean(data_test$logconcentration_plus_median >= log(5))
```


#### Indirect Method

```{r}
threshold_val = log(5)
```

```{r}
indirect_policy_RKHS <- compute_indirect_policy(smoothers = smoothers_RKHS, trt_bounds = depth_range, threshold_val = threshold_val, krige_adjust = krige_values_RKHS, spatial = TRUE)

indirect_policy_validation_RKHS <- compute_indirect_policy(smoothers = smoothers_validation_RKHS, trt_bounds = depth_range, threshold_val = threshold_val, krige_adjust = krige_values_validation_RKHS, spatial = TRUE)

indirect_policy_test_RKHS <- compute_indirect_policy(smoothers = smoothers_test_RKHS, trt_bounds = depth_range, threshold_val = threshold_val, krige_adjust = krige_values_test_RKHS, spatial = TRUE)
```

```{r}
Metrics::accuracy(log(data_train$concentration_plus_median) > threshold_val, indirect_policy_RKHS > data_train$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_train$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_RKHS > data_train$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_test_RKHS > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_test_RKHS > data_test$logWellDepth))
```

```{r}
Metrics::accuracy(log(data_validation$concentration_plus_median) > threshold_val, indirect_policy_validation_RKHS > data_validation$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_validation$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_validation_RKHS > data_validation$logWellDepth))
```

#### initialization

##### one starting point

```{r}
# we use the following one for now
GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
kernel_design_matrix <- model.matrix(~ . , data = cbind(data_train[, GPS_covariate_names], U = krige_values_RKHS))[,-1]
kernel_design_matrix_test <- model.matrix(~ . , data = cbind(data_test[, GPS_covariate_names], U = krige_values_test_RKHS))[,-1]

train_means <- colMeans(kernel_design_matrix)
train_sds <- apply(kernel_design_matrix, 2, sd)

kernel_design_matrix <- scale(kernel_design_matrix)
kernel_design_matrix_test <- scale(kernel_design_matrix_test, center = train_means, scale = train_sds)
```

```{r}
resids <- data_train$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
scaled_resids <- resids * weights_cbps$weights
# plot(scaled_resids)
kernel_bw <- 1.06 * sd(data_train$logWellDepth) *nrow(data_train)^(-1/5)
```


```{r}
# initial_policy <- cbind(rep(1,nrow(data)), diag(nrow(data))) %*% params_initial_glmnet
initial_policy <- rep(mean(depth_range), nrow(data))

direct_individual <- c()
for(i in 1:nrow(data)){
  cat(i,'-th sample\n')
  direct_trt <- optim(# par = indirect_policy_RKHS[sub_indices[i]],
                      # par = mean(depth_range),
                      # par = depth_range[1],
                      par = initial_policy[i],
                      fn = compute_total_loss_smooth,
                      subject_idx = i,
                      T = data_train$logWellDepth,
                      krige_adjust = krige_values_RKHS,
                      outcome_resid = resids,
                      propensity_est = 1/weights_cbps$weights,
                      cumint_smoothers = cumint_smoothers_RKHS, 
                      trt_bounds = depth_range,
                      threshold_val = log(5), 
                      kernel_bw = kernel_bw, 
                      clip_epsilon = clip_epsilon ,
                      surrogate_type = "Gaussian",
                      loss_type = "db",
                      method = "L-BFGS-B", 
                      control = list(maxit = 100, trace = 1))
  # cat('optim trt: ', direct_trt$par,'\n\n')
  direct_individual <- c(direct_individual, direct_trt$par)
  # cat('indirect RKHS: ',indirect_policy_RKHS[sub_indices[i]],'\n\n')
  # cat('indirect DB: ', indirect_policy_DB[sub_indices[i]],'\n\n')
}

```

```{r}
Metrics::accuracy(log(data_train$concentration_plus_median) > threshold_val, direct_individual > data_train$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_train$concentration_plus_median) > threshold_val), estimate = as.factor( direct_individual > data_train$logWellDepth))
```

```{r}
plot(indirect_policy_RKHS,direct_individual)
```

##### multi-starting point

```{r}
clip_epsilon <- 1
resids <- data_train$logconcentration_plus_median - predict(svm_auto, newdata = design_matrix) - krige_values_RKHS
```


```{r}
kernel_bw <- 1.06 * sd(data_train$logWellDepth) *nrow(data_train)^(-1/5)
direct_individual_multi_initial <- c()

initial_value_set <- c(depth_range[1] - 1, mean(depth_range), depth_range[2] + 1)

for(i in 1:nrow(data_train)){
  cat(i,'-th sample\n')
  tmp_loss <- 1e4
  for(initial_point in initial_value_set){
    direct_trt <- optim(par = initial_point,
                        fn = compute_total_loss_smooth,
                        subject_idx = i,
                        T = data_train$logWellDepth,
                        krige_adjust = krige_values_RKHS,
                        outcome_resid = resids,
                        propensity_est = 1/weights_cbps$weights,
                        cumint_smoothers = cumint_smoothers_RKHS, 
                        trt_bounds = depth_range,
                        threshold_val = log(5), 
                        kernel_bw = kernel_bw, 
                        clip_epsilon = clip_epsilon ,
                        surrogate_type = "Gaussian",
                        loss_type = "db",
                        method = "L-BFGS-B", 
                        control = list(maxit = 300, trace = 1))
    if(direct_trt$value < tmp_loss){
      tmp_loss <- direct_trt$value
      tmp_direct_policy <- direct_trt$par
    }
  }
  direct_individual_multi_initial <- c(direct_individual_multi_initial, tmp_direct_policy)
}

```

```{r}
Metrics::accuracy(log(data_train$concentration_plus_median) > threshold_val, direct_individual_multi_initial > data_train$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_train$concentration_plus_median) > threshold_val), estimate = as.factor( direct_individual_multi_initial > data_train$logWellDepth))
```


### cross-validation with validation/test set

first we build grid for hyper-parameter, the metrics that we use are the accuracy and mcc in validation set

```{r}
# kernel_bw_grid <- c(0.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5),  1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 1.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 2 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5))
# 
# lambda_grid <- c(0,1,2,5)
# 
# m_grid <- c(-2,-1.5,-1,-0.5, 0)

kernel_bw_baseline <- 1.06 * sd(data_train$logWellDepth) *nrow(data_train)^(-1/5)
hyperparameters <- list(
  kernel_bw = c(0.5 * kernel_bw_baseline,  1 * kernel_bw_baseline, 1.5 * kernel_bw_baseline, 2 * kernel_bw_baseline),
  lambda = c(0,1,2,5),
  m = c(-2,-1.5,-1,-0.5, 0)
)

param_grid <- expand.grid(hyperparameters)

param_grid$acc_initial <- NA
param_grid$mcc_initial <- NA
param_grid$acc_validation <- NA
param_grid$mcc_validation <- NA
param_grid$acc_test <- NA
param_grid$mcc_test <- NA

# View the grid
print(param_grid)
```

```{r}
GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
# one type of kernel matrix using U as covariats
kernel_design_matrix <- model.matrix(~ . , data = cbind(data_train[, GPS_covariate_names], U = krige_values_RKHS))[,-1]
kernel_design_matrix_test <- model.matrix(~ . , data = cbind(data_test[, GPS_covariate_names], U = krige_values_test_RKHS))[,-1]
kernel_design_matrix_validation <- model.matrix(~ . , data = cbind(data_validation[, GPS_covariate_names], U = krige_values_validation_RKHS))[,-1]
```


```{r}
for(i in 1:nrow(param_grid)){
  m <- param_grid$m[i]
  lambda <- param_grid$lambda[i]
  kernel_bw <- param_grid$kernel_bw[i]
  
  result <- real_data_cross_validation(m = m, 
                             lambda = lambda, 
                             kernel_bw = kernel_bw, 
                             kernel_design_matrix = kernel_design_matrix, 
                             kernel_design_matrix_test = kernel_design_matrix_test,
                             kernel_design_matrix_validation = kernel_design_matrix_validation)
      
  param_grid$acc_initial[i] <- result$acc_initial
  param_grid$mcc_initial[i] <- result$mcc_initial
  param_grid$acc_validation[i] <- result$acc_validation
  param_grid$mcc_validation[i] <- result$mcc_validation
  param_grid$acc_test[i] <- result$acc_test
  param_grid$mcc_test[i] <- result$mcc_test
  
  # Optional: Print progress to the console
  cat(sprintf("Finished run %d/%d | m=%.2f, lambda=%.2f, bw=%.2f | Val MCC: %.4f\n | Test MCC: %.4f\n", 
              i, nrow(param_grid), m, lambda, kernel_bw, result$mcc_validation, result$mcc_test))    
     
}
```



### Cross-Validation

#### Cross-Validation for non-DC algorithm

Cross-validation for ##### Optimize Sum of DB loss: non-DC algorithm, non-identity-kernel, whole dataset##### on policy_learning_nitrate_regional_test.rmd

##### kernel bandwidth

```{r}
for(m in c(-1.7,-1.6,-1.5,-1.4,-1.3)){
  cat("m is: ",m,'\n')
  gamma <- 2^(m) * median(rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)

  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  
  
  print_acc_mcc(T = data_test$logWellDepth, 
                Y = data_test$logconcentration_plus_median, 
                policy = k_matrix_test %*% kernel_optim_trt_DB$par, 
                threshold_val = log(5))

}
```

We see that -1.5 is the best

##### RKHS coefficient penalty

```{r}
for(lambda in c(0, 0.25, 0.5, 1, 2)){
  cat("lambda is: ",lambda,'\n')
  m <- -1.5
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)

  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  
  kernel_optim_trt_DB <- optim(par = params_initial_glmnet, 
      fn = compute_total_loss_smooth_RKHS,
      gr = d_compute_total_loss_smooth_RKHS,
      K = k_matrix,
      # K = cbind(rep(1,nrow(data)), diag(nrow(data))),
      # K = diag(nrow(data)),
      T = data_train$logWellDepth,
      krige_adjust = krige_values_RKHS,
      outcome_resid = resids,
      propensity_est = 1/weights_cbps$weights,
      lambda = lambda, 
      smoothers = smoothers_RKHS, 
      cumint_smoothers = cumint_smoothers_RKHS, 
      trt_bounds = depth_range,
      threshold_val = log(5), 
      kernel_bw = kernel_bw, 
      clip_epsilon = clip_epsilon,
      surrogate_type = "Gaussian",
      loss_type = "db",
      method = "L-BFGS-B",
      control = list(maxit = 1000, trace = 3)) 
  
  print_acc_mcc(T = data_test$logWellDepth, 
              Y = data_test$logconcentration_plus_median, 
              policy = k_matrix_test %*% kernel_optim_trt_DB$par, 
              threshold_val = log(5))
  

}
```

lambda similar to 0.25, test acc is 0.572922 

##### sorrogate kernel bandwidth

```{r}

kernel_bw <- 

for(lambda in c(0, 0.25, 0.5, 1, 2)){
  cat("lambda is: ",lambda,'\n')
  m <- -1.5
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)

  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  
  kernel_optim_trt_DB <- optim(par = params_initial_glmnet, 
      fn = compute_total_loss_smooth_RKHS,
      gr = d_compute_total_loss_smooth_RKHS,
      K = k_matrix,
      # K = cbind(rep(1,nrow(data)), diag(nrow(data))),
      # K = diag(nrow(data)),
      T = data_train$logWellDepth,
      krige_adjust = krige_values_RKHS,
      outcome_resid = resids,
      propensity_est = 1/weights_cbps$weights,
      lambda = lambda, 
      smoothers = smoothers_RKHS, 
      cumint_smoothers = cumint_smoothers_RKHS, 
      trt_bounds = depth_range,
      threshold_val = log(5), 
      kernel_bw = kernel_bw, 
      clip_epsilon = clip_epsilon,
      surrogate_type = "Gaussian",
      loss_type = "db",
      method = "L-BFGS-B",
      control = list(maxit = 1000, trace = 3)) 
  
  print_acc_mcc(T = data_test$logWellDepth, 
              Y = data_test$logconcentration_plus_median, 
              policy = k_matrix_test %*% kernel_optim_trt_DB$par, 
              threshold_val = log(5))
  

}
```

##### (Optimal) for non-DC algorithm

```{r}
# cat("lambda is: ",lambda,'\n')
m <- -1.5
lambda <- 0.25
gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))
k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)

k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)

initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual, alpha = 0, lambda = 0)
params_initial_glmnet <- coef(initial_glmnet)
params_initial_glmnet <- as.numeric(params_initial_glmnet)

kernel_optim_trt_DB <- optim(par = params_initial_glmnet, 
    fn = compute_total_loss_smooth_RKHS,
    gr = d_compute_total_loss_smooth_RKHS,
    K = k_matrix,
    # K = cbind(rep(1,nrow(data)), diag(nrow(data))),
    # K = diag(nrow(data)),
    T = data_train$logWellDepth,
    krige_adjust = krige_values_RKHS,
    outcome_resid = resids,
    propensity_est = 1/weights_cbps$weights,
    lambda = lambda, 
    smoothers = smoothers_RKHS, 
    cumint_smoothers = cumint_smoothers_RKHS, 
    trt_bounds = depth_range,
    threshold_val = log(5), 
    kernel_bw = kernel_bw, 
    clip_epsilon = clip_epsilon,
    surrogate_type = "Gaussian",
    loss_type = "db",
    method = "L-BFGS-B",
    control = list(maxit = 1000, trace = 3)) 

print_acc_mcc(T = data_test$logWellDepth, 
            Y = data_test$logconcentration_plus_median, 
            policy = k_matrix_test %*% kernel_optim_trt_DB$par, 
            threshold_val = log(5))
  
```

The loss function has to be bounded. Smoothness of L_minus

Stochastic gradient descent.
Many non-convex algorithm: different advantages and drawbacks.

Maybe try CXVR. 


#### Cross-Validation for DC algorithm

##### build covariate matrix

using kriging values as covariate

```{r}
# we use the following one for now
GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
# one type of kernel matrix using U as covariats
kernel_design_matrix <- model.matrix(~ . , data = cbind(data_train[, GPS_covariate_names], U = krige_values_RKHS))[,-1]
kernel_design_matrix_test <- model.matrix(~ . , data = cbind(data_test[, GPS_covariate_names], U = krige_values_test_RKHS))[,-1]

train_means <- colMeans(kernel_design_matrix)
train_sds <- apply(kernel_design_matrix, 2, sd)

kernel_design_matrix <- scale(kernel_design_matrix)
kernel_design_matrix_test <- scale(kernel_design_matrix_test, center = train_means, scale = train_sds)
```

```{r}
# the other type of kernel matrix using coordinates
GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","latitude","longitude")
kernel_design_matrix <- model.matrix(~ . , data = data_train[, GPS_covariate_names])[,-1]
kernel_design_matrix_test <- model.matrix(~ . , data = data_test[, GPS_covariate_names])[,-1]

train_means <- colMeans(kernel_design_matrix)
train_sds <- apply(kernel_design_matrix, 2, sd)

kernel_design_matrix <- scale(kernel_design_matrix)
kernel_design_matrix_test <- scale(kernel_design_matrix_test, center = train_means, scale = train_sds)
```


##### Cross-Validation for RKHS kernel bandwidth in initialization

```{r}
for(m in c(-2,-1.5,-1,-0.5)){
  cat('value of m is ',m,'\n')
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual_multi_initial, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  print_acc_mcc(T = data_test$logWellDepth, 
            Y = data_test$logconcentration_plus_median, 
            policy = k_matrix_test %*% params_initial_glmnet,
            # policy = k_matrix_test %*% params_old, 
            threshold_val = log(5))
}

```

m equal to -0.5 seems to be better, we will further try it with a finer grid near  m=-0.5


```{r}
for(m in c(-0.7,-0.5,-0.3,-0.1)){
  cat('value of m is ',m,'\n')
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual_multi_initial, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  print_acc_mcc(T = data_test$logWellDepth, 
            Y = data_test$logconcentration_plus_median, 
            policy = k_matrix_test %*% params_initial_glmnet,
            # policy = k_matrix_test %*% params_old, 
            threshold_val = log(5))
}
```

It turns out -0.5 is the best


##### Cross-Validation for surrogate kernel bandwidth in initialization

```{r}
# This function contains the work for a single iteration of your original loop
find_best_policy <- function(i, initial_value_set, kernel_bw) {
  
  # Use Inf for a more robust starting loss value
  tmp_loss <- Inf
  tmp_direct_policy <- NA # Default return value in case of errors
  
  # Inner loop to find the best initial point for the optimization
  for (initial_point in initial_value_set) {
    
    # Use try() to catch potential errors during optimization
    
      direct_trt_result <- optim(par = initial_point,
            fn = compute_total_loss_smooth,
            gr = d_compute_total_loss_smooth,
            subject_idx = i,
            # Pass other necessary arguments using ...
            T = data_train$logWellDepth,
            krige_adjust = krige_values_RKHS,
            outcome_resid = resids,
            propensity_est = 1/weights_cbps$weights,
            smoothers = smoothers_RKHS, 
            cumint_smoothers = cumint_smoothers_RKHS, 
            trt_bounds = depth_range,
            threshold_val = log(5), 
            kernel_bw = kernel_bw, 
            clip_epsilon = clip_epsilon ,
            surrogate_type = "Gaussian",
            loss_type = "db",
            method = "L-BFGS-B",
            # Set trace = 0 to prevent garbled console output from parallel workers
            control = list(maxit = 300, trace = 0)) 
    
    
    # Check if optim succeeded and if the new loss is an improvement
    if (direct_trt_result$value < tmp_loss) {
      tmp_loss <- direct_trt_result$value
      tmp_direct_policy <- direct_trt_result$par
    }
  }
  
  return(tmp_direct_policy)
}
```

```{r}
kernel_bw_lst <- c(0.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5),  1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 1.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 2 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5))

for(kernel_bw_try in kernel_bw_lst){
  cat('The kernel bandwith is ',kernel_bw_try, '\n')
  direct_individual_multi_initial <- mcmapply(FUN = find_best_policy, i = 1:nrow(data), MoreArgs = list(
initial_value_set = initial_value_set, kernel_bw = kernel_bw_try), SIMPLIFY = TRUE)
  mcc_metric <- yardstick::mcc_vec(truth = as.factor(log(data_train$concentration_plus_median) > threshold_val), estimate = as.factor( direct_individual_multi_initial > data_train$logWellDepth))
  cat('mcc is ',mcc_metric,'\n')
}
```
It seems that a larger bandwidth is better

```{r}
kernel_bw_lst <- c(2 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 2.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 3 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5), 3.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5))

for(kernel_bw_try in kernel_bw_lst){
  cat('The kernel bandwith is ',kernel_bw_try, '\n')
  direct_individual_multi_initial <- mcmapply(FUN = find_best_policy, i = 1:nrow(data), MoreArgs = list(
initial_value_set = initial_value_set, kernel_bw = kernel_bw_try), SIMPLIFY = TRUE)
  mcc_metric <- yardstick::mcc_vec(truth = as.factor(log(data_train$concentration_plus_median) > threshold_val), estimate = as.factor( direct_individual_multi_initial > data_train$logWellDepth))
  cat('mcc is ',mcc_metric,'\n')
}
```
It seems that 2.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5) is the best

```{r}
kernel_bw_try <- 2.5 * 1.06 * sd(data_train$logWellDepth) *nrow(data)^(-1/5)
direct_individual_multi_initial <- mcmapply(FUN = find_best_policy, i = 1:nrow(data), MoreArgs = list(
initial_value_set = initial_value_set, kernel_bw = kernel_bw_try), SIMPLIFY = TRUE)
```


```{r}
for(m in c(-0.7,-0.5,-0.3,-0.1)){
  cat('value of m is ',m,'\n')
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
  
  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
  initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual_multi_initial, alpha = 0, lambda = 0)
  params_initial_glmnet <- coef(initial_glmnet)
  params_initial_glmnet <- as.numeric(params_initial_glmnet)
  print_acc_mcc(T = data_test$logWellDepth, 
            Y = data_test$logconcentration_plus_median, 
            policy = k_matrix_test %*% params_initial_glmnet,
            # policy = k_matrix_test %*% params_old, 
            threshold_val = log(5))
}
```

The chosen surrogate bandwith does not increase the performance, what we do does not make sense, I think we need a valide cross-validation.

##### Cross-Validation with ...


```{r}
initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual_multi_initial, alpha = 0, lambda = 0)
params_initial_glmnet <- coef(initial_glmnet)
params_initial_glmnet <- as.numeric(params_initial_glmnet)
```

```{r}
print_acc_mcc(T = data_test$logWellDepth, 
            Y = data_test$logconcentration_plus_median, 
            policy = k_matrix_test %*% params_initial_glmnet,
            # policy = k_matrix_test %*% params_old, 
            threshold_val = log(5))
```


```{r}
for(m in c(-2,,-1.5,-1,-0.5)){
  gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
  rbf <- rbfdot(sigma = 1/(gamma^2))
  k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
  k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)

  k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
  k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
  
}
```


##### optimal

```{r}
m <- -0.5
gamma <- 2^(m) * median(fields::rdist(kernel_design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))
k_matrix <- kernelMatrix(rbf, kernel_design_matrix)
k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)

k_matrix_test <- kernelMatrix(rbf, kernel_design_matrix_test, kernel_design_matrix)
k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
initial_glmnet <- glmnet(x = k_matrix[,-1], y = direct_individual_multi_initial, alpha = 0, lambda = 0)
params_initial_glmnet <- coef(initial_glmnet)
params_initial_glmnet <- as.numeric(params_initial_glmnet)
print_acc_mcc(T = data_test$logWellDepth, 
          Y = data_test$logconcentration_plus_median, 
          policy = k_matrix_test %*% params_initial_glmnet,
          # policy = k_matrix_test %*% params_old, 
          threshold_val = log(5))
```

```{r}
hist(k_matrix_test %*% params_initial_glmnet)
```


```{r}
compute_total_loss_smooth_RKHS(params =  c(0, direct_individual_multi_initial),
                                K = cbind(rep(0, nrow(data)), diag(nrow(data))),
                                T = data_train$logWellDepth,
                                krige_adjust = krige_values_RKHS,
                                outcome_resid = resids,
                                propensity_est = 1/weights_cbps$weights,
                                lambda = 0, 
                                cumint_smoothers = cumint_smoothers_RKHS,
                                smoothers = smoothers_RKHS,
                                trt_bounds = depth_range,
                                threshold_val = log(5), 
                                kernel_bw= kernel_bw, 
                                clip_epsilon = clip_epsilon * 20,
                                surrogate_type = "Gaussian",
                                loss_type = "db")
```


```{r}
compute_total_loss_smooth_RKHS(params =  params_initial_glmnet,
                                K = k_matrix,
                                T = data_train$logWellDepth,
                                krige_adjust = krige_values_RKHS,
                                outcome_resid = resids,
                                propensity_est = 1/weights_cbps$weights,
                                lambda = 0, 
                                cumint_smoothers = cumint_smoothers_RKHS,
                                smoothers = smoothers_RKHS,
                                trt_bounds = depth_range,
                                threshold_val = log(5), 
                                kernel_bw= kernel_bw, 
                                clip_epsilon = clip_epsilon * 20,
                                surrogate_type = "Gaussian",
                                loss_type = "db")
```

```{r}
compute_total_loss_smooth_dc_approx_RKHS(params =  params_initial_glmnet,
                                         params_old = params_initial_glmnet,
                                K = k_matrix,
                                T = data_train$logWellDepth,
                                krige_adjust = krige_values_RKHS,
                                outcome_resid = resids,
                                propensity_est = 1/weights_cbps$weights,
                                lambda = 0, 
                                cumint_smoothers = cumint_smoothers_RKHS,
                                smoothers = smoothers_RKHS,
                                trt_bounds = depth_range,
                                threshold_val = log(5), 
                                kernel_bw= kernel_bw, 
                                clip_epsilon = clip_epsilon * 20,
                                clip_epsilon_bar = 20 * clip_epsilon,
                                surrogate_type = "Gaussian",
                                loss_type = "db")
```


```{r}
max_iter <- 50
tol <- 1e-3

params_old <- params_initial_glmnet
for(k in 1:max_iter) {
  sub_result <- optim(par = params_old, 
                      fn = compute_total_loss_smooth_dc_approx_RKHS, # for multiple of it, we have to use sum
                      gr = d_compute_total_loss_smooth_dc_approx_RKHS, # we try the case with/without gr, we found
                      params_old = params_old, 
                      K = k_matrix,
                      T = data_train$logWellDepth,
                      krige_adjust = krige_values_RKHS,
                      outcome_resid = resids,
                      propensity_est = 1/weights_cbps$weights,
                      lambda = 2, 
                      cumint_smoothers = cumint_smoothers_RKHS, 
                      smoothers = smoothers_RKHS,
                      trt_bounds = depth_range,
                      threshold_val = log(5), 
                      kernel_bw= kernel_bw, 
                      # clip_epsilon = 1,
                      # clip_epsilon_bar = 1000,
                      clip_epsilon = clip_epsilon * 20,
                      clip_epsilon_bar = 20 * clip_epsilon,
                      surrogate_type = "Gaussian",
                      loss_type = "db",
                      method = "L-BFGS-B" , 
                      control = list(maxit = 50, trace = 1))
  
  params <- sub_result$par
  
  
  # 3. Check for convergence
  change <- sqrt(sum((params_old - params)^2))
  cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
  
  if (change < tol) {
    cat("--- DC Algorithm Converged ---\n")
    params_old <- params
    break
  }
  params_old <- params
  
  print_acc_mcc(T = data_test$logWellDepth, 
          Y = data_test$logconcentration_plus_median, 
          # policy = k_matrix_test %*% params_initial_glmnet,
          policy = k_matrix_test %*% params_old,
          threshold_val = log(5))
}
```


I think we need to do official cross-validation of the parameter



