---
title: "policy_learning_nitrate"
author: "Xindi Lin"
date: "2025-08-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# library(ranger)
# library(CausalGPS)
library(SuperLearner)
library(sf)
# library(sl3) # The modern Super Learner framework
library(dplyr)
library(ggplot2)
# library(WeightIt)
library(GpGp)
library(Metrics)
library(rnaturalearth)
library(ggplot2)
library(see)
library(yardstick)
library(xgboost)
library(glmnet)
```

```{r}
source('data_generation.R')
source('funcs_kriging.R')
source('loss_functions.R')
source('loss_functions_RKHS.R')
```

#### load & filtering the nitrate data

```{r}
data = read.csv("/Users/xindilin/Desktop/2024 summer/groundwater_pesticide/data/data_Nitrate_with_covar_median_well.csv")
```

```{r}
dim(data)
sum(is.na(data))
```

```{r}
head(data)
```

##### Well Type

```{r}
table(data$WellUse)
```
We might only keep the private wells

```{r}
data <- data[data$WellUse == "PRIVATE POTABLE",]
```

##### Localize Area

```{r}
unique(data$County)
```

"Central": "Wood", "Portage", "Waupaca", "JUNEAU", "Adams", "Waushara", "Marquette", "Green lake"
"West Central": "Buffalo", "Trempealeau", "La Crosse", "Monraoe"
"South Central": "Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock"
"East Central": 

```{r}
data$area = ""
```

```{r}
indices_central <- data$County %in% c("Wood", "Portage", "Waupaca", "Juneau", "Adams", "Waushara", "Marquette", "Green Lake")
data[indices_central, "area"] = "Central"

indices_westCentral <- data$County %in% c("Buffalo", "Trempealeau", "La Crosse", "Monroe")
data[indices_westCentral, "area"] = "West Central"

indices_southCentral <- data$County %in% c("Columbia", "Dodge", "Dane", "Jefferson", "Green", "Rock")
data[indices_southCentral, "area"] = "South Central"
```

```{r}
unique(data[data$area == "Central", "County"])
unique(data[data$area == "West Central", "County"])
unique(data[data$area == "South Central", "County"])
```

```{r}
dim(data[data$area == "Central",])
dim(data[data$area == "West Central",])
dim(data[data$area == "South Central",])
```



##### keep the latest record

by WellID

```{r}
latest_records <- data %>%
  # Make sure the date column is a true Date type for correct sorting
  mutate(SampleDate = as.Date(SampleDate)) %>%
  # Group the data by the well ID
  group_by(WellID) %>%
  # Within each group, keep only the row with the maximum date
  slice_max(order_by = SampleDate, n = 1) %>%
  # Optional: remove the grouping
  ungroup()
```

```{r}
dim(latest_records)
```

keep the latest record by "welldepth","latitude","longitude"

```{r}
latest_records_depth <- latest_records %>%
  # Ensure the date column is a true Date type
  mutate(SampleDate = as.Date(SampleDate)) %>%
  # Group by all four identifying columns
  group_by(WellDepth, latitude, longitude) %>%
  # Keep the row with the latest date within each group
  slice_max(order_by = SampleDate, n = 1) %>%
  ungroup()

latest_records_depth <- latest_records %>%
  # Ensure the date column is a true Date type
  # mutate(SampleDate = as.Date(SampleDate)) %>%
  # Group by all four identifying columns
  group_by(WellDepth, latitude, longitude) %>%
  # Keep the row with the latest date within each group
  slice_max(order_by = SampleYear, n = 1) %>%
  ungroup()
```

```{r}
dim(latest_records_depth)
```

```{r}
data <- latest_records_depth
```

```{r}
summary(data$WellDepth)
```

##### Bin depth

Bin depth and get the median, I think we might ignore this step.
Because this step actually makes it multi-categorical, we might consider later

```{r}
# 2. Bin depths and calculate the median for each bin
binned_data <- latest_records_depth %>%
  # Create a new column 'depth_bin' representing the start of each 5-unit interval
  # e.g., 101, 102, 104 all become 100. 115 becomes 115.
  mutate(depth_bin = floor(WellDepth / 5) * 5) %>%
  # Now group by that new depth bin as well
  group_by(depth_bin,latitude, longitude) %>% # Group by the location first
  # Calculate the median measurement for each group
  # Also, let's count the records and find the median depth in that bin
  mutate(median_measurement_in_bin = median(concentration_plus_median)) 
```

```{r}
dim(binned_data)
```

```{r}
binned_data <-  binned_data %>%
  group_by(depth_bin,latitude, longitude) %>%
  slice_head(n=1)
```

```{r}
dim(binned_data)
```

```{r}
data <- binned_data
```

```{r}
data$WellDepth <- data$depth_bin
data$concentration_plus_median <- data$median_measurement_in_bin
```


```{r}
dim(data)
```

##### filter data close to zero/threshold

we may filter out the nitrate level is close to zero

```{r}
dim(data[data$concentration_plus_median>=1,])
data <- data[data$concentration_plus_median>=1,]
```

There are too many of them, so I sample a subset

```{r}
# There are too many of them, so I sample a subset
data = data[sample(nrow(data),15000),]
```

##### we may filter out the nitrate level around the threshold

```{r}
summary(log(data$concentration_plus_median) - log(10))
```

```{r}
sum(log(data$concentration_plus_median) - log(10) >= 0 & log(data$concentration_plus_median) - log(10) <= 3)
```

```{r}
sum(log(data$concentration_plus_median) - log(10) <= 0 & log(data$concentration_plus_median) - log(10) >= -1)
```

```{r}
filtering_index <- log(data$concentration_plus_median) - log(10) <= 2 & log(data$concentration_plus_median) - log(10) >= -1

data <- data[filtering_index,]
```

##### filter for a reasonable well depth range

```{r}
summary(data$WellDepth)
```

```{r}
data <- data[data$WellDepth!=1599.0 ,]
```


##### filter data before 2020

```{r}
data <- data[data$SampleYear>=2020,]
```

#### Split to Training and Test Data

```{r}
set.seed(1998)

data_full <- data
train_indices <- sample(nrow(data_full), nrow(data_full) * 0.6)
data <- data_full[train_indices,]
data_test <- data_full[-train_indices,]
```

#### Visualization

##### Distribution of Contaminants and Depth

```{r}
plot(density(log(data$WellDepth)))
# so we can roughly assume the log of wellDepth, i.e., the treatment, is normally distributed.
```


```{r}
plot(density(log(data$concentration_plus_median)))
```

```{r}
plot(density(log(data$concentration_plus_median[data$concentration_plus_median!=0.5])))
```

##### How does it change with depth for individual locations

```{r}
# 1. Find the occurrence of each unique combination
occurrence_counts <- data %>%
  count(longitude, latitude, sort = TRUE)

# 2. Get just the most frequent combination
most_frequent <- occurrence_counts %>%
  # slice_head() is a reliable way to get the top row
  slice_head(n = 1)
```

```{r}
i = 3
original_records <- inner_join(data, occurrence_counts[i,], by = c("longitude", "latitude"))
```

```{r}
plot(original_records$WellDepth, original_records$concentration_plus_median)
```
Now I may want to take temporal 


#### Estimate nuisance parameter

```{r}
data$logWellDepth = log(data$WellDepth)
data_test$logWellDepth = log(data_test$WellDepth)
data$logconcentration_plus_median = log(data$concentration_plus_median)
data_test$logconcentration_plus_median = log(data_test$concentration_plus_median)
```

##### Estimate Outcome Regression

```{r}
# OR_covariate_names <- c("logWellDepth","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","longitude","latitude","SampleYear")

OR_covariate_names <- c("logWellDepth","precipitation","cafolog","crop_type_combine","drainagecl")
```

Adding latitude, longitude does not change the prediction result a lot.

```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])
```


###### Using Random Forest

```{r}
OR_mean_model <- SuperLearner(
  Y = log(data$concentration_plus_median),
  X = data[, OR_covariate_names],
  family = gaussian(),
  SL.library = c("SL.ranger")
)
```

```{r}
plot(log(data$concentration_plus_median), OR_mean_model$library.predict, cex = 0.1)
abline(0,1,col='red')
```
```{r}
test_OR_pred <- predict(OR_mean_model, newdata = data_test)$pred
```

```{r}
plot(log(data_test$concentration_plus_median), test_OR_pred)
abline(0,1,col='red')
```

###### Using GAM

I think a treatment free component model and a blip model seems to match to the mechanism of the nitrate

$$E(Y|A,X,U) = f(X_1)+f(A,X_2)+U$$
Here, $X_1$ is the source of the nitrate, including 

```{r}
OR_covariate_names <- c("logWellDepth","StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
```

```{r}
OR_base_covariates <- c("crop_type_combine", "cafolog")
OR_interaction_covariates <- c("logWellDepth", "drainagecl","precipitation")
```

```{r}
data$crop_type_combine <- as.factor(data$crop_type_combine)
data$drainagecl <- as.factor(data$drainagecl)
data$logconcentration_plus_median <- log(data$concentration_plus_median)
```

```{r}
OR_mean_model_gam <- mgcv::gam(logconcentration_plus_median ~ s(cafolog, by = crop_type_combine) + te(logWellDepth, precipitation, by = drainagecl), data = data)
```

```{r}
plot(data$logconcentration_plus_median, OR_mean_model_gam$fitted.values)
abline(0,1,col='red')
```

###### Using RKHS

```{r}
library(kernlab)
library(glmnet)
```

```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])
# we need to scale the continuous variables
mean_logWellDepth <- mean(design_matrix[, 2])
sd_logWellDepth <- sd(design_matrix[, 2])

mean_precipitation <- mean(design_matrix[, 3])
sd_precipitation <- sd(design_matrix[, 3])

mean_cafolog <- mean(design_matrix[, 4])
sd_cafolog <- sd(design_matrix[, 4])
```

```{r}
design_matrix[, 2] <- (design_matrix[, 2] - mean_logWellDepth)/sd_logWellDepth
design_matrix[, 3] <- (design_matrix[, 3] - mean_precipitation)/sd_precipitation
design_matrix[, 4] <- (design_matrix[, 4] - mean_cafolog)/sd_cafolog
# standardize for test data
design_matrix_test[, 2] <- (design_matrix_test[, 2] - mean_logWellDepth)/sd_logWellDepth
design_matrix_test[, 3] <- (design_matrix_test[, 3] - mean_precipitation)/sd_precipitation
design_matrix_test[, 4] <- (design_matrix_test[, 4] - mean_cafolog)/sd_cafolog
```

```{r}
gamma <- median(rdist(design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))
k_matrix <- kernelMatrix(rbf, design_matrix)
# build kernel matrix for test data
k_matrix_test <- kernelMatrix(rbf, design_matrix_test,design_matrix)
```

```{r}
OR_mean_model_RKHS <- glmnet(x = k_matrix, y = data$logconcentration_plus_median, alpha = 0)

OR_mean_model_RKHS.cv <- cv.glmnet(x = k_matrix, y = data$logconcentration_plus_median, alpha = 0)
```

```{r}
OR_pred_RKHS <- predict(OR_mean_model_RKHS, newx = k_matrix, s=OR_mean_model_RKHS.cv$lambda.min)

OR_pred_RKHS_test <- predict(OR_mean_model_RKHS, newx = k_matrix_test, s=OR_mean_model_RKHS.cv$lambda.min)
```

I actually don't know why it fails

```{r}
plot(data$logconcentration_plus_median, OR_pred_RKHS, cex = 0.1)
abline(0,1,col='red')
plot(data_test$logconcentration_plus_median, OR_pred_RKHS_test, cex = 0.1)
abline(0,1,col='red')
```





###### Using XGBoost



The performance of XGboost is similar to random forest, but we may add monotone constraints, or use linear model rather than tree

We draw the prediction curve of indivisual predictions

```{r}
i = 10
treatment_values <- seq(min(design_matrix[,"logWellDepth"]), max(design_matrix[,"logWellDepth"]), by = 0.01)
data_expanded <- design_matrix[rep(i, each = length(treatment_values)), ]
data_expanded[,"logWellDepth"] <- rep(treatment_values, times = 1)
```

```{r}
plot(treatment_values, predict(OR_mean_model_xgboost, data_expanded))
```

##### Estimate Residual using Gaussian Process

```{r}
data$OR_residuals = log(data$concentration_plus_median) - OR_mean_model$Z[,1]
```

$$Y_{ij} = f(A_{ij},X_i)+U_i+\epsilon_{ij}, i \text{ is the block}, j \text{ every well in the block}$$
If most section only has one well, we should create representative well for every location.

Add noise to the locations (latitude, longitude). 

Add a random noise to the latitude, longitude to avoid overlap points. A degree of latitude is about 69 miles.
A section is a 1 mile by 1 mile block. So we add a uniform disturbation -1/100,1/100

```{r}
data$longitude <- data$longitude + runif(nrow(data), -0.01,0.01)
data$latitude <- data$latitude+ runif(nrow(data), -0.01,0.01)
```

If we add noise, we might skip this step of getting median

```{r}
data_plss_residual = data %>%
  group_by(longitude, latitude)%>%
  summarise(
    median_value = median(OR_residuals, na.rm = TRUE,.groups = 'keep')
  )
```

```{r}
data_num_of_wells = data %>%
  group_by(longitude, latitude)%>%
  summarise(
    count = n()
  )
```

```{r}
hist(data_num_of_wells$count)
mean(data_num_of_wells$count == 1)
```

```{r}
nrow(data_plss_residual)
nrow(data)
```

```{r}
y = log(data$concentration_plus)
locs = data[,c("longitude", "latitude")]
# y = residuals(gam.model_interact)
```

```{r}
gpfit = GpGp::fit_model(data_plss_residual$median_value,
                        locs = data_plss_residual[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")
```

```{r}
# get prediction
data$Gp_pred = GpGp::predictions(fit = gpfit, 
                  locs_pred = data[,c("longitude", "latitude")], 
                  X_pred = rep(1,nrow(data)))
```

```{r}

```

Become very slow when there are lots of points

```{r}
gpfit = GpGp::fit_model(data$logconcentration_plus_median - loocv_model$pred,
                        locs = data[,c("longitude", "latitude")],
                        covfun_name = "matern_sphere")
```

##### Estimate Residual using Thin Plate Splines

```{r}
# this expression has to use mgcv but not gam package, gam in 'mgcv' is different from gam in 'gam' package
spline_smooth <- mgcv::gam(OR_residuals~s(longitude,latitude,k=300,bs='tp',fx=TRUE),data=data,method="REML") #,method="REML"
```

```{r}
plot(data$OR_residuals, spline_smooth$fitted.values)
```

```{r}
spline_test_values <- predict(spline_smooth, newdata = data_test)
```

##### OR prediction on test set

```{r}
krige_values <- GpGp::predictions(fit = gpfit, 
                      locs_pred = data[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data)))
```

```{r}
krige_values_test <- GpGp::predictions(fit = gpfit, 
                      locs_pred = data_test[,c("longitude", "latitude")], 
                      X_pred = rep(1,nrow(data_test)))
```

```{r}
# OR_pred <- predict(OR_mean_model, newdata = data_test)$pred
OR_pred_test <- predict(OR_mean_model_xgboost, newdata = design_matrix_test)
OR_pred <- predict(OR_mean_model_xgboost, newdata = design_matrix)
```

```{r}
plot(log(data_test$concentration_plus_median)- OR_pred, spline_test_values, main = 'Spline predictions')
abline(0,1,col='red')
```

```{r}
plot(log(data_test$concentration_plus_median)- OR_pred_test, krige_values_test, main = 'GP predictions')
abline(0,1,col='red')
```

```{r}
plot(log(data_test$concentration_plus_median), OR_pred, cex = 0.1)
abline(0, 1, col='red')

plot(log(data_test$concentration_plus_median), OR_pred + krige_values_test, cex = 0.1)
abline(0, 1, col='red')
```

rmse for test data

```{r}
Metrics::rmse(log(data_test$concentration_plus_median), OR_pred_test)
Metrics::rmse(log(data_test$concentration_plus_median), krige_values_test + OR_pred_test)
```

rmse for training data

```{r}
# rmse(log(data$concentration_plus_median), OR_mean_model$library.predict)
# rmse(log(data$concentration_plus_median), data$Gp_pred + OR_mean_model$library.predict)
Metrics::rmse(log(data$concentration_plus_median), OR_pred)
Metrics::rmse(log(data$concentration_plus_median), krige_values + OR_pred)
```


##### Estimate Generalized Propensity Score

```{r}
data$krige_adjust <- krige_values
data_test$krige_adjust <- krige_values_test
```


```{r}
# GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","krige_adjust")

GPS_covariate_names <- c("crop_type_combine","drainagecl","precipitation","cafolog")
# "StaticLevel","WellUse" seems to be very useful?
# GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
# GPS_covariate_names <- c("StaticLevel","WellUse","crop_type_combine","drainagecl","precipitation","cafolog","latitude","longitude")
## Based on results, directly adding latitude, longitude tend to bring over-fitting
```

###### Covariate Balancing

```{r}
library(independenceWeights)
```

```{r}
weights <- independence_weights(A = data$logWellDepth, X = data[, GPS_covariate_names])
```

It is very slow and there are some problem with it.

###### Nornal density

We assume the log of well depth is normally distributed with its mean and variance depending on covariates.


Following: directly using SuperLearner
Modeling the conditional mean (with ranger)

```{r}
gps_mean_model <- SuperLearner(
  Y = data$logWellDepth,
  X = data[, GPS_covariate_names],
  family = gaussian(),
  SL.library = c("SL.ranger")
)
```

Plot the fitted value in training data

```{r}
# visualize the residual of the prediction
plot(data$logWellDepth, gps_mean_model$SL.predict)
abline(0,1,col='red')
```

Plot the prediction in test data

```{r}
# visualize the residual of the prediction
plot(data_test$logWellDepth, predict(gps_mean_model, newdata = data_test[, GPS_covariate_names])$pred)
abline(0,1,col='red')
```

```{r}
plot(data$logWellDepth - gps_mean_model$Z)
qqnorm(data$logWellDepth - gps_mean_model$Z)
qqline(data$logWellDepth - gps_mean_model$Z, col = "red") # Adds the reference line
```

```{r}
gps_sigma = sd(data$logWellDepth - gps_mean_model$Z)
gps_sigma
```

We can also further model the heteroskedasticity by fitting the log squared residual to the covariance

##### Fit Gp for well depth

```{r}
depth_residual_df <- data.frame(depth_residual = data$logWellDepth - gps_mean_model$Z, longitude = data$longitude, latitude = data$latitude)
```

```{r}
depth_residual_sf <- st_as_sf(depth_residual_df, coords = c('longitude','latitude'), crs = 4326)
```

```{r}
gpfit_depth <- GpGp::fit_model(y = depth_residual_df$depth_residual, 
                          locs = depth_residual_df[,c("longitude", "latitude")],
                          covfun_name = "matern_sphere")
```

```{r}
depth_krige_values <- leave_one_out_kriging(locs = depth_residual_df[,c("longitude", "latitude")], 
                                      y_obs = depth_residual_df$depth_residual, 
                                      gp_model = gpfit_depth$covfun_name,
                                      gp_params = gpfit_depth$covparms,
                                      order = c("coordinate")) 
```


```{r}
depth_krige_values_test <- GpGp::predictions(fit = gpfit_depth, locs_pred = data_test[,c("longitude","latitude")], X_pred = rep(1, nrow(data_test)))
```

```{r}
depth_pred <-  gps_mean_model$SL.predict
depth_pred_test <- predict(gps_mean_model, data_test[, GPS_covariate_names])$pred
```

```{r}
plot(depth_krige_values_test, data_test$logWellDepth - depth_pred_test, main='depths residuals v.s. Gp prediction')
abline(0,1,col='red')
plot(data_test$logWellDepth,depth_pred_test)
abline(0,1,col='red')
plot(data_test$logWellDepth,depth_pred_test + depth_krige_values_test)
abline(0,1,col='red')
```
For training data
```{r}
Metrics::rmse(data$logWellDepth,depth_pred + depth_krige_values)
Metrics::rmse(data$logWellDepth,depth_pred)
```

For test data
```{r}
Metrics::rmse(data_test$logWellDepth,depth_pred_test + depth_krige_values_test)
Metrics::rmse(data_test$logWellDepth,depth_pred_test)
```

```{r}
depth_residual_df_test <- data.frame(depth_residual_gp = data_test$logWellDepth - depth_pred_test - depth_krige_values_test, longitude = data_test$longitude, latitude = data_test$latitude)
```

```{r}
depth_residual_sf_test <- st_as_sf(depth_residual_df_test, coords = c('longitude','latitude'), crs = 4326)
```

Let's plot the residuals

```{r}
qq_plot <- ggplot(depth_residual_df_test, aes(sample =  depth_residual_gp)) +
  stat_qq(color = "blue", alpha = 0.7) +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(
    title = "Normal Q-Q Plot of Depth Residual",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles (Residuals)"
  ) +
  theme_minimal()

depth_residual_df_test$index = 1:nrow(depth_residual_df_test)
depth_residual_df_test$logWellDepth = data_test$logWellDepth
  
index_plot <- ggplot(depth_residual_df_test, aes(x = logWellDepth, y = depth_residual_gp)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residuals vs. Well Depth",
    x = "log Well Depth",
    y = "Residuals"
  ) +
  theme_minimal()

print(qq_plot)
print(index_plot)
```

##### Test for spatial unconfoundedness: Moran Test 

```{r}
depth_residual_df <- data.frame(depth_residual = data$logWellDepth - gps_mean_model$Z, longitude = data$longitude, latitude = data$latitude)
```

```{r}
depth_residual_sf <- st_as_sf(depth_residual_df, coords = c('longitude','latitude'), crs = 4326)
```

```{r}
usa_states <- ne_states(country = "united states of america", returnclass = "sf")
wisconsin <- usa_states[usa_states$name == "Wisconsin", ]
```


```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "white", color = "black") +  # Wisconsin boundary
  geom_sf(data=depth_residual_sf, aes(color= depth_residual),size=0.1) +
  scale_color_material_c()
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()
```

```{r}
# Find the nearest neighbors for each point
neighbors <- knn2nb(knearneigh(depth_residual_sf, k = 10))
weights <- nb2listw(neighbors, style = "W") # 'W' means row-standard
```

```{r}
# Test the spatially smooth variable
moran.test(depth_residual_sf$depth_residual, listw = weights)
```


```{r}
ggplot() +
  geom_sf(data = wisconsin, fill = "white", color = "black") +  # Wisconsin boundary
  geom_sf(data=depth_residual_sf_test, aes(color= depth_residual_gp),size=0.1) +
  scale_color_material_c()
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()
```


#### Pre-compute Fitted Outcome Mean for integral
##### For traning data

The range of the treatment

```{r}
# depth_range <- c(min(log(data_full$WellDepth)), max((log(data_full$WellDepth))))
# since the depth can be 0
# depth_range <- c(2, max((log(data_full$WellDepth))))
```


```{r}
# xgboost use the design matrix
cf_predictions <- precompute_cf_predictions(data_obs = design_matrix, 
                              fit_outcome = OR_mean_model_xgboost,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
i = 3
plot(cf_predictions$treatment_values, cf_predictions$pred_matrix[,i])
```

```{r}
smoothers <- make_prediction_smoothers(treatment_grid = cf_predictions$treatment_values,
                                            pred_matrix_cf = cf_predictions$pred_matrix,
                                            # smooth_method = "gam",
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers <- make_cumint_smoothers(treatment_grid = cf_predictions$treatment_values,
                                         pred_matrix_cf = cf_predictions$pred_matrix,
                                         smooth_method = "splinefun",
                                         # smooth_method = "gam",
                                         # smooth_method = "smooth.spline",
                                         spar = NULL)
```

```{r}
d_cumint_smoothers <- make_d_cumint_smoothers(treatment_grid = cf_predictions$treatment_values,
                                               pred_matrix_cf = cf_predictions$pred_matrix,
                                               # smooth_method = "splinefun",
                                               # smooth_method = "gam",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```

I feel like the trend is very similar, only the referenced level are different.

Also, since the outcome mean prediction, which is the slope of the integral, changes a little, the integral looks like a line

```{r}
i = 7

plot(cf_predictions$treatment_values, cf_predictions$pred_matrix[,i])

plot(cf_predictions$treatment_values, sapply(cf_predictions$treatment_values, function(t){smoothers[[i]](t)}), cex = 0.1)

plot(cf_predictions$treatment_values, sapply(cf_predictions$treatment_values, function(t){cumint_smoothers[[i]](t)}), cex = 0.1)

plot(cf_predictions$treatment_values, sapply(cf_predictions$treatment_values, function(t){d_cumint_smoothers[[i]](t)}), cex = 0.1)
```

```{r}
i = 5
plot(seq(3.1,6.9,by = 0.01), sapply(seq(3.1,6.9,by = 0.01), function(t){smoothers[[i]](t)}), cex = 0.1)
```

##### For test data

```{r}
# for superlearner
cf_predictions_test <- precompute_cf_predictions(data_obs = data_test[,OR_covariate_names], 
                              fit_outcome = OR_mean_model,
                              treatment_range = c(min(log(data_full$WellDepth)), max((log(data_full$WellDepth)))),
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
# xgboost use the design matrix
cf_predictions_test <- precompute_cf_predictions(data_obs = design_matrix_test, 
                              fit_outcome = OR_mean_model_xgboost,
                              treatment_range = depth_range,
                              treatment_step = 0.02,
                              treatment_name = "logWellDepth")
```

```{r}
# xgboost is already monotone, so we don't need isotomic adjustment
cf_predictions_test <- isotomic_correction(treatment_grid = cf_predictions_test$treatment_values,
                                          pred_matrix_cf = cf_predictions_test$pred_matrix)
```

```{r}
i = 9
plot(cf_predictions_test$treatment_values, cf_predictions_test$pred_matrix[,i])
```

```{r}
smoothers_test <- make_prediction_smoothers(treatment_grid = cf_predictions_test$treatment_values,
                                            pred_matrix_cf = cf_predictions_test$pred_matrix,
                                            smooth_method = "smooth.spline",
                                            spar = NULL)
```

```{r}
cumint_smoothers_test <- make_cumint_smoothers(treatment_grid = cf_predictions_test$treatment_values,
                                               pred_matrix_cf = cf_predictions_test$pred_matrix,
                                               # smooth_method = "splinefun",
                                               smooth_method = "smooth.spline",
                                               spar = NULL)
```

```{r}
i = 5
plot(cf_predictions_test$treatment_values, cf_predictions_test$pred_matrix[,i])
```

```{r}
i = 5
plot(seq(3.1,6.9,by = 0.01), sapply(seq(3.1,6.9,by = 0.01), function(t){smoothers_test[[i]](t)}), cex = 0.1)
```

I think maybe we should estimate the GP after smoothing the outcome regression

#### Indirect Method

```{r}
threshold_val = log(10)
# depth_range <- c(min(log(data_full$WellDepth)), max((log(data_full$WellDepth))))
```

##### For Training Data

```{r}
indirect_policy_nonsp <- sapply(1:nrow(data), function(i) {
  if(smoothers[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
indirect_policy <- sapply(1:nrow(data), function(i) {
  if(smoothers[[i]](depth_range[1]) + krige_values[i]<= threshold_val) return(depth_range[1])
  if(smoothers[[i]](depth_range[2]) + krige_values[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers[[i]](t) + krige_values[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

##### For Test Data

```{r}
indirect_policy_nonsp_test <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test[[i]](depth_range[1])<= threshold_val) return(depth_range[1])
  if(smoothers_test[[i]](depth_range[2]) > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test[[i]](t)- threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

```{r}
indirect_policy_test <- sapply(1:nrow(data_test), function(i) {
  if(smoothers_test[[i]](depth_range[1]) + krige_values_test[i]<= threshold_val) return(depth_range[1])
  if(smoothers_test[[i]](depth_range[2]) + krige_values_test[i] > threshold_val) return(depth_range[2])
  res <- uniroot(f = function(t) {smoothers_test[[i]](t) + krige_values_test[i] - threshold_val},
        lower = depth_range[1],
        upper = depth_range[2])
  return(res$root)
})
```

##### Metrics for Training data

Let's see the metrics of indirect policy

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy > data$logWellDepth)

Metrics::f1(log(data$concentration_plus_median) > threshold_val, indirect_policy > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy > data$logWellDepth))
```

Let's see the metrics of indirect policy nonsp

```{r}
Metrics::accuracy(log(data$concentration_plus_median) > threshold_val, indirect_policy_nonsp > data$logWellDepth)

Metrics::f1(log(data$concentration_plus_median) > threshold_val,indirect_policy_nonsp > data$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_nonsp > data$logWellDepth))
```

##### Metrics for Test Data

spatial indirect

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_test > data_test$logWellDepth)

Metrics::f1(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_test > data_test$logWellDepth))
```


non-spatial indirect

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val, indirect_policy_nonsp_test > data_test$logWellDepth)

Metrics::f1(log(data_test$concentration_plus_median) > threshold_val,indirect_policy_nonsp_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( indirect_policy_nonsp_test > data_test$logWellDepth))
```


#### Direct Method: Outcome Regression Parametric Policy

We have already define the design_matrix_test and design_matrix when fitting xgboosting

We need to exclude the well depth from it

```{r}
policy_covariate_names <- c("WellUse","crop_type_combine","drainagecl","precipitation","cafolog","longitude","latitude")
# policy_covariate_names <- c("WellUse","crop_type_combine","drainagecl","precipitation","cafolog")
```

```{r}
model_temp <- mgcv::gam(concentration_plus_median~s(longitude,latitude,k=50,bs='tp',fx=TRUE) + crop_type_combine + s(precipitation) + s(cafolog) + drainagecl, data=data, method="REML")

```

```{r}
design_matrix_policy <- model.matrix(model_temp)
design_matrix_policy_test <- predict(model_temp, newdata = data_test, type = "lpmatrix")
```

```{r}
design_matrix_policy <- model.matrix(~ . , data = data[,policy_covariate_names])
design_matrix_policy_test <- model.matrix(~ . , data = data_test[, policy_covariate_names])
```

```{r}
design_matrix_gp <- cbind(design_matrix, gp_adjust = krige_values)
design_matrix_test_gp <- cbind(design_matrix_test, gp_adjust = krige_values_test)
```

initialize using

```{r}
# initial_lm <- lm(indirect_policy~ -1 + design_matrix_gp)
# coef(initial_lm)
initial_lm <- lm(indirect_policy~ -1 + design_matrix_policy)

```

```{r}
plot(indirect_policy, initial_lm$fitted.values)
```


```{r}
# design_matrix <- model.matrix(~ X1 + X2 + X3 + X4, cbind(data$X))
# the derivative that I write here is wrong with large probability
coefs <- coef(initial_lm)
parametric_optim_result_or <- optim(par = coef(initial_lm),
      fn = function(coefs){
        assigned_trt <- design_matrix_policy %*% coefs
        return(compute_total_loss_db_sum_smooth(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data),
                                 T = data$logWellDepth,
                                 krige_adjust = rep(0, nrow(data)),
                                 outcome_resid = rep(0, nrow(data)),
                                 propensity_est = rep(1,nrow(data)),
                                 cumint_smoothers = cumint_smoothers,
                                 trt_bounds = depth_range,
                                 threshold_val = threshold_val,
                                 kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                                 clip_epsilon = 1,
                                 loss_type = "Integral",
                                 surrogate_type = "Gaussian"))},
      method = "L-BFGS-B",
      # gr = function(coefs){t(design_matrix_policy) %*% d_compute_total_loss_db_smooth(assigned_trt = assigned_trt,
      #                            subject_idx = 1:nrow(data),
      #                            T = data$logWellDepth,
      #                            krige_adjust = rep(0, nrow(data)),
      #                            outcome_resid = rep(0, nrow(data)),
      #                            propensity_est = rep(1,nrow(data)),
      #                            smoothers = smoothers,
      #                            trt_bounds = depth_range,
      #                            threshold_val = threshold_val,
      #                            kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
      #                            clip_epsilon = 1,
      #                            surrogate_type = "Gaussian")},
      control = list(trace = 3))
```

```{r}
direct_policy_parametric_or <- design_matrix_gp %*% parametric_optim_result_or$par
direct_policy_parametric_or_test <- design_matrix_test_gp %*% parametric_optim_result_or$par
```

Training data

```{r}
plot(indirect_policy, direct_policy_parametric_or)
```

Test Data

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val,direct_policy_parametric_or_test > data_test$logWellDepth)

Metrics::f1(log(data_test$concentration_plus_median) > threshold_val, direct_policy_parametric_or_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( direct_policy_parametric_or_test > data_test$logWellDepth))
```

I think a parametric policy seems to be insufficient in this case
We will try the RKHS

#### IPW Parametric Policy

```{r}
# design_matrix <- model.matrix(~ X1 + X2 + X3 + X4, cbind(data$X))

# propensity_est <- dnorm(x = assigned_trt, mean = gps_mean_model$SL.predict, sd = gps_sigma)
# coefs <- coef(initial_lm)
# design_matrix_gp %*% coefs

parametric_optim_result_ipw <- optim(par = coef(initial_lm),
      fn = function(coefs){
        assigned_trt <- design_matrix_gp %*% coefs
        return(compute_total_loss_sum_smooth(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data),
                                 T = data$logWellDepth,
                                 krige_adjust = krige_values,
                                 outcome_resid = data$logconcentration_plus_median,
                                 propensity_est = as.numeric(dnorm(x = assigned_trt, 
                                                        mean = gps_mean_model$SL.predict, sd = gps_sigma)),
                                 cumint_smoothers = cumint_smoothers,
                                 trt_bounds = depth_range,
                                 threshold_val = threshold_val,
                                 kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                                 clip_epsilon = 1,
                                 loss_type = "ipw",
                                 surrogate_type = "Gaussian"))},
      method = "L-BFGS-B",
      gr = function(coefs){t(design_matrix_gp) %*% d_compute_total_loss_smooth(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data),
                                 T = data$logWellDepth,
                                 krige_adjust = krige_values,
                                 outcome_resid = data$logconcentration_plus_median,
                                 propensity_est = dnorm(x = assigned_trt, 
                                                        mean = gps_mean_model$SL.predict, sd = gps_sigma),
                                 smoothers = smoothers,
                                 trt_bounds = depth_range,
                                 threshold_val = threshold_val,
                                 kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                                 clip_epsilon = 1,
                                 surrogate_type = "Gaussian",
                                 loss_type = "ipw")},
      control = list(trace = 3))
```
##### Test the derivative for individual observation

```{r}

subject_idx <- 2

loss_type <- "ipw"

y_d <- d_compute_total_loss_smooth(assigned_trt = seq(depth_range[1]-2,depth_range[2]+2, length.out = 100),
                           subject_idx = rep(1,100),
                           T = data$logWellDepth,
                           krige_adjust = krige_values,
                           outcome_resid = data$logconcentration_plus_median,
                           propensity_est = dnorm(x = assigned_trt, 
                                                  mean = gps_mean_model$SL.predict, sd = gps_sigma),
                           smoothers = smoothers,
                           trt_bounds = depth_range,
                           threshold_val = threshold_val,
                           kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                           clip_epsilon = 1,
                           surrogate_type = "Gaussian",
                           loss_type = loss_type)

h = 0.01
y_plus <- compute_total_loss_smooth(assigned_trt = seq(depth_range[1] - 2,depth_range[2] + 2, length.out = 100) + h,
                         subject_idx = rep(1,100),
                         T = data$logWellDepth,
                         krige_adjust = krige_values,
                         outcome_resid = data$logconcentration_plus_median,
                         propensity_est = as.numeric(dnorm(x = assigned_trt, 
                                                mean = gps_mean_model$SL.predict, sd = gps_sigma)),
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         threshold_val = threshold_val,
                         kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                         clip_epsilon = 1,
                         loss_type = loss_type,
                         surrogate_type = "Gaussian")

y_minus <- compute_total_loss_smooth(assigned_trt = seq(depth_range[1] - 2,depth_range[2] + 2, length.out = 100) - h,
                         subject_idx = rep(1,100),
                         T = data$logWellDepth,
                         krige_adjust = krige_values,
                         outcome_resid = data$logconcentration_plus_median,
                         propensity_est = as.numeric(dnorm(x = assigned_trt, 
                                                mean = gps_mean_model$SL.predict, sd = gps_sigma)),
                         cumint_smoothers = cumint_smoothers,
                         trt_bounds = depth_range,
                         threshold_val = threshold_val,
                         kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                         clip_epsilon = 1,
                         loss_type = loss_type,
                         surrogate_type = "Gaussian")
```

```{r}
plot(seq(depth_range[1],depth_range[2], length.out = 100), y_d)
plot(seq(depth_range[1],depth_range[2], length.out = 100), (y_plus -y_minus)/(2*h))
```


```{r}
# Create a sequence of x-values to test
x_vals <- seq(-3, 3, by = 0.1)

# Calculate the analytical derivative for all x-values
y_analytical <- g_analytical(x_vals)

# Calculate the numerical derivative for all x-values
y_numerical <- (f(x_vals + h) - f(x_vals - h)) / (2 * h)
```


```{r}
# design_matrix <- model.matrix(~ X1 + X2 + X3 + X4, cbind(data$X))

# propensity_est <- dnorm(x = assigned_trt, mean = gps_mean_model$SL.predict, sd = gps_sigma)
coefs <- coef(initial_lm)
parametric_optim_result_ipw <- optim(par = coef(initial_lm),
      fn = function(coefs){
        assigned_trt <- design_matrix_gp %*% coefs
        return(compute_total_loss_sum_smooth(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data),
                                 T = data$logWellDepth,
                                 krige_adjust = krige_values,
                                 outcome_resid = data$logconcentration_plus_median,
                                 propensity_est = dnorm(x = assigned_trt, 
                                                        mean = gps_mean_model$SL.predict, sd = gps_sigma),
                                 cumint_smoothers = cumint_smoothers,
                                 trt_bounds = depth_range,
                                 threshold_val = threshold_val,
                                 kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                                 clip_epsilon = 1,
                                 loss_type = "ipw",
                                 surrogate_type = "Gaussian"))},
      method = "L-BFGS-B",
      gr = function(coefs){t(design_matrix_gp) %*% d_compute_total_loss_smooth(assigned_trt = assigned_trt,
                                 subject_idx = 1:nrow(data),
                                 T = data$logWellDepth,
                                 krige_adjust = krige_values,
                                 outcome_resid = data$logconcentration_plus_median,
                                 propensity_est = dnorm(x = assigned_trt, 
                                                        mean = gps_mean_model$SL.predict, sd = gps_sigma),
                                 smoothers = smoothers,
                                 trt_bounds = depth_range,
                                 threshold_val = threshold_val,
                                 kernel_bw = 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5),
                                 clip_epsilon = 1,
                                 surrogate_type = "Gaussian",
                                 loss_type = "ipw")},
      control = list(trace = 3))
```

```{r}
direct_policy_parametric_ipw <- design_matrix_gp %*% parametric_optim_result_ipw$par
direct_policy_parametric_ipw_test <- design_matrix_test_gp %*% parametric_optim_result_ipw$par
```

Training data

```{r}
plot(indirect_policy, direct_policy_parametric_ipw)
```

```{r}
Metrics::accuracy(log(data_test$concentration_plus_median) > threshold_val,direct_policy_parametric_ipw_test > data_test$logWellDepth)

Metrics::f1(log(data_test$concentration_plus_median) > threshold_val, direct_policy_parametric_ipw_test > data_test$logWellDepth)
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( direct_policy_parametric_ipw_test > data_test$logWellDepth))
```

#### Doubly Robust Parametric Policy

```{r}
# doubly robust pseudo outcome
pesudo_outcome = OR_mean_model$library.predict + ((log(data$concentration_plus_median) - OR_mean_model$library.predict)/gps_obj_2$.data$gps)
```

```{r}
# Dummy coding categorical variables
X <- model.matrix(~ . , data = data[, GPS_covariate_names])
```

```{r}
loss_nitrate <- function(X,coefs, data){
  # get predicted 
  policy = X %*% coefs
  
  # get tmp data
  data.tmp = data
  data.tmp$logWellDepth = policy
  # get fitted outcome regression
  or_pred = predict(OR_mean_model, newdata = data.tmp)$pred
  
  gps_mean = predict(gps_mean_model, newdata = data.tmp[, GPS_covariate_names])$pred
  # get fitted GPS
  gps_mean = predict(gps_mean_model, newdata = data.tmp[, GPS_covariate_names])$pred
  
  gps_est = sapply(gps_mean - policy, function(i){dnorm(1, mean = i,sd = gps_sigma)})
  # get db pseudo-outcome
  
}
```


```{r}
pesudo_outcome_nitrate = function()
  
  (resids[col_idx][!index_left_clip]/ prop_estimation[col_idx][!index_left_clip]) * kernel_integral_vectorized(trt = trt_assigned_right_clipped[!index_left_clip], trt_obs = data[col_idx,"trt"][!index_left_clip], bandwidth = bandwidth)
```


```{r}
plot(log(data$concentration_plus_median),pesudo_outcome,xlim = c(-1,5),ylim = c(-5,10))
abline(0,1,col='red')
```

##### Least Square Error

$$(T - \text{pseudo_outcome})^2$$

```{r}

```






#### Direct Method: location RKHS

```{r}
gps_mean_model
gps_sigma
```

We will use location as the covariates.

```{r}
latlon_matrix <- data[,c('longitude','latitude')]
latlon_matrix_test <- data_test[,c('longitude','latitude')]
# we need to scale the continuous variables
mean_lon <- mean(data[,c('longitude')])
sd_lon <- sd(data[,c('longitude')])
mean_lat <- mean(data[,c('latitude')])
sd_lat <- sd(data[,c('latitude')])
```

```{r}
latlon_matrix[, 'longitude'] <- (latlon_matrix[, 'longitude'] - mean_lon)/sd_lon
latlon_matrix[, 'latitude'] <- (latlon_matrix[, 'latitude'] - mean_lat)/sd_lat
# standardize for test data
latlon_matrix_test[, 'longitude'] <- (latlon_matrix_test[, 'longitude'] - mean_lon)/sd_lon
latlon_matrix_test[, 'latitude'] <- (latlon_matrix_test[, 'latitude'] - mean_lat)/sd_lat
latlon_matrix <- as.matrix(latlon_matrix)
latlon_matrix_test <- as.matrix(latlon_matrix_test)
```

```{r}
# to reduce the computational burden, we choose landmarks 
landmark_indices <- sample(1:nrow(data), 1000)
landmarks <- latlon_matrix[landmark_indices,]

gamma <- median(rdist(latlon_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))

k_matrix <- kernelMatrix(rbf, latlon_matrix, landmarks)
# build kernel matrix for test data
k_matrix_test <- kernelMatrix(rbf, latlon_matrix_test, landmarks)
k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
```


```{r}
initial_RKHS_cv <- cv.glmnet(k_matrix[,-1], indirect_policy_nonsp, alpha = 0) 
```

```{r}
coef(initial_RKHS_cv)
```

```{r}
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.numeric(k_matrix %*% coef(initial_RKHS_cv, s = "lambda.min")) > data$logWellDepth))
```

```{r}
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( as.numeric(k_matrix_test %*% coef(initial_RKHS_cv, s = "lambda.min")) > data_test$logWellDepth))
```

```{r}
max_iter <- 7
tol <- 1e-3
clip_epsilon_bar <- 100
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)

surrogate_type <- "Gaussian"

loss_type <- "ipw"
resids <- data$logconcentration_plus_median

# loss_type <- 'or'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict - krige_values

# loss_type <- 'db'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict - krige_values


gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

params_old <- as.numeric(coef(initial_RKHS_cv, s = "lambda.min"))
params <- as.numeric(coef(initial_RKHS_cv, s = "lambda.min"))

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = params_old, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = params_old, 
                        K = k_matrix, 
                        T = data$logWellDepth,
                        krige_adjust = rep(0, nrow(data)),
                        outcome_resid = resids,
                        propensity_est = gps_est,
                        lambda = 10, 
                        smoothers = smoothers, 
                        cumint_smoothers = cumint_smoothers, 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = log(10), 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```

```{r}
yardstick::mcc_vec(truth = as.factor(log(data_test$concentration_plus_median) > threshold_val), estimate = as.factor( as.numeric(k_matrix_test %*% params_old) > data_test$logWellDepth))
```



#### Direct Method: covariate RKHS


```{r}
library(kernlab)
library(glmnet)
```

##### get the kernel matrices


```{r}
design_matrix <- model.matrix(~ . , data = data[, OR_covariate_names])
design_matrix_test <- model.matrix(~ . , data = data_test[, OR_covariate_names])
# we need to scale the continuous variables
mean_logWellDepth <- mean(design_matrix[, 2])
sd_logWellDepth <- sd(design_matrix[, 2])

mean_precipitation <- mean(design_matrix[, 3])
sd_precipitation <- sd(design_matrix[, 3])

mean_cafolog <- mean(design_matrix[, 4])
sd_cafolog <- sd(design_matrix[, 4])
```

```{r}
design_matrix[, 2] <- (design_matrix[, 2] - mean_logWellDepth)/sd_logWellDepth
design_matrix[, 3] <- (design_matrix[, 3] - mean_precipitation)/sd_precipitation
design_matrix[, 4] <- (design_matrix[, 4] - mean_cafolog)/sd_cafolog
# standardize for test data
design_matrix_test[, 2] <- (design_matrix_test[, 2] - mean_logWellDepth)/sd_logWellDepth
design_matrix_test[, 3] <- (design_matrix_test[, 3] - mean_precipitation)/sd_precipitation
design_matrix_test[, 4] <- (design_matrix_test[, 4] - mean_cafolog)/sd_cafolog
```

```{r}
gamma <- median(rdist(design_matrix))
rbf <- rbfdot(sigma = 1/(gamma^2))


landmark_indices <- sample(1:nrow(data), 2000)
landmarks <- design_matrix[landmark_indices,]

k_matrix <- kernelMatrix(rbf, design_matrix, landmarks)
# build kernel matrix for test data
k_matrix_test <- kernelMatrix(rbf, design_matrix_test, landmarks)

k_matrix <- cbind(rep(1,nrow(k_matrix)), k_matrix)
k_matrix_test <- cbind(rep(1,nrow(k_matrix_test)), k_matrix_test)
```

```{r}
initial_RKHS_cv <- cv.glmnet(k_matrix[,-1], indirect_policy_nonsp, alpha = 0) 
```


##### dc algorithm

```{r}
max_iter <- 30
tol <- 1e-4
clip_epsilon_bar <- 100
clip_epsilon <- 1
kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)
depth_range <- c(min(data$logWellDepth), max(data$logWellDepth))

surrogate_type <- "Gaussian"

loss_type <- "ipw"
resids <- data$logconcentration_plus_median

# loss_type <- 'or'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict - krige_values

# loss_type <- 'db'
# resids <- data$Y - outcome_mean_nonsp_model$SL.predict - krige_values


gps_est <- dnorm(data$logWellDepth, mean = gps_mean_model$SL.predict, sd = gps_sigma)

params_old <- as.numeric(coef(initial_RKHS_cv, s = "lambda.min"))
params <- as.numeric(coef(initial_RKHS_cv, s = "lambda.min"))

for (k in 1:max_iter) {
    
    # Solve the convex subproblem
    cat(k,'-th iteration of DC algorithm','\n')
    sub_result <- optim(par = params_old, 
                        fn = compute_total_loss_smooth_dc_approx, 
                        gr = d_compute_total_loss_smooth_dc_approx, # we try the case with/without gr
                        params_old = params_old, 
                        K = k_matrix, 
                        T = data$logWellDepth,
                        krige_adjust = rep(0, nrow(data)),
                        outcome_resid = resids,
                        propensity_est = gps_est,
                        lambda = 10, 
                        smoothers = smoothers, 
                        cumint_smoothers = cumint_smoothers, 
                        kernel_bw = kernel_bw, 
                        trt_bounds = depth_range,
                        threshold_val = log(10), 
                        clip_epsilon = clip_epsilon,
                        clip_epsilon_bar = clip_epsilon_bar,
                        surrogate_type = surrogate_type,
                        loss_type = loss_type,
                        method = "L-BFGS-B", 
                        control = list(maxit = 60, trace = 3))
    
    params <- sub_result$par
    
    # 3. Check for convergence
    change <- sqrt(sum((params_old - params)^2))
    cat(sprintf("Iter: %d, Change: %f, Objective: %f\n", k, change, sub_result$value))
    
    if (change < tol) {
      cat("--- DC Algorithm Converged ---\n")
      params_old <- params
      break
    }
    
    params_old <- params
  }
```

```{r}
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.numeric(k_matrix %*% coef(initial_RKHS_cv, s = "lambda.min")) > data$logWellDepth))
```

```{r}
yardstick::mcc_vec(truth = as.factor(log(data$concentration_plus_median) > threshold_val), estimate = as.factor( as.numeric(k_matrix %*% params_old) > data$logWellDepth))
```

#### Visualize loss

```{r}
surrogate_type <- "Gaussian"
loss_type <- "ipw"

clip_epsilon_bar <- 100
clip_epsilon <- 1

kernel_bw <- 1.06 * sd(data$logWellDepth) *nrow(data)^(-1/5)

knots <- seq(depth_range[1], depth_range[2], length.out = 100)

which(data$logconcentration_plus_median>=log(20))[1] # 309
which(data$logconcentration_plus_median>=log(30))[1] # 735
which(data$logconcentration_plus_median>=log(30))[2] # 5037
i <- 5037

plot(knots, sapply(knots, compute_total_loss_smooth,  subject_idx = i,
                            T = data$logWellDepth,
                            krige_adjust = rep(0, nrow(data)),
                            outcome_resid = data$logconcentration_plus_median,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = depth_range,
                            threshold_val = log(10),
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

plot(knots, sapply(knots, compute_total_loss_smooth_plus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type) - sapply(knots, compute_total_loss_smooth_minus,  subject_idx = i,
                             T = data$T,
                            krige_adjust = krige_values,
                            outcome_resid = data$Y,
                            propensity_est = gps_est,
                            cumint_smoothers = cumint_smoothers,
                            trt_bounds = trt_bounds,
                            threshold_val = threshold_val,
                            kernel_bw = kernel_bw,
                            clip_epsilon = 1,
                            clip_epsilon_bar,
                            surrogate_type,
                            loss_type))

```
```

#### use spatial neural network

```{r}
library(reticulate)
```


```{r}
reticulate::use_condaenv("policy_learning", required = TRUE)
use_condaenv("policy_learning_arm64", required = TRUE)
```

```{python}
import torch
```

